### 2015-01-15 doc/tutorial/README.Rmd
small changes

# Tutorial

**SVMBridge** is a small framework for the controlling several SVM packages like LIBSVM, SVMperf, CVM from inside R. 
The goal is to allow for a rather easy way to do experiments with these different SVM packages. As the framework interacts with these packages by I/O, the focus is not on speed.

In the following we provide a short introduction to **SVMBridge**. An introductory example serves as a quickstart guide.
Note that our focus is on your comprehension of the basic functions and
applications. For detailed technical information and manual pages, please refer to
the package's [manual pages](http://aydindemircioglu.github.io/SVMBridge/man/). They are regularly updated and reflect the documentation of the current package.

1. [Quickstart](quickstart.md)
2. [In Depth Introduction](introduction.md)
3. [Wrappers](wrappers.md)

### 2015-01-15 R/LASVM_wrapper.R

@@ -1,10 +1,5 @@
 #!/usr/bin/Rscript  --vanilla 
 
-source ("software/helpers/universalWrapper.R")
-
-library(BBmisc)
-
-
 
 
 evalLASVM = function(...)  {   

### 2015-01-15 R/SVMperf_wrapper.R

@@ -1,11 +1,5 @@
 #!/usr/bin/Rscript  --vanilla 
 
-source ("software/helpers/universalWrapper.R")
-
-library(BBmisc)
-
-
-
 
 evalSVMperf = function(...)  {   
     universalWrapper (

### 2015-01-28 doc/tutorial/introduction.Rmd
objectify 1/3

Writing your own wrapper is also possible, and not too complicated. You must provide several callback functions. The most complex part will be provide functions to read and write the model. As a starter, you might want to look into the wrappers, **SVMBridge** provides.

# SVM Parameter

Every SVM has some parameter, tha usually must be tuned. While LIBSVM has (for RBF kernel) two prominent parameter, the regularization term $C$ and the kernel bandwidth $\gamma$, this might not be true for other methods. For example, Budgeted SVM operates with $\lambda$ instead of $C$, with $\lambda = \frac{1}{2 C}$, and the $\gamma$ is twice the gamma of LIBSVM. Because of this, **SVMBridge** does not have a prebuild set of parameters. Every wrapper is free to use the parameter it needs. This entails that to use a wrapper, you must first know which parameter it uses. You can use the help function to get a description of a wrapper.

LIBSVMHelp() { }

# Training

Training can be either performed from data in memory (probably the most useful case), from the file system by using the ```trainSVM``` function. 
Note: In case you specify data from memory and from disc, **SVMBridge** will throw an error, as it will be confused about which source to use.

# Testing

Testing is done very similar to training by calling the ```testSVM``` function You have the option to use data from memory, from file system. Furthermore, predictions can be provided. By default, the predictions are written into a temporary file and are discarded after testing. 
Note: In case you specify data from memory and from disc, **SVMBridge** will throw an error, as it will be confused about which source to use.

# Model file

If you do not specify any model file for training, ```trainSVM``` will ask R for a temporary file and will provide this path to the SVM package. Internally, the ```writeModel``` routine of the SVM package will be used to create a model file from a model in memory.  In case you have not specified a model file path, after training the model will be read from the temporary model file into the SVM Results Object. Contrary, if you did specify a model file path, then by default, the model is not read back into the SVM Results Object. To enforce reading the model, you can set the argument ```readModel``` to true.

At testing time, you must provide the trained model. If you pass this as an in-memory model, ```testSVM``` will call the ```readModel``` routine of the SVM package first to dump the model to a temporary file on the filesystem.

skipModel == TRUE

```splus
library(SVMBridge)
```

# Predictions

After testing, the predictions will again be stored in the SVM Object and can be accessed by e.g. ```SVMObj$predictions```. Depending on the purpose, the predictions are not always needed, for example when tuning the SVM parameters. To avoid extra I/O and memory consumption, reading back the predictions can be turned of by using the ```skipPredictions``` options. By default, ```skipPredictions``` is false, so ```callSVM``` will return the predictions in the SVM Object.

skipPredictions == TRUE

# Writing your own wrapper

The method name will define all function names! This is very important, if you do not abide to this rule, callSVM will not find the corresponding callbacks! Suppose you want to write a wrapper for mySVM. In this case, you will need to create a file exactly called "mySVM_wrapper.R'. Please note that the filename is case-sensitive. This is also true for all the callback functions.

Your mySVM software must be split into two parts: Training and Testing. This might be the same executable, but be aware, that testing must work without training data. In case your package needs testing data at training time (e.g. if the model only specifies the index of the support vector in the training file instead of copying them into the model), you need to perform quite heavy tricks to get this going. In general, using other packages and skip **SVMBridge** might be a good idea, as **SVMBridge** strictly follows the LIBSVM way of training and testing.

## The training parameters

You will get a bunch of parameters. You will need to decide which ones you want to handle. E.g. the LIBSVM wrapper handles $C$, $\gamma$ and $\epsilon$, but also uses $kernelCacheSize$. On the other hand, the LIBSVM wrapper currently has no option to control the shrinking, although LIBSVM has a parameter for this. Similarly you will need to decide, whether you want the user to have control over all parameters or only a relevant subset. The callback will then gather all the parameter and create a command-line string.

# SVM Result Object

Actually we do enforce that.
We do not enforce a structure on the returned SVM Return Object. Nonetheless, we encourage you to use the same structure we have provided in our example wrapper, if that is applicable to your case. The SVM Return Object contains the following fields: 
 -execTime
 -error
Depending on training or testing, this is the training or testing time, measured in wall time by microbenchmark.

## Test parameters

Very similar to training parameters, the test parameter callback will assemble a given model file and test data and possibly other options.

extraParameters for testing??

# Data Formats

As nearly all SVM Packages work with the LIBSVM/SVMlight Sparse Format, this is the basic file format in SVMBridge. Notice that SVMBridge actually does neither enforce nor check this explicitly. You can support your own data format for your own SVM wrapper. In this case, keep in mind that other SVM packages might not be able to work with your data.

To implement your own data format, recall that you have two ways to provide data to your SVM software. The first is by specifiying a path to a data file on the filesystem. In this case, SVMBridge will not explicitly work with the data, so you are free to use any format, as long as your SVM software can handle this. The second way is to pass data as a matrix to the SVMBridge. Here, SVMBridge needs to dump the given data first. By default, the SVMBridge will write the data in the LIBSVM Sparse Format. You are free to overwrite this behaviour, by rewriting the convertData function of the SVM Wrapper Object.

Let us give an easy example.

LABELS..
MULTICLASS..

# Vocabulary

The SVM software is the compiled object on the file system, e.g. when you download and compile LIBSVM you will end  up with two binaries, one for training, one for testing. These files is an SVM software for us.

Glue code that handles controlling the SVM software is called an SVM wrapper. This contains e.g. routines for reading and writing an SVM model.

Finally, an SVM package is for us the internal object in the SVMBridge, that has basically consists of the wrapper and the software (e.g. a path to the binary).

### 2015-01-28 R/LASVM_wrapper.R

@@ -1,113 +1,125 @@
 #!/usr/bin/Rscript  --vanilla 
 
+#
+# SVMBridge 
+#		(C) 2015, by Aydin Demircioglu
+#
+#		LASVM_wrapper.R
+# 
+# SVMBridge is free software: you can redistribute it and/or modify
+# it under the terms of the GNU Lesser General Public License as published
+# by the Free Software Foundation, either version 3 of the License, or
+# (at your option) any later version.
+#
+# SVMBridge is distributed in the hope that it will be useful,
+# but WITHOUT ANY WARRANTY; without even the implied warranty of
+# MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
+# GNU Lesser General Public License for more details.
+#
+# Please do not use this software to destroy or spy on people, environment or things.
+# All negative use is prohibited.
+#
+ 
 
+	
+	createTrainingArguments.LASVM = function (
+		trainDataFile = "",
+        modelFile = "",
+        extraParameter = "",
+        kernelCacheSize = 1024,
+		cost = 1, 
+        useBias = FALSE,
+        gamma = 1,
+        epochs = 1,
+        epsilon = 0.001, 
+        ...) 
+	{
+		# count training examples
+		N = countLines(trainDataFile)
 
-evalLASVM = function(...)  {   
-    universalWrapper (
-        modelName = "LASVM",
-        trainingParameterCallBack = LASVMTrainingParameterCallBack,
-        testParameterCallBack = LASVMTestParameterCallBack,
-        extractInformationCallBack  = LASVMExtractInformationCallBack,
-        trainBinary = LASVMTrainBinary(),
-        testBinary = LASVMTestBinary (),
-        bindir = LASVMBinDir(),
-        ...
-    );
-}
+		biasParameter = "-b 0"
+		if (useBias == TRUE)
+			biasParameter = "-b 1"
 
+		args = c(
+			sprintf("-m %d", kernelCacheSize), # in MB 
+			biasParameter,
+			sprintf("-g %.16f", gamma),
+			sprintf("-c %.16f", cost), 
+			sprintf("-e %.16f", epsilon),
+			sprintf("-p %.16f", epochs),
+			extraParameter,
+			trainDataFile,
+			modelFile
+		)
 
+		return (args)
+	}
 
-LASVMTrainingParameterCallBack = function (trainfile = "",
-                                            modelFile = "",
-                                            extraParameter = "",
[…]

### 2015-01-28 R/LIBSVM_wrapper.R

@@ -4,7 +4,7 @@
 # SVMBridge 
 #		(C) 2015, by Aydin Demircioglu
 #
-#		SVMperf_wrapper.R
+#		LIBSVM_wrapper.R
 # 
 # SVMBridge is free software: you can redistribute it and/or modify
 # it under the terms of the GNU Lesser General Public License as published
@@ -20,244 +20,207 @@
 # All negative use is prohibited.
 #
  
+	
+	createTrainingArguments.LIBSVM = function (x,
+		trainDataFile = "",
+        modelFile = "",
+        extraParameter = "",
+        kernelCacheSize = 1024,
+		cost = 1, 
+        useBias = FALSE,
+        gamma = 1,
+        epsilon = 0.001, 
+        ...) 
+	{
+		args = c(
+			"-s 0",                         # c classification
+			"-t 2",
+			sprintf("-m %d", kernelCacheSize), # in MB 
+			sprintf("-c %.16f", cost),         # rbf kernel
+			sprintf("-g %.16f", gamma),        # gamma
+			sprintf("-e %.16f", epsilon),      # epsilon tolerance
+			extraParameter,
+			trainDataFile,
+			modelFile
+		)
 
-#' LIBSVMTrainBinary 
-#'
-#' return full path of training binary.
-#
-
-LIBSVMTrainBinary <- function() {
-    return ("svm-train")
-}
-
-
-
-#' LIBSVMTestBinary 
-#'
-#' return full path of test binary.
-#
-
-LIBSVMTestBinary <- function() {
-    return ("svm-predict")
-}
-
-
-
-#' LIBSVMTrainBinaryOutputPattern
-#'
-#' return pattern that identifies the training binary.
-#
-
-LIBSVMTrainBinaryOutputPattern <- function() {
-    return ('Usage: svm-train .options. training_set_file .model_file.')
-}
-
-
-
-#' LIBSVMTestBinaryOutputPattern
-#'
-#' return pattern that identifies the test binary.
-#
-
-LIBSVMTestBinaryOutputPattern <- function() {
-    return ('for one-class SVM only 0 is supported')
-}
-
+		return (args)
+	}
 
 	
-#' evalLIBSVM
-#'
-#' @param     ...		parameters that will be passed on training and/or test callbacks
-#
-
-evalLIBSVM = function(...)  {   
-
-	parameterList = list(..., 
-		trainingParameterCallBack = LIBSVMTrainingParameterCallBack,
[…]

### 2015-01-28 R/SVMWrapper.R

@@ -1,0 +1,114 @@
+#!/usr/bin/Rscript  --vanilla 
+#
+# SVMBridge 
+#		(C) 2015, by Aydin Demircioglu
+#
+#		SVMWrapper.R
+# 
+# SVMBridge is free software: you can redistribute it and/or modify
+# it under the terms of the GNU Lesser General Public License as published
+# by the Free Software Foundation, either version 3 of the License, or
+# (at your option) any later version.
+#
+# SVMBridge is distributed in the hope that it will be useful,
+# but WITHOUT ANY WARRANTY; without even the implied warranty of
+# MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
+# GNU Lesser General Public License for more details.
+#
+# Please do not use this software to destroy or spy on people, environment or things.
+# All negative use is prohibited.
+#
+ 
+ 
+
+#' General class for any SVM wrapper.
+#'
+#' Description... 
+#'
+
+    createSVMWrapper = function(method, trainingBinary = '', ...) {
+        
+        # check that trainingBinary exists, if given
+        # ...    
+        
+        # register the newly created object in our environment
+        
+        # create S3 Object
+        svmPackage = makeS3Obj(c(method, "SVMWrapper"),
+    		method = method,
+            trainingBinary = trainingBinary
+    	)
+
+    	# TODO: replace existing object in package list
+    	
+    	# add object to our internal list of created wrappers
+	#	SVMBridgeEnv$packages = c(SVMBridgeEnv$packages, svmPackage)
+		
+		return (svmPackage)
+    }
+    
+
+    
+
+	createTrainingArguments = function (x,
+		trainDataFile = "",
+        modelFile = "",
+        extraParameter = "",
+        kernelCacheSize = 1024,
+		cost = 1, 
+        useBias = FALSE,
+        gamma = 1,
+        epsilon = 0.001, 
+        ...) 
+	{
+		UseMethod("createTrainingArguments")
+	}
+
+	
+	
+	createTestArguments = function (x,
+		testDataFile = "",
+		modelFile = "", 
+		predictionsFile = "",
+		...) 
+	{
+		UseMethod("createTestArguments")
+	}
+
+
+	
+	extractTrainInfo = function (x, output) {
+		UseMethod ("extractTrainInfo")
+	}
+	
+	
+	
+	extractTestInfo = function (x, output) {
[…]

### 2015-01-28 R/addSVMPackage.R

@@ -20,41 +20,44 @@
 #
  
  
-
-
-# every wrapper has its own object that is derived from the universalWrapper.
-# we create all these at initialization time of the package.
-# then, when e.g. paths are updated, we have a nice place to store these.
-
-
-addSVMPackage <- function (filePath = NA, softwarePath = NA, verbose = FALSE)
+ 
+addSVMPackage <- function (method = NA, filePath = NA, softwarePath = NA, verbose = FALSE)
 {
-	if (is.na(filePath) == TRUE) {
-		stopf("Please provide a file path for SVM package to add.")
+	# we always need a method
+	if (is.na(method) == TRUE) {
+		stopf("Please provide the name of the SVM package to create")
 	}
+
+	# load definition of SVM object
+	if (is.na(filePath) == FALSE) {
+		source (filePath, local = FALSE)
+	}
+
 	
-	source (filePath, local = FALSE)
+	# TODO: test if the object specified by the name exists
 	
-	# now we make a crucial assumption here,
-	# to make everything 'magical',
-	# assume that the filename already denotes the
-	# method name and (all the callback names if needed)
-	# as well as the caller name (evalXXX).
-	baseName = basename (filePath)
-	methodName = strsplit(baseName, "_")[[1]][1]
-	SVMBridgeEnv$packages = c(SVMBridgeEnv$packages, methodName)
-	SVMBridgeEnv[[methodName]] = list()
 	
-	# FIXME 
-	SVMBridgeEnv[[methodName]]$trainBinaryPattern = "svm-train"
-	SVMBridgeEnv[[methodName]]$trainBinaryOutputPattern = "Usage: svm-train .options. training_set_file .model_file."
-	SVMBridgeEnv[[methodName]]$testBinaryPattern = "svm-predict"
-	SVMBridgeEnv[[methodName]]$testBinaryOutputPattern = "for one-class SVM only 0 is supported"
+	# create an object now
+	if (verbose == TRUE) {
+		messagef ("Creating SVM Wrapper for method %s: ", method)
+	}
+	SVMObject = createSVMWrapper( method = method)
+	SVMBridgeEnv$packages[[method]] = SVMObject
+	if (verbose == TRUE) {
+		messagef ("Registred Objects")
+		for (i in seq(1, length(SVMBridgeEnv$packages))) {
+			s = SVMBridgeEnv$packages[i]
+			print (s)
+		}
+	}
 
 	
 	# add software path
[…]

### 2015-01-28 R/testSVM.R

@@ -57,7 +57,7 @@ testSVM = function(
 
 	# model 
 	model = NULL,
-	modelFilePath = NULL,
+	modelFile = NULL,
 
 	# rest
 	extraParameter = "",
@@ -68,9 +68,11 @@ testSVM = function(
 		messagef("--- Testing...")
 	}
 
-		# FIXME get the correct object
+	# get the correct object
+	SVMObject = SVMBridgeEnv$packages[[method]]
+	
 	# ask object for its path
-	testBinaryPath = SVMBridgeEnv[[method]]$testBinaryPath
+	testBinaryPath = SVMObject$testBinaryPath
 	testBinary = basename(testBinaryPath)
 	
 	
@@ -98,10 +100,10 @@ testSVM = function(
 
 
 	# test model
-	if ( (is.null(model) == FALSE) && (is.null(modelFilePath) == FALSE))
+	if ( (is.null(model) == FALSE) && (is.null(modelFile) == FALSE))
 		stopf("Given a model in memory and specified a model file name. Confused. Stopping.")
 			
-	if ( (is.null(model) == TRUE) && (is.null(modelFilePath) == TRUE))
+	if ( (is.null(model) == TRUE) && (is.null(modelFile) == TRUE))
 		stopf("Neither given a model nor given a path to the model. Stopping.")
 
 
@@ -123,19 +125,20 @@ testSVM = function(
 	
 	# if the user specified a model in memory, we first need to write that
 	if (is.null(model) == FALSE) {
-		modelFilePath = tempfile()
+		modelFile = tempfile()
 		if (verbose == TRUE)
-			messagef("  Writing model in memory to disk as %s", modelFilePath)
-		LIBSVMWriteModelCallBack (model = model, modelFilePath = modelFilePath, verbose = verbose)
+			messagef("  Writing model in memory to disk as %s", modelFile)
+		
+		writeModel (SVMObject, modelFile = modelFile, verbose = verbose)
 	}
 
 	# TODO: add prediction as an option, so people can have them permanently on disk??
 	predictionsFilePath = tempfile()
 	
 	# retrieve test parameters
-	args = LIBSVMTestParameterCallBack( 
+	args = createTestArguments (SVMObject, 
 		testDataFile = testDataFile, 
-		modelFile = modelFilePath, 
+		modelFile = modelFile, 
 		predictionsFilePath = predictionsFilePath,
 		...)
     
@@ -145,10 +148,10 @@ testSVM = function(
 		messagef("Testing took %f seconds.", testTime);
[…]

### 2015-01-28 demo.R

@@ -19,26 +19,31 @@
 # All negative use is prohibited.
 #
 
-library(devtools)
-load_all (".")
 
 #library (SVMBridge)
+	library(devtools)
+	load_all (".")
 
-# add new wrappers -- this is done in initPackage for wrappers known to the package
-# (so this is currently done twice)
-addSVMPackage (filePath = "./R/LIBSVM_wrapper.R", softwarePath = "../../../svm_large_scale/software/libSVM", verbose = TRUE)
-addSVMPackage (filePath = "./R/CVM_wrapper.R", softwarePath = "../../../svm_large_scale/software/libCVM", verbose = FALSE)
+	
+# as the libary already loads default wrappers this works
+	addSVMPackage (method = "LIBSVM", verbose = FALSE)
+	findSVMSoftware ("LIBSVM", searchPath = "../../../svm_large_scale/software/libSVM", verbose = FALSE)
+
+# but this works also
+	addSVMPackage (method = "LIBSVM", filePath = "./R/LIBSVM_wrapper.R", softwarePath = "../../../svm_large_scale/software/libSVM", verbose = FALSE)
+	
 
 
 # alternatively, addSVMPackage without softwarePath and search all in a given path
-#findAllSVMSoftware (searchPath = "../../../../svm_large_scale/software/", verbose = FALSE) 
+	findAllSVMSoftware (searchPath = "../../../svm_large_scale/software/", verbose = FALSE) 
 
 
 # wird alle bekannte software-pakete suchen also SVMperf, libSVM, ...
 #FIXME: allow path like ~/
-outputAllSVMSoftwarePackages ()
+	outputAllSVMSoftwarePackages ()
 
-
+	
+	
 # load iris  for now
 	shufflediris = iris[sample(nrow(iris)),]
 

### 2015-01-28 src/avl.c

@@ -1,589 +1,0 @@
-/*****************************************************************************
-
-    avl.c - Source code for the AVL-tree library.
-
-    Copyright (C) 1998  Michael H. Buselli <cosine@cosine.org>
-    Copyright (C) 2000-2002  Wessel Dankers <wsl@nl.linux.org>
-
-    This library is free software; you can redistribute it and/or
-    modify it under the terms of the GNU Lesser General Public
-    License as published by the Free Software Foundation; either
-    version 2.1 of the License, or (at your option) any later version.
-
-    This library is distributed in the hope that it will be useful,
-    but WITHOUT ANY WARRANTY; without even the implied warranty of
-    MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the GNU
-    Lesser General Public License for more details.
-
-    You should have received a copy of the GNU Lesser General Public
-    License along with this library; if not, write to the Free Software
-    Foundation, Inc., 59 Temple Place, Suite 330, Boston, MA 02111-1307  USA
-
-    Augmented AVL-tree. Original by Michael H. Buselli <cosine@cosine.org>.
-
-    Modified by Wessel Dankers <wsl@nl.linux.org> to add a bunch of bloat to
-    the sourcecode, change the interface and squash a few bugs.
-    Mail him if you find new bugs.
-
-*****************************************************************************/
-
-#include <stdio.h>
-#include <stdlib.h>
-#include <errno.h>
-#include "avl.h"
-
-static void avl_rebalance(avl_tree_t *, avl_node_t *);
-
-#ifdef AVL_COUNT
-#define NODE_COUNT(n)  ((n) ? (n)->count : 0)
-#define L_COUNT(n)     (NODE_COUNT((n)->left))
-#define R_COUNT(n)     (NODE_COUNT((n)->right))
-#define CALC_COUNT(n)  (L_COUNT(n) + R_COUNT(n) + 1)
-#endif
-
-#ifdef AVL_DEPTH
-#define NODE_DEPTH(n)  ((n) ? (n)->depth : 0)
-#define L_DEPTH(n)     (NODE_DEPTH((n)->left))
[…]

### 2015-01-28 src/hv.c

@@ -1,529 +1,0 @@
-/*************************************************************************
-
- hypervolume computation
-
- ---------------------------------------------------------------------
-
-                       Copyright (c) 2010
-                  Carlos Fonseca <cmfonsec@ualg.pt>
-             Manuel Lopez-Ibanez <manuel.lopez-ibanez@ulb.ac.be>
-                    Luis Paquete <paquete@dei.uc.pt>
-
- This program is free software (software libre); you can redistribute
- it and/or modify it under the terms of the GNU General Public License
- as published by the Free Software Foundation; either version 2 of the
- License, or (at your option) any later version.
-
- This program is distributed in the hope that it will be useful, but
- WITHOUT ANY WARRANTY; without even the implied warranty of
- MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the GNU
- General Public License for more details.
-
- You should have received a copy of the GNU General Public License
- along with this program; if not, you can obtain a copy of the GNU
- General Public License at:
-                 http://www.gnu.org/copyleft/gpl.html
- or by writing to:
-           Free Software Foundation, Inc., 59 Temple Place,
-                 Suite 330, Boston, MA 02111-1307 USA
-
- ----------------------------------------------------------------------
-
- Relevant literature:
-
- [1]  C. M. Fonseca, L. Paquete, and M. Lopez-Ibanez. An
-      improved dimension-sweep algorithm for the hypervolume
-      indicator. In IEEE Congress on Evolutionary Computation,
-      pages 1157-1163, Vancouver, Canada, July 2006.
-
- [2]  L. Paquete, C. M. Fonseca and M. Lopez-Ibanez. An optimal
-      algorithm for a special case of Klee's measure problem in three
-      dimensions. Technical Report CSI-RT-I-01/2006, CSI, Universidade
-      do Algarve, 2006.
-
-*************************************************************************/
-
-#include "hv.h"
-#include "avl.h"
-#include <R.h>
-#include <stdlib.h>
-#include <stdio.h>
-#include <limits.h>
-#include <float.h>
-#include <assert.h>
-
[…]

### 2015-01-28 src/hv.h

@@ -1,48 +1,0 @@
-/*************************************************************************
-
- hv.h 
-
- ---------------------------------------------------------------------
-
-                       Copyright (c) 2005, 2006
-                  Carlos Fonseca <cmfonsec@ualg.pt>
-             Manuel Lopez-Ibanez <manuel.lopez-ibanez@ulb.ac.be>
-                    Luis Paquete <paquete@dei.uc.pt>
-
- This program is free software (software libre); you can redistribute
- it and/or modify it under the terms of the GNU General Public License
- as published by the Free Software Foundation; either version 2 of the
- License, or (at your option) any later version.
-
- This program is distributed in the hope that it will be useful, but
- WITHOUT ANY WARRANTY; without even the implied warranty of
- MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the GNU
- General Public License for more details.
-
- You should have received a copy of the GNU General Public License
- along with this program; if not, you can obtain a copy of the GNU
- General Public License at:
-                 http://www.gnu.org/copyleft/gpl.html
- or by writing to:
-           Free Software Foundation, Inc., 59 Temple Place,
-                 Suite 330, Boston, MA 02111-1307 USA
-
- ----------------------------------------------------------------------
-
-
-*************************************************************************/
-#ifndef HV_H_
-#define HV_H_
-
-#ifdef __cplusplus
-extern "C" {
-#endif
-
-extern int stop_dimension;
-double fpli_hv(double *data, int d, int n, const double *ref);
-
-#ifdef __cplusplus
-}
-#endif
-
-#endif

### 2015-01-28 src/macros.h

@@ -1,23 +1,0 @@
-#ifndef FOO_MACROS_H
-#define FOO_MACROS_H
-
-#include <R.h>
-#include <Rinternals.h>
-#include <Rdefines.h>
-
-#define MAX(a,b) (((a)>(b))?(a):(b))
-
-#define UNPACK_REAL_VECTOR(S, D, N) \
-    double *D = REAL(S); \
-    const R_len_t N = length(S);
-
-#define UNPACK_REAL_MATRIX(S, D, N, K) \
-    double *D = REAL(S); \
-    const R_len_t N = nrows(S); \
-    const R_len_t K = ncols(S);
-
-#define UNPACK_REAL_MATRIX_2(S, D, N) \
-    double *D = REAL(S); \
-    const R_len_t N = nrows(S); 
-
-#endif

### 2015-01-28 doc/tutorial/introduction.Rmd
make predictions option

@@
 Finally, an SVM package is for us the internal object in the SVMBridge, that has basically consists of the wrapper and the software (e.g. a path to the binary).
+
+## Speed
+
+It is possible to bypass any extra-I/O and directly communicate with the SVM Package by specifying the training and test data as well as the model and prediction as a file.
+If SVMBridge gets file pathes, it will not try to reread the written model nor the predictions.
+Sometimes you do not need the predictions, neither on memory nor on disk.
+In this case you can pass "/dev/null" as prediction file.
+This will make the SVM Test write to null, so it will not take any I/O time,  and therefore the SVMBridge will not reread the predictions, as it was already written on disk.

### 2015-01-28 R/LIBSVM_wrapper.R

@@ -218,6 +218,7 @@
 	}
 
 	
+	
 	print.LIBSVM = function(x) {
 		messagef("--- Object: %s", x$method)
 		messagef("       Training Binary at %s", x$trainBinaryPath)

### 2015-01-28 demo.R

@@ -42,7 +42,8 @@
 #FIXME: allow path like ~/
 	outputAllSVMSoftwarePackages ()
 
-	
+
+	verbose = FALSE
 	
 # load iris  for now
 	shufflediris = iris[sample(nrow(iris)),]
@@ -57,10 +58,8 @@
 	testDatay[testDatay==3] = 1
 	trainDatay[trainDatay==2] = -1
 	testDatay[testDatay==2] = -1
+
 	
-#  this will train from memory.
-# as no modelfile was given, callSVM will create a temporay model file
-# and read this after training, so it will be in svmObj$model
 	messagef("\n\n\n======= Train LIBSVM, Traindata from Memory, Model to Memory")
 	svmObj =  trainSVM(
 		method = "LIBSVM",
@@ -69,35 +68,78 @@
 		cost = 1, 
 		gamma = 1, 
 		epsilon = 0.01, 
-        verbose = TRUE
+        verbose = verbose
     )  
 
 
-# or, if preferred, load sparse data from disk
 	messagef("\n\n\n======= Train LIBSVM, Traindata from File, Model to Memory")
+    trainObj =  trainSVM(
+		method = "LIBSVM",
+		trainDataFile = './tests/data/australian.train',
+		cost = 1, 
+		gamma = 1, 
+		epsilon = 0.01, 
+        verbose = verbose
+    )  
+
+    
+    messagef("\n\n\n======= Test LIBSVM, Testdata from Disk, Model from Memory, Predictions to Memory")
+    testObj =  testSVM(
+		method = "LIBSVM",
+		testDataFile = './tests/data/australian.train',
+        model = trainObj$model,
+        verbose = verbose
+    )  
+
+    
+	messagef("\n\n\n======= Test LIBSVM, Testdata from Memory, Model from Memory, Predictions to Memory")
+    testObj =  testSVM(
+		method = "LIBSVM",
+		testDataX = testDataX, 
+		testDatay = testDatay, 
+        verbose = verbose,
+        model = trainObj$model
+    )  
+
+    
+    messagef("\n\n\n======= Test LIBSVM, Testdata from Disk, Model from Memory, Predictions to Disk")
+    testObj =  testSVM(
+		method = "LIBSVM",
+		testDataFile = './tests/data/australian.train',
+        model = trainObj$model,
+        predictionsFile = "./tmp/predictions.txt",
+        verbose = verbose
+    )  
+
+    
[…]

### 2015-03-22 doc/tutorial/introduction.Rmd
add find svm wrapper

@@
 Suppose you want to write a wrapper for mySVM.
-In this case, you will need to create a file exactly called "mySVM_wrapper.R'.
+In this case, you will need to create a file exactly called 'mySVM_wrapper.R'.
 Please note that the filename is case-sensitive.
@@
 In general, using other packages and skip **SVMBridge** might be a good idea, as **SVMBridge** strictly follows the LIBSVM way of training and testing.
+
+## Adding the wrapper
+
+After you have written a wrapper, say 'mySVM_wrapper.R', you must add your wrapper to the **SVMBridge**.
+This you can either manually or automatically (see the next section for information about this).
+There are two parts of this procedure: First you must tell **SVMBridge** where to find your wrapper 'mySVM_wrapper.R'.
+After this you must specify where the corresponding binaries for mySVM are to be found.
+Remember that the **SVMBridge** acts mainly as a convenient tool to call a SVM solver via command line.
+
+To add your wrapper to the **SVMBridge** simply call ```addSVMPackage```, e.g.
+if mySVM resides in the mySVM subdirectory, you can call ```addSVMPackage (method = "mySVM", filePath = "./mySVM/mySVM_wrapper.R", softwarePath = "./mySVM/bin")```.
+Notice that filePath is mandatory, and the **SVMBridge** will source the specified wrapper.
+Though you are free to name your wrapper as you wish, it is a good idea to stick to the convention chosen and name the wrapper for your SVM package by adding '_wrapper.R', e.g.
+the wrapper for LASVM is called 'LASVM_wrapper.R' inside **SVMBridge**.
+The path for the binaries of mySVM is not mandotory.
+Instead of specifiying it, **SVMBridge** can automatically find it (based on the code in the mySVM_wrapper.R file).
+
+## Automatically finding the software
+
+After you added your wrapper, if you did not specified a softwarePath, **SVMBridge** can find the corresponding binary files for you.
+Remember that if you forget to specifiy the softwarePath,  **SVMBridge** will not understand where to find the binaries.
+
+The corresponding function is ```findSVMSoftware```.
+To tell the **SVMBridge** where to find the mySVM software, call it like this: ```findSVMSoftware ("mySVM", searchPath = "./mySVM", verbose = FALSE).
+
+For convience, there is also a findAllSVMSoftware function that will try to find the software of all registered SVM packages based on the given search path, e.g., if all your SVM software can be found inside the ./software directory, you can call ```findAllSVMSoftware(searchPath = "./software")```.
+
+## Automatically finding the wrapper
+
+Sometimes it is convenient to allow the user to specify the SVM software he needs.
+In this case it is no good to specify neither the wrapper nor the software by hand.
+Therefore **SVMBridge** allows you not only to automatically find the software, but also the wrapper.
+This only works, if the wrapper follows the naming convention, i.e.
+the wrapper for mySVM needs to be called mySVM_wrapper.R.
+To initiate the search, just call ```findSVMWrapper (method = "mySVM", searchPath = "./mySVM")```.
+The specified searchPath will, just as ```findSVMSoftware```, be searched recursively.
+Notice also that if multiple such wrappers exist inside a directory (in subdirectories), **SVMBridge** will stop searching after the very first hit.
 

### 2015-03-22 R/findSVMWrapper.R

@@ -1,0 +1,106 @@
+#!/usr/bin/Rscript  --vanilla 
+#
+# SVMBridge 
+#		(C) 2015, by Aydin Demircioglu
+#
+#		findSVMWrapper.R
+# 
+# SVMBridge is free software: you can redistribute it and/or modify
+# it under the terms of the GNU Lesser General Public License as published
+# by the Free Software Foundation, either version 3 of the License, or
+# (at your option) any later version.
+#
+# SVMBridge is distributed in the hope that it will be useful,
+# but WITHOUT ANY WARRANTY; without even the implied warranty of
+# MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
+# GNU Lesser General Public License for more details.
+#
+# Please do not use this software to destroy or spy on people, environment or things.
+# All negative use is prohibited.
+#
+ 
+
+
+# findSVMWrapper
+#		given a search path, it will try to find the corresponding wrapper for the given method.
+#
+# @param 	searchPath	 	search the given path for the SVM binaries of all known SVM packages.
+# @param	verbose			print messages while searching?
+#
+# @note		To make sure that the binary is correct, it will be executed! (see findBinary for more infos)
+# @note		If multiple binaries are found, the last one will be taken. Overwrite by hand, if necessary.
+
+	findSVMWrapper <- function (method = NA, name = NA, searchPath = NA, verbose = FALSE) {
+		if (verbose == TRUE) {
+			BBmisc::messagef("API: Finding wrapper for %s", method)
+		}
+		
+		if (is.na(searchPath)) {
+			BBmisc::stopf("No search path is given!")
+		}
+		
+		if (is.na(method)) {
+			BBmisc::stopf ("No method name is given")
+		}
+		
+		if (verbose == TRUE) {
+			BBmisc::messagef("  Trying to find wrapper for %s", method) 
+		}
+
+		if (is.na (name) == TRUE) {
+			pattern = paste( "^", method, "_wrapper.R$", sep = "")
+		} else {
+			pattern = paste( "^", name, "$", sep = "")
+		}
+
+		if (verbose == TRUE) {
+			BBmisc::messagef("  Looking for a wrapper with regex %s.", pattern)
+		}
+		
+		files <- list.files (searchPath, pattern = pattern, recursive = TRUE)
+		foundWrapper = ''
+		for (file in files) {
+			wrapperPath = file.path(searchPath, file)
+			if (verbose == TRUE) { 
[…]

### 2015-03-22 doc/tutorial/introduction.Rmd
extract findbinary, add fix for svmperf

@@
 This will make the SVM Test write to null, so it will not take any I/O time,  and therefore the SVMBridge will not reread the predictions, as it was already written on disk.
+
+# FAQ
+
+## Searching for an unknown solver
+
+If you try to find a wrapper that does not exist, FIXME: fix that.
+for now the error later on:
+Class 'try-error'  atomic [1:1] Error in UseMethod("findSoftware") : 
+  no applicable method for 'findSoftware' applied to an object of class "c('CSVM_walltime', 'SVMWrapper')"
+
+## FindBinary stalls
+
+If the binary expects inputs, executing the binary will stall.
+Please make sure that your binary does not expect input upon being called without parameters.
+If necessary, turn off automatic search and provide **SVMBridge** with the direct path to the software.
+FIXME: WILL SEND 1 to the program, if the program is called SVMperf. will still stall on other wait-for-key-programs.

### 2015-03-22 R/findSVMSoftware.R

@@ -20,42 +20,6 @@
 #
  
 
-# findBinary
-#		given a file pattern and an std-out-pattern it will try to find a binary matching
-#		the file pattern (which contains std-out-pattern as output if called without arguments)
-#		in the path given.
-#
-# @param 	searchPath	 	search the given path for the SVM binaries recursively.
-# @param	pattern		pattern for the binary file
-# @param	outputPattern		pattern for the stdout output of the binary
-# @param	verbose			print messages while searching?
-#
-# @note		To make sure that the binary is correct, it will be executed!
-# 					Furthermore, many SVM packages derive from libSVM. as such, they
-#					often do not change the prediction binary. We will try to sort these out,
-#					but it might be hopeless. With luck, the found binary will be left untouched,
-#					and thus work, if not, you must set the path by hand.
-# @note		If multiple binaries are found, the last one will be taken. Overwrite by hand, if necessary.
-
-findBinary <- function (searchPath, pattern, outputPattern, verbose = FALSE) {
-	if (verbose == TRUE) { BBmisc::messagef("  Checking for pattern %s", pattern) }
-	files <- list.files (searchPath, pattern = pattern, recursive = TRUE)
-    foundBinary = ''
-    for (binary in files) {
-		binaryPath = file.path(searchPath, binary)
-		if (verbose == TRUE) { BBmisc::messagef("    -Found binary at %s", binaryPath) }
-		stdout = system3(binaryPath, args = c(), verbose = FALSE)
-		if (length(grep(outputPattern, stdout$output)) != 0) {
-			foundBinary = binaryPath
-		}
-    } 
-    
-    # TODO:: should we here throw an error? or leave it to the upper class
-    
-    return (foundBinary)
-}
-
-
 
 # findAllSVMSoftware 
 #		given a search path, it will try to find the corresponding software packages

### 2015-03-24 doc/tutorial/introduction.Rmd
make sure readmodel only works if someone forces it

@@
 
-# Model file
+# Models
 
-If you do not specify any model file for training, ```trainSVM``` will ask R for a temporary file and will provide this path to the SVM package.
-Internally, the ```writeModel``` routine of the SVM package will be used to create a model file from a model in memory.
-In case you have not specified a model file path, after training the model will be read from the temporary model file into the SVM Results Object.
-Contrary, if you did specify a model file path, then by default, the model is not read back into the SVM Results Object.
-To enforce reading the model, you can set the argument ```readModel``` to true.
+Usually the underlying SVM solvers will create a model on disk after training. You have two options in ```trainSVM``` to work with these models: modelFile and readModel.
+
+First assume that you provide ```trainSVM``` with a path in the modelFile argument.
+In that case this path will be passed to the corresponding SVM solver, which (if everything runs without an error) will create a model file at the specified path.
+The second option for ```trainSVM``` is readModel.
+If this is set to TRUE, the ```trainSVM``` routine will reread the model back into memory and put it into the resulting trainObject.
+This will allow you work with the model directly in R, e.g.
+by reading out the alpha vector of suppor vector coefficients (if the model provides these information).
+Contrary, if you do not want to work with the model, e.g.
+as you are only interested in the test results of a subsequent ```testSVM``` call, there is no need to parse the model file, and you might want to set readModel to FALSE.
+Note that reading a model back into memory is currently very slow and a resource-hog, so only read back the model if this is really necessary.
+
+In the second case, you do not provide a path to ```trainSVM```.
+In that case ```trainSVM``` has to create a temporary path, as the SVM solver necessarily will need a place to write its model.
+In this case, the readModel flag work exactly as before.
+Note that if you set readModel to FALSE, the model will be lost, as it was only written temporarily, but not read back.
+This might only make sense in very rare, special applications.
+
+Note that in very rare cases you might want to train a SVM, but you are not interested in the model, e.g.
+you only need the time to train or want to check if the SVM will run without an error.
+In that case it is a good idea to pass "/dev/null" as modelFile and set readModel to FALSE.
+This will make the SVM solver to write its model into "/dev/null", effectively discarding it.
+Similarly, ```trainSVM``` will subsequently ignore the not existing model file.
 

### 2015-03-24 R/trainSVM.R

@@ -64,6 +64,7 @@ trainSVM = function(
 	modelFile = NULL,
 	verbose = FALSE,
 
+	readModelFile = FALSE,
 	...) {
 
 	# get the correct object
@@ -85,8 +86,6 @@ trainSVM = function(
 		BBmisc::stopf("No method name is given, this should never happen.")
 	}
 				
-				# FIXME: make this an option
- 	readModelFile = TRUE
  	
 	# did the user specify a modelfile?
 	if (is.null(modelFile) == TRUE) {
@@ -100,8 +99,11 @@ trainSVM = function(
 	# finally, everything is dumped to disk.
   
 	if ( (is.null(trainDataX) == FALSE) && (is.null(trainDataFile) == FALSE))
+	{
+		print(trainDataX)
+		print(trainDataFile)
 		BBmisc::stopf("Given a data frame as training data and specified a training file name. Confused. Stopping.")
-			
+	}		
 	if ( (is.null(trainDataX) == TRUE) && (is.null(trainDataFile) == TRUE))
 		BBmisc::stopf("Neither specified training data path nor given training data. Stopping.")
 

### 2015-05-05 R/RcppExports.R
snapshot, package building should work on linux now

@@ -11,7 +11,7 @@
 #' @return		the data is read into an R matrix and an R vector, containing the data
 #'					and the labels. note, that these are not in sparse format, but are dense.
 #' @examples	
-#'					readSpareData ("./australian.data")
+#'					#readSpareData ("./australian.data")
 #' @export
 readSparseData <- function(filename, verbose = FALSE, zeroBased = FALSE) {
     .Call('SVMBridge_readSparseData', PACKAGE = 'SVMBridge', filename, verbose, zeroBased)
@@ -26,9 +26,9 @@ readSparseData <- function(filename, verbose = FALSE, zeroBased = FALSE) {
 #' @keywords	IO 
 #' @return		NULL.
 #' @examples	
-#'	X = as.matrix(iris[,1:4])
-#'	Y = as.matrix(as.numeric(iris[,5]))
-#'	writeSparseData (X, Y, "./australian.data")
+#'	#X = as.matrix(iris[,1:4])
+#'	#Y = as.matrix(as.numeric(iris[,5]))
+#'	#writeSparseData (X, Y, "./australian.data")
 #' @export
 writeSparseData <- function(filename, X, Y, verbose = FALSE, zeroBased = FALSE) {
     .Call('SVMBridge_writeSparseData', PACKAGE = 'SVMBridge', filename, X, Y, verbose, zeroBased)

### 2015-05-05 examples/examples/ex_2d_numeric.R

@@ -1,25 +1,0 @@
-##### optimizing branin in 2D with mbo / EI #####
-
-library(mlrMBO)
-library(ggplot2)
-library(soobench)
-set.seed(1)
-configureMlr(show.learner.output = FALSE)
-
-set.seed(423)
-
-obj.fun = branin_function()
-
-par.set = extractParamSetFromSooFunction(obj.fun)
-
-ctrl = makeMBOControl(init.design.points = 10, iters = 10, propose.points = 1)
-ctrl = setMBOControlInfill(ctrl, crit = "ei", opt = "focussearch", opt.focussearch.points = 2000)
-
-lrn = makeLearner("regr.km", predict.type = "se", covtype = "matern3_2")
-
-run = exampleRun(obj.fun, par.set = par.set, learner = lrn, control = ctrl, 
-	points.per.dim = 50, show.info = TRUE)
-
-print(run)
-
-res = autoplot(run, pause = TRUE)

### 2015-05-05 examples/examples/ex_2d_numeric_multipoint.R

@@ -1,28 +1,0 @@
-##### optimizing branin in 2D with multipoint proposal #####
-
-library(mlrMBO)
-library(ggplot2)
-library(soobench)
-set.seed(2) # FIXME: does not work for seed == 1
-configureMlr(show.learner.output = FALSE)
-
-obj.fun = branin_function()
-
-par.set = extractParamSetFromSooFunction(obj.fun)
-
-ctrl = makeMBOControl(init.design.points = 10, iters = 10, propose.points = 5)
-ctrl = setMBOControlMultiPoint(ctrl,
-  method = "multicrit",
-  multicrit.objective = "ei.dist",
-  multicrit.dist = "nearest.neighbor",
-  multicrit.maxit = 200
-)
-
-lrn = makeLearner("regr.km", predict.type = "se", covtype = "matern3_2")
-
-run = exampleRun(obj.fun, par.set = par.set, learner = lrn, control = ctrl,
-	points.per.dim = 50, show.info = TRUE)
-
-print(run)
-
-res = autoplot(run, pause = TRUE)

### 2015-05-05 examples/examples/ex_multicrit.R

@@ -1,28 +1,0 @@
-
-library(mlrMBO)
-library(ggplot2)
-library(mco)
-
-set.seed(1)
-configureMlr(show.learner.output = FALSE)
-
-# f = makeMBOFunction(function(x) {
-  # c(x^2, (x - 2)^2)
-# })
-f = makeMBOFunction(zdt1)
-ps = makeNumericParamSet(len = 2L, lower = 0, upper = 1)
-
-learner = makeLearner("regr.km", nugget.estim = FALSE,
-  predict.type = "se")
-
-iters = 10L
-ctrl = makeMBOControl(iters = iters, number.of.targets = 2L,
-  init.design.points = 5L, save.on.disk.at = integer(0L))
-ctrl = setMBOControlInfill(ctrl, crit = "dib",
-  opt.focussearch.points = 10000)
-ctrl = setMBOControlMultiCrit(ctrl, parego.s = 100)
-
-run = exampleRunMultiCrit(f, ps, learner, ctrl, points.per.dim = 50,
-  show.info = TRUE, nsga2.args = list(), ref.point = c(11, 11))
-
-autoplot(run, pause = TRUE, iters = 1:iters)

### 2015-05-05 src/SparseData.cpp

@@ -83,7 +83,7 @@ static char* readline(FILE *input)
 //' @return		the data is read into an R matrix and an R vector, containing the data
 //'					and the labels. note, that these are not in sparse format, but are dense.
 //' @examples	
-//'					readSpareData ("./australian.data")
+//'					#readSpareData ("./australian.data")
 //' @export
 // [[Rcpp::export]] 
 List readSparseData (std::string filename, bool verbose = false, bool zeroBased = false) {
@@ -256,9 +256,9 @@ List readSparseData (std::string filename, bool verbose = false, bool zeroBased 
 //' @keywords	IO 
 //' @return		NULL.
 //' @examples	
-//'	X = as.matrix(iris[,1:4])
-//'	Y = as.matrix(as.numeric(iris[,5]))
-//'	writeSparseData (X, Y, "./australian.data")
+//'	#X = as.matrix(iris[,1:4])
+//'	#Y = as.matrix(as.numeric(iris[,5]))
+//'	#writeSparseData (X, Y, "./australian.data")
 //' @export
 // [[Rcpp::export]] 
 List writeSparseData (std::string filename, NumericMatrix X, NumericVector Y, bool verbose = false, bool zeroBased = false) {

### 2015-08-18 doc/tutorial/windowsHowTo.Rmd
few fixes

***SVMBridge Windows Tutorial***

1) Download R 3.2.x or later for Windows (32/64 bit) (https://cran.r-project.org/bin/windows/base/)

2) Download Rtools for R 3.2.x and later (https://cran.r-project.org/bin/windows/Rtools/)

3) Download Git for Windows (https://git-scm.com/downloads)

3) Install R and use the 32 bit option only

4) Install Rtools and use the 32 bit option only

5) Install Git

6) Start up the Git Bash and navigate into a directory in which you want to download SVMbridge to. For best compatibility,
   install SVMBridge into your home directory. Some tests may fail if you put it elsewhere.

7) Download the SVMBridge repository by using git clone in the git console 
   (git clone https://github.com/aydindemircioglu/SVMBridge.git)

8) Delete the makefile.win file, which is located in the src directory in SVMBridge.

9) Create a folder structure "shark/svm_large_data/software/" in your home directory, download and install the latest version of the SVM you want to use in 
   the newly created shark folder. In this directory you will find executables with names like "svm-train.exe" and "svm-predict.exe" 
   for each SVM.(By example if you use LIBSVM and downloaded the latest version from "https://www.csie.ntu.edu.tw/~cjlin/libsvm/"). 
   At this point you have to look into some files, which are located in the SVMBridge\R folder.

   The file you need is called *_wrapper.R whereas the * stands for the SVM you're using.
   In this file, look for a function called findSoftware.* whereas the * stands for the SVM you're using. A few lines below, two variables
   called trainBinaryPattern and testBinaryPattern are defined with a certain string for the windows platform. Be sure that the binaries 
   of the SVM you are using are named the same way, or else the function won't be able to find them. Be sure as well to remember the path 
   to where you installed the SVM, since it is a needed argument for usage.
   (For the example with LIBSVM you don't need to change the names of the binaries)

9) Start up R Gui and install following packages by typing "install.packages("package-name")" into the R Gui: BBmisc, stringr, 
    plyr, Rcpp, microbenchmark, R.utils, e1071, SparseM, testthat, knitr, devtools

10) Your Environment is now ready to get started:

11) Find the option "change directory" and navigate to the directory where you downloaded SVMBridge to and type "devtools::test()" into your R gui 
    for a full test run.

12) Type in the following small demo into your R console for a quick start in how to use the wrapper. This demo uses the LIBSV machine as stated earlier. (More demos can be found in demo.R in SVMBridge).

trainFile = "*realtive path to the directory where your R executable lies in, to the directory of your dataset (e.g for c:/ProgramData/R-3.2.1/bin/ you need to jump three directories
			back to be in directory c:/) or absolute path to your dataset. For this example you can use the file sparse.data located in SVMBridge/tests/data/. (You may also use
			tilde character.)*"

addSVMPackage(method="LIBSVM", verbose=TRUE)

findSVMSoftware(
	"LIBSVM", 
	searchPath = "*relative path or path to some parent directory of the used SVM binaries*", 	
	verbose = TRUE
)

trainObject = trainSVM( 
		method = "LIBSVM",
		trainDataFile = trainFile,
		cost = runif(1),
		gamma = runif(1),
		readModelFile = TRUE,
		verbose = verbose
)

testObject = testSVM(
		method = "LIBSVM",
		testDataFile = trainFile,
		modelFile = trainObject$model,
		verbose = verbose
)

### 2015-09-15 paper/article.tex
first sketch of a paper

\documentclass[article]{jss}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% declarations for jss.cls %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%% almost as usual
\author{Ayd\i n Demircio\u{g}lu\\Ruhr-Universit\"at Bochum \And 
Hanna Houphouet\\Ruhr-Universit\"at Bochum \And 
        Daniel Horn\\TU Dortmund \AND
        Tobias Glasmachers\\Ruhr-Universit\"at Bochum \And 
        Bernd Bischl\\TU M\"unchen \And 
        Claus Weihs\\TU Dortmund}
\title{\pkg{SVMBridge}: Using SVM Solvers from R}

%% for pretty printing and a nice hypersummary also set:
\Plainauthor{Ayd\i n Demircio\u{g}lu, Daniel Horn} %% comma-separated
\Plaintitle{SVMBridge: Using SVM Solvers from R} %% without formatting
\Shorttitle{\pkg{SVMBridge}: Using SVM Solvers from R} %% a short title (if necessary)

%% an abstract and keywords
\Abstract{
  Most SVM Solver do not come with a wrapper in R.
  In case the SVM Solver is written in C++, it can be linked to
  a package via Rcpp. Although possible, this entails a lot of work to do,
  and is not possible, if the SVM Solver is written in other languages.
  Alternatively, one can call the SVM Solver from within R by a system command.
  SVMBridge eases this calls by providing a framework and ready wrappers
  for several SVM Solvers like LASVM, SVMperf, LLSVM and BVM/CVM.  
}
\Keywords{support vector machines, command line}
\Plainkeywords{support vector machines, command line} %% without formatting
%% at least one keyword must be supplied

%% publication information
%% NOTE: Typically, this can be left commented and will be filled out by the technical editor
%% \Volume{50}
%% \Issue{9}
%% \Month{June}
%% \Year{2012}
%% \Submitdate{2012-06-04}
%% \Acceptdate{2012-06-04}

%% The address of (at least) one author should be given
%% in the following format:
\Address{
  Ayd\i n Demircio\u{g}lu\\
  Institut f\"ur Neuroinformatik\\
  Ruhr-Universit\"at Bochum\\
  44790 Bochum
  E-mail: \email{aydin.demircioglu@ini.rub.de}\\
  URL: \url{http://www.ini.rub.de/PEOPLE/aydind/}
}
%% It is also possible to add a telephone and fax number
%% before the e-mail in the following format:
%% Telephone: +43/512/507-7103
%% Fax: +43/512/507-2851

%% for those who use Sweave please include the following line (with % symbols):
%% need no \usepackage{Sweave.sty}

%% end of declarations %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{document}

%% include your article here, just as usual
%% Note that you should use the \pkg{}, \proglang{} and \code{} commands.

%% Note: If there is markup in \(sub)section, then it has to be escape as above.

\section{Introduction}

\subsection{Support Vector Machines}

Support Vector machines in general.

\subsection{SVM packages in R}

There are only a few SVM libraries readily available from within R.
This includes the kernlab, the e1071 and the lasvmR packages.
All of these link existing C++ sources directly from R and need
more or less extensive wrappers to allow the user direct access to
the options of the underlying SVM solver.

Other SVM solver, like the BudgetedSVM package, which contains
BSGD and LLSVM, or SVMperf cannot be called directly from R.
Instead, a system command has to be issued.
To ease this calls, the \pkg{SVMBridge} package has been developed.

\section{The SVMBridge Package}

\subsection{Wrapper}

A wrapper consists of several routines: It must provide routines
for assembling the command line call for training as well as testing,
reading and writing the model and a few general routines like searching
the binaries and printing.

Note that the search for binaries might have a huge toll on HPC clusters,
so it is adviced to specify the paths directly instead of relying on the 
automatism.

\section{Examples}

To use the SVMBridge, one usually perform the following steps:
Add the external wrapper to the SVMBridge, load the data, train a model, use the
model for prediction.

Notice that the data must be written by the SVMBridge before calling 
the corresponding SVM Solver, as there is no way to pass data via memory
to command line tools. Therefore

\subsection{Adding the Wrapper to the SVMBridge}

\subsection{Training a model}

\subsection{Predicions}

\subsection{Optimization Values}

At times, some of the training values are of interest,
e.g. when comparing different solvers, the primal value
of the SVM problem as well as the dual value might  be of interest.
These can be computed via the \code{optimizationValues()} routine.

\subsection{Performance Considerations}

We sum up the points to keep in mind when performance is of high priority:
Do not use the automatism to find the binaries, specify the paths by hand.
Do not load the data into memory, specify the path of the data when training instead.
Do not re-read the model into memory, let it on disk.

\section{Conclusion}

We provided a simple framework to attach many SVM solvers to the R Subsystem.
This allows for a systematic call of them.
By providing a wrapper, linking any SVM solver to R become much easier.

\section*{Acknowledgments}

\end{document}

### 2015-09-15 paper/article.tex
small update

@@
 
+\subsection{Reading Sparse Data}
+
+Most SVM solver work with the sparse data format.
+This format consists of:
+Each line (delimited by a CR/LF) is given by a label
+and the non-zero components of the data point,
+e.g. to encode the vector six dimensional vector 
+$0 0 0 0.2 0 1.4$ belonging to class $3$, the sparse 
+data format would contain the line {3 4:0.2 6:1.4}.
+
+Although reading these files into R is possible via either
+the kernlab or the e1071 routines, both only provide R solutions,
+and therefore suffer from suboptimal performance.
+The SVMBridge package provides a simple sparse data format 
+reading and writing, which are implemented in C++ and
+therefore are nearly two orders of magnitudes faster than
+the corresponding e1071 routines.
+Notice, that currently only dense matrices are supported,
+TODO: work around this with the Sparse Matrix Package.
+
 \subsection{Training a model}
 
+Multiclass is supported, where possible.
+Note that several SVM packages only support binary problems.
+In these cases, training a one-vs-all machine is rather easy, if the
+data is loaded into R first.
+
+Reading a model is possible via the \code{readModelFromFile ()} function.
+This will try to detect the format of the file by calling the 
+\code{isModelFile} routine of each known wrapper.
+Unluckily, several SVM solver use similar model formats, so that
+a direct detection is not possible, e.g. CVM/BVM follow the LIBSVM
+format, but add a comment (with '\#')about the running time to the
+bottom of the model file. Apart from this change, there is no difference.
+From a practical viewpoint we extended the LIBSVM model reader
+to cope with this extra line, so that there is no need for an extra CVM/BVM reader.
+This means that reading a CVM model will make the SVMBridge to detect a LIBSVM
+model (if the LIBSVM wrapper is loaded).
+In these cases, if multiple models claims ownership, a "default" model can be provided,
+which will take precedence over other models. Without a default,
+the model will be random.
+
 \subsection{Predicions}
+
+Predicting from a trained model is rather easy, as there are usually not many options.
+As with training, the predict routine will accept a model either in memory or from a file.
 
@@
 These can be computed via the \code{optimizationValues()} routine.
+
+\subsection{Helper Routines}
+
+...
 
@@
 
+We acknowledge support by the Mercator Research Center Ruhr,
+under grant Pr-2013-0015 \textit{Support-Vektor-Maschinen f{\"u}r extrem gro{\ss}e Datenmengen} and
+partial support by the German Research Foundation (DFG) within the Collaborative Research Centers SFB 823
+\textit{Statistical modelling of nonlinear dynamic processes}, Project C2.
+
 \end{document}

### 2015-09-15 paper/article.tex
update paper

@@
-\documentclass[article]{jss}
+\documentclass[article, shortnames]{jss}
+
+% TODO:
+% C++/R -> prolang{.}
+% SVMBridge -> pkg{.}
+
+\usepackage{amsmath, amssymb}
+\usepackage{mathtools}
+
+\newcommand{\sign}{\mathop{\mathrm{sign}}}
+\newcommand{\N}{\mathbb{N}}
+\newcommand{\R}{\mathbb{R}}
+\newcommand{\Order}{\mathcal{O}}
 
@@
 \Address{
-  Ayd\i n Demircio\u{g}lu\\
+  Ayd\i n Demircio\u{g}lu, Hanna Houphouet, Tobias Glasmachers\\
   Institut f\"ur Neuroinformatik\\
   Ruhr-Universit\"at Bochum\\
-  44790 Bochum
-  E-mail: \email{aydin.demircioglu@ini.rub.de}\\
-  URL: \url{http://www.ini.rub.de/PEOPLE/aydind/}
+  44790 Bochum\\
+  Germany\\
+  E-mail: \email{\{aydin.demircioglu, hanna.houphouet, tobias.glasmachers\}@ini.rub.de}\\
+  URL: \url{http://www.ini.rub.de}\\
+  \\
+
+  Daniel Horn, Claus Weihs\\
+  Fakult\"at Statistik\\
+  Technische Universit\"at Dortmund\\
+  44221 Dortmund\\
+  Germany \\ 
+  E-mail: \email{\{dhorn, bischl, weihs\}@statistik.tu-dortmund.de} \\
+  URL: \url{https://www.statistik.tu-dortmund.de/computationalstats.html}\\
+\\
+
+  Bernd Bischl\\
+Institut f\"ur Statistik \\
+Ludwig-Maximilians-Universit\"at M\"unchen\\
+80539 M\"unchen\\
+  Germany \\
+E-mail: \email{bernd.bischl@stat.uni-muenchen.de}\\
+URL: \url{http://www.statistik.lmu.de/~bischl}
 }
+
 %% It is also possible to add a telephone and fax number
@@
 
-Support Vector machines in general.
+Support Vector Machines (SVM) are linear classifiers
+that try to maxime the margin of a binary problem \citep{cortes1995support}.
+As SVMs are only linear classifiers, a kernelized version
+is used for more complex data. Basically, they work by
+solving the  following convex problem:
+\begin{equation}
+        \min_{w \in \mathcal{H}, b \in \R} \quad \frac{1}{2} ||w||^2 + C \cdot \sum_{i=1}^n \max \Big( 0, 1 - y_i \big( \langle w, \varphi(x_i) \rangle_{\mathcal{H}} + b \big) \Big).
+        \label{SVMproblemPrimal}
+\end{equation}
+Here 
+$\mathcal{D}=\{(x_1, y_1), \dots, (x_n, y_n)\} \in (X \times \{ \pm 1\})^n$
+is the labeled data,
+and  $\varphi : X \to \mathcal{H}$ is a feature map into a reproducing
+kernel Hilbert space $\mathcal{H}$, corresponding to a positive definite
+(Mercer) kernel function $k : X \times X \to \R$, fulfilling
+$\langle \varphi(x), \varphi(x') \rangle_{\mathcal{H}} = k(x, x')$ for
+all $x, x' \in X$. Furthermore, $C > 0$ is a regularization parameter.
+It controls the complexity of the SVM model.
+
+Often, the problem is not directly solved, but dualized via Lagrangian theory.
+Then the problem can be stated as 
+\begin{align}
+        \max_{\alpha \in \R^n} \quad & \sum_{i=1}^n \alpha_i - \frac{1}{2} \sum_{i,j=1}^n \alpha_i \alpha_j y_i y_j k(x_i, x_j) \label{SVMproblemDual} \\
+        \text{s.t.} \quad & \sum_{i=1}^{n} y_i \alpha_i = 0 \text{~~and~~} 0 \leq \alpha_i \leq C \,\,\, \forall \, i \in \{1, \dots, n\}. \notag
+        \enspace
+\end{align}
+The solution then takes the form $w = \sum_{i=1}^n \alpha_i y_i \varphi(x_i)$
+and the offset~$b$ can be computed from the Karush-Kuhn-Tucker (KKT)
+complementarity conditions.
+
+SVMs classify by using the model given by
+$h_{w,b}(x) = \textrm{sign}(\langle w, \phi(x)\rangle_{\mathcal{H}} + b)$
+
+Different kernels can be used, e.g. the polynomial kernel or the 
+RBF kernel $k(x,x') = e^{-\gamma || x-x' ||^2}$, where
+$\gamma$ and $d$ are parameters of the kernel, that must
+be tuned to the given data.
+
+SMO.
+
+On the practical side, LIBSVM can be regarded as the 
+reference implementation of the SMO and the 'golden standard'
+of SVM solvers \citep{cc01a}.
 
@@
 
-Other SVM solver, like the BudgetedSVM package, which contains
-BSGD and LLSVM, or SVMperf cannot be called directly from R.
+Other SVM solver, like the BudgetedSVM package~\citep{djuric2014budgetedsvm}, 
+which contains BSGD and LLSVM, or SVMperf~\citep{joachims2009sparse}, cannot be called directly from R.
 Instead, a system command has to be issued.
@@
 
-\section{Examples}
+We provided default wrappers for LIBSVM, LASVM, CVM/BVM,
+SVMperf, BSGD and LLSVM.
+
+\section{Using a Wrapper}
 
@@
 
+The first step is to make the SVMBridge aware of a new wrapper.
+This is simply done by calling \code{addSVMPackage()}.
+There are several options.
+The SVMBridge comes with an easy mechanism to search 
+the corresponding binaries (specified in the wrapper) to ease
+the usage. To use this, one can instead call \code{findSVMSoftware}.
+
 \subsection{Reading Sparse Data}
@@
 e.g. to encode the vector six dimensional vector 
-$0 0 0 0.2 0 1.4$ belonging to class $3$, the sparse 
+ $ ( 0 , 0 , 0 , 0.2 , 0 , 1.4 ) $
+belonging to class $3$, the sparse 
 data format would contain the line {3 4:0.2 6:1.4}.
@@
 Although reading these files into R is possible via either
-the kernlab or the e1071 routines, both only provide R solutions,
+the \pkg{kernlab} or the \pkg{e1071} routines, both only provide R solutions,
 and therefore suffer from suboptimal performance.
@@
 
+\begin{CodeChunk}
+\begin{CodeInput}
+R> predictions = testSVM (model, data = '')
+R> head(predictions)
+\end{CodeInput}
+\begin{CodeOutput}
+...
+\end{CodeOutput}
+\end{CodeChunk}
+
 \subsection{Optimization Values}
@@
 
+\subsection{Platform Considerations}
+
+Although the SVMBridge was meant to be cross-platform, this goal
+is hard to achieve. From a users perspective, it can be used on all three
+major platforms (Linux, MacOS, Windows). Additional tests should
+be started to make sure the package works as intended.
+
 \subsection{Performance Considerations}
@@
 Do not re-read the model into memory, let it on disk.
+
+\section{Creating a Wrapper}
+
+Adding your own SVM solver to the SVMBridge boils down to
+writing a S3 class with several routines. In the inst folder you will find
+a template that you can fill with your own code. Here we will go
+through the details of adding the software package ...
+
+\subsection{Training Call}
+
+\subsection{Testing Call}
+
+\subsection{Handling Models}
+
+As only LIBSVM models are being used, one can readily use the 
+inbuild functions of the SVMBridge.
 
@@
 
+\bibliography{collected}
+
 \end{document}

### 2015-09-15 paper/article.tex
update paper

@@
 
+\begin{CodeChunk}
+\begin{CodeInput}
+R> model = trainSVM (method = "LIBSVM", cost = 1.0, gamma = 2.0,
+			trainDataFile = "./data.sparse",
+			epsilon = 0.042,
+			modelFile = modelFile,
+			useBias = TRUE,
+			verbose = verbose
+		)
+\end{CodeInput}
+\begin{CodeOutput}
+...
+\end{CodeOutput}
+\end{CodeChunk}
+
+The training data can be passed on via a variable in memory or by specifying the path 
+of the data set file. 
+Note that there are two sets of variables: Those who belong to the SVMBridge,
+like the trainDataFile variable or verbose flag, and those that are passed further
+to the underlying SVM wrapper.
+Not all options that are provided by the underlying SVM solver
+are supported by the wrappers. As our focus lies on binary classification, because of time constraints 
+we opted  to support only these options, e.g. we dropped the one-class SVDD in LIBSVM.
+In case there is need for other options, it is 
+easy to enhance the wrappers.
+There is furthermore the subsampling option, which can be used for larger data sets.
+
 \subsection{Predicions}
@@
 As with training, the predict routine will accept a model either in memory or from a file.
+The same is true for the data to predict.  Again notice that data and models in memory 
+must be written to disk prior to calling the prediction binary.
 
@@
 
+Specifically, this routine compute the following values:
+\code{weight} is given by $\frac{1}{2} w^T w$, with
+$w = \sum_i \alpha_i s_i$. Furthermore, we have
+that $C \sum_i max (0, 1 - ..) $, and 
+\code{primal} is then given by $primal = w + CHinge$.
+\code{dual} on the other hand is simply
+$w - \sum_i |\alpha_i|$. In general it must hold
+that $primal \geq dual$. 
+WHAT IF NOT? what does it mean?
+
 \subsection{Helper Routines}
 
-...
+A few routines have been placed into the package that were helpful
+with testing. Namely, there are several generators for synthetical data sets,
+collected from different publications. 
+WHAT ELSE? can routines be used outside? detectTilde e.g.?
 
@@
 
+The flow is as follows: The SVMBridge will call 
+createTrainingArguments method of any wrapper. The wrapper
+will return a string that contains all parameters that need to be
+passed to the binary of the underlying SVM solver. Note that
+this includes the model, prediction and training files.
+
 \subsection{Training Call}
 
+A simple S3 method has to be created that needs to be called
+\code{createTrainingArguments.LIBSVM}. Apart from the very
+first argument (taking the object itself), all other paramters
+can be chosen freely. After training, from the output several
+training information will be extracted by the 
+\code{extractTrainingInfo()} function.
+
 \subsection{Testing Call}
+
+Similar to the training call, \code{createTestArguments()} needs to
+writen. Again, it should accept three variables called testDataFile and
+modelFile as well as predictionFile. There is also a
+\code{extractTestInfo()} that will extract the important informations
+from the output of the test binary.
 
@@
 As only LIBSVM models are being used, one can readily use the 
-inbuild functions of the SVMBridge.
+inbuild functions of the SVMBridge. If the solver works with its
+own model, one needs to rewrite the readModel method of the wrapper.
+This will only obtain the modelFile and will return the model
+in the model format (S3 Object with XY, describe it somewhere else)
+SVMBridge needs also a writeModel functions, as models passed in
+memory needs to be written to disk. .
+Finally it is necessary to access the predictions itself. For this one
+needs to write the \code{readPredictions()} function. For LIBSVM this
+will read the prediction files that is being written by LIBSVM to disk.
+
+\subsection{Other functions}
+
+Every wrapper should also be able to find tiself in a given directory tree.
+For this \code{findSoftware} is provided.
+
+At last, for querying reasons, there is also a print routine
+so that wrapper can print out any further information about itself.
 

### 2015-09-15 test_SVMperfOptValue.R

@@ -31,11 +31,13 @@ C = 99.1
 	)  
 
 	model = trainedObj$model
+	writeModelToFile (model, "./libsvm.model")
+
+	model$alpha = model$alpha 
 	data = readSparseData (trainFile)
 	oV = optimizationValues (X = data$X, Y = data$Y, model = model, C = C)
 	print (oV)
 	
-	writeModelToFile (model, "./libsvm.model")
 	model2 = readModelFromFile ("./libsvm.model")
 	oV = optimizationValues (X = data$X, Y = data$Y, model = model2, C = C)
 	print (oV)

### 2015-09-16 paper/article.tex
snapshot

@@
 % TODO:
-% C++/R -> prolang{.}
+% \proglang{C++}/R -> prolang{.}
 % SVMBridge -> pkg{.}
@@
 \Abstract{
-  Most SVM Solver do not come with a wrapper in R.
-  In case the SVM Solver is written in C++, it can be linked to
+  Most SVM Solver do not come with a wrapper in  \proglang{R}.
+  In case the SVM Solver is written in \proglang{C++}, it can be linked to
   a package via  Rcpp. Although possible, this entails a lot of work to do,
   and is not possible, if the SVM Solver is written in other languages.
-  Alternatively, one can call the SVM Solver from within R by a system command.
+  Alternatively, one can call the SVM Solver from within \proglang{R} by a system command.
   SVMBridge eases this calls by providing a framework and ready wrappers
@@
 
-\subsection{SVM packages in R}
+\subsection[SVM packages in R]{SVM packages in \proglang{R}}
 
-There are only a few SVM libraries readily available from within R.
-This includes the kernlab, the e1071 and the lasvmR packages.
-All of these link existing C++ sources directly from R and need
+There are only a few SVM libraries readily available from within \proglang{R}.
+This includes the well-known packages like \pkg{kernlab} (via \code{ksvm()})~\citep{kernlab}, 
+and \pkg{e1071} (via \code{svm()})~\citep{dimitriadou2008misc}
+as well as other less known ones, e.g. \pkg{SwarmSVM}~\citep{SwarmSVM} 
+and \pkg{lasvmR}~\citep{lasvmR}.
+All of these link existing \proglang{C++} sources directly from within \proglang{R} and need
 more or less extensive wrappers to allow the user direct access to
@@
 
-Other SVM solver, like the BudgetedSVM package~\citep{djuric2014budgetedsvm}, 
-which contains BSGD and LLSVM, or SVMperf~\citep{joachims2009sparse}, cannot be called directly from R.
+Other SVM solvers, like the BudgetedSVM package~\citep{djuric2014budgetedsvm}, 
+which contains BSGD and LLSVM, or SVMperf~\citep{joachims2009sparse}, cannot be called directly from  \proglang{R}.
 Instead, a system command has to be issued.
@@
 
+The \pkg{SVMBridge} provides a framework to incorporate 
+SVM solvers by means of direct calls (via \code{system2()} commands).
+This is different to \pkg{e1071} and \pkg{kernlab}, which link (and
+thus copy) existing \proglang{C++} code directly inside the package.
+\pkg{SVMBridge} works thus differently: Each SVM solver must provide
+its own socalled {\it Wrapper}. This is a piece of R code that contains
+code e.g. to allow the proper calling of the system binary of a SVM solver
+as well as handling the corresponding models. Such wrapper can be
+registered in the SVMBridge and henceforth can be used opaquely.
+
 \subsection{Wrapper}
 
-A wrapper consists of several routines: It must provide routines
+A wrapper is a S3 object: It must provide routines
 for assembling the command line call for training as well as testing,
@@
 
-Although reading these files into R is possible via either
-the \pkg{kernlab} or the \pkg{e1071} routines, both only provide R solutions,
+Although reading these files into \proglang{R} is possible via either
+the \pkg{kernlab} or the \pkg{e1071} routines, both only provide \proglang{R} solutions,
 and therefore suffer from suboptimal performance.
 The SVMBridge package provides a simple sparse data format 
-reading and writing, which are implemented in C++ and
+reading and writing, which are implemented in \proglang{C++} and
 therefore are nearly two orders of magnitudes faster than
@@
 In these cases, training a one-vs-all machine is rather easy, if the
-data is loaded into R first.
+data is loaded into \proglang{R} first.
 
@@
 
-Every wrapper should also be able to find tiself in a given directory tree.
-For this \code{findSoftware} is provided.
+Every wrapper should also be able to find itself in a given directory tree.
+For this the S3 method \code{findSoftware()} must be provided.
 
@@
 so that wrapper can print out any further information about itself.
+To see the currently loaded wrapper, one can call 
+\code{overviewAllSVMSolver()}.
 
@@
 
-We provided a simple framework to attach many SVM solvers to the R Subsystem.
-This allows for a systematic call of them.
-By providing a wrapper, linking any SVM solver to R become much easier.
+We provided a simple framework, \pkg{SVMBridge},  to attach many SVM solvers to the \proglang{R} landscape.
+\pkg{SVMBridge} does this by an easily extensible wrappers that manage the exchange of data and models
+via command line.
+By providing such a wrapper, linking any SVM solver to \proglang{R} becomes much easier.
 
@@
 
-We acknowledge support by the Mercator Research Center Ruhr,
+We acknowledge support by the Mercator  \proglang{R}.search Center  \proglang{R}.hr,
 under grant Pr-2013-0015 \textit{Support-Vektor-Maschinen f{\"u}r extrem gro{\ss}e Datenmengen} and
-partial support by the German Research Foundation (DFG) within the Collaborative Research Centers SFB 823
+partial support by the German  \proglang{R}.search Foundation (DFG) within the Collaborative  \proglang{R}.search Centers SFB 823
 \textit{Statistical modelling of nonlinear dynamic processes}, Project C2.

### 2015-09-16 R/addSVMPackage.R

@@ -158,13 +158,24 @@ addSVMPackage <- function (method = NA, filePath = NA, softwarePath = NA,
 
 
 
-#' dump all known package infos
+#' Dump all known package infos
+#'
+#' @param	verbose		be verbose in output?
+#' @return	A vector of all known SVM solvers as string.
+#'
 #' @export	outputAllSVMSoftwarePackages
-outputAllSVMSoftwarePackages <- function () {
-	BBmisc::messagef("Currently known solver:")
+
+outputAllSVMSoftwarePackages <- function (verbose = TRUE) {
+	knownPackages = c()
+	if (verbose == TRUE)
+		cat("Currently known solver:\n")
 	for (package in SVMBridgeEnv$packages) {
-		print (package)
+		if (verbose == TRUE)
+			print (package)
+		knownPackages = c(knownPackages, package)
 	}
+	
+	return (knownPackages)
 }
 
 

### 2015-09-16 src/OptimizationValues.cpp

@@ -425,6 +425,7 @@ Rcpp::List computeOptimizationValues (NumericMatrix X, NumericVector Y, double C
 			// predict value
 			double currentMarginO = svm_predict_values (gamma, X(i,_), SV, nSV, alpha, rho, label);
 			double currentMargin = 1.0 - double(Y[i]) * currentMarginO; 
+//			Rcout << currentMargin << "\n";
 			if (currentMargin > 0)
 				hingeLoss += currentMargin;
 			if ((currentMarginO <= 0.0) && (Y[i] > 0))
@@ -467,7 +468,7 @@ Rcpp::List computeOptimizationValues (NumericMatrix X, NumericVector Y, double C
 		
 		// output 
 		if (verbose == true) 
-			Rcout << "Current weight (0.5*||w||): " <<  sqrt(2*weight) << "\n";
+			Rcout << "Current weight (||w||): " <<  sqrt(weight) << "\n";
 		if (verbose == true) 
 			Rcout << "Training error (" << trainingError << "/" << l << "): " << trainingError/X.nrow() << "\n";
 		if (verbose == true) 

### 2016-01-04 doc/tutorial/README.Rmd
snapshot

@@
 
-**SVMBridge** is a small framework for the controlling several SVM packages like LIBSVM, SVMperf, CVM from inside R. 
-The goal is to allow for a rather easy way to do experiments with these different SVM packages. As the framework interacts with these packages by I/O, the focus is not on speed.
+**SVMBridge** is a small framework for the controlling several SVM packages like LIBSVM, SVMperf, CVM via command-line from inside R.
+The goal is to allow for a rather easy way to do experiments with these different SVM packages, e.g.
+by replacing one variant with another one without nearly zero effort.
+As the framework interacts with these packages by file I/O, the focus is not on speed.
 
-In the following we provide a short introduction to **SVMBridge**. An introductory example serves as a quickstart guide.
-Note that our focus is on your comprehension of the basic functions and
-applications. For detailed technical information and manual pages, please refer to
-the package's [manual pages](http://aydindemircioglu.github.io/SVMBridge/man/). They are regularly updated and reflect the documentation of the current package.
+In the following we provide a short introduction to **SVMBridge**. An introductory example serves as a quickstart guide. Note that our focus is on your comprehension of the basic functions and
+applications.
 
-1. [Quickstart](quickstart.md)
-2. [In Depth Introduction](introduction.md)
+1. [Introduction](introduction.md)
+2. [In Depth Introduction](detailed.md)
 3. [Wrappers](wrappers.md)
+4. [Helper functions](helpers.md)
+5. [Windows How-To](windowsHowTo.md)

### 2016-01-04 doc/tutorial/detailed.Rmd
snapshot

--- 2015-03-24 doc/tutorial/introduction.Rmd
@@
 
+# Finding a software package
+
+There are actually two ways to search for wrappers and software..
+
 # Writing your own wrapper
@@
 
+v
 The corresponding function is ```findSVMSoftware```.
 To tell the **SVMBridge** where to find the mySVM software, call it like this: ```findSVMSoftware ("mySVM", searchPath = "./mySVM", verbose = FALSE).
+It will then look into
 
 For convience, there is also a findAllSVMSoftware function that will try to find the software of all registered SVM packages based on the given search path, e.g., if all your SVM software can be found inside the ./software directory, you can call ```findAllSVMSoftware(searchPath = "./software")```.
+
+Notice that if you write your own findSoftware routine for your wrapper, you might need some platform dependent checks, e.g.
+the LIBSVM binray is simply called "svm-train" on Linux and Mac OSX, but "svm-train.exe" on Windows.
+You can use the functionality provided by R to check for the platform.
+An example can be seen in the LIBSVM wrapper code.
 
@@
 
+Conveniently, the **SVMBridge** contains fast routines to read and write Sparse Data Format.
+Contrary to the similar functionality in e1071, our routines work with dense matrices internally, i.e.
+reading a Sparse File will result in a dense matrix object, while writing works only with dense matrices and will produce a Sparse File again.
+In case you need to work with the data in R and it is too big to read it densely, please use e1071 functionality.
+If the sparsity is not too high, our routines will work much faster than e1071, as we adapted the LIBSVM reader to our needs, so we use Rcpp for this.
+
+## Implementing your own data format
+
 To implement your own data format, recall that you have two ways to provide data to your SVM software.
@@
 
-## Speed
-
-It is possible to bypass any extra-I/O and directly communicate with the SVM Package by specifying the training and test data as well as the model and prediction as a file.
-If SVMBridge gets file pathes, it will not try to reread the written model nor the predictions.
-Sometimes you do not need the predictions, neither on memory nor on disk.
-In this case you can pass "/dev/null" as prediction file.
-This will make the SVM Test write to null, so it will not take any I/O time,  and therefore the SVMBridge will not reread the predictions, as it was already written on disk.
-
 # FAQ
@@
 
-## FindBinary stalls
+## FindBinary stalls when called with parameter execute = TRUE
 
-If the binary expects inputs, executing the binary will stall.
+If the binary expects any kind of input, executing the binary will stall.
 Please make sure that your binary does not expect input upon being called without parameters.
-If necessary, turn off automatic search and provide **SVMBridge** with the direct path to the software.
-FIXME: WILL SEND 1 to the program, if the program is called SVMperf. will still stall on other wait-for-key-programs.
+If necessary, turn off automatic search and provide **SVMBridge** the direct path to the software.
+In some cases you might work around this behaviour by explicitly sending an input to the binary, e.g.
+the SVMperf wrapper will send a single key to the SVMperf binary, which will print out some information and waits for a key when called without any other argument.
+Every wrapper provided with the **SVMBridge** in the installation directory should not suffer from this.
+
+# Speed
+
+It is possible to bypass any extra-I/O and directly communicate with the SVM Package by specifying the training and test data as well as the model and prediction as a file.
+If SVMBridge gets file pathes, it will not try to reread the written model nor the predictions.
+Sometimes you do not need the predictions, neither on memory nor on disk.
+In this case you can pass "/dev/null" as prediction file.
+This will make the SVM Test write to null, so it will not take any I/O time,  and therefore the SVMBridge will not reread the predictions, as it was already written on disk.

### 2016-01-04 doc/tutorial/helpers.Rmd
snapshot

**SVMBridge** has some helper functions that might be of use outside the SVM context. We give a short list here, for more, please refer to the corresponding documentation.

- **expandTilde**: This will expand a path containing a tilde (=home directory on Unix-like OSs) into a full path, e.g. '~/R' might expand to '/home/user/R' on Unix or '/Users/user/R' on Windows.

- **..**

Notice that the LIBSVM wrapper is the only one that is directly available when loading the **SVMBridge**. The reason for not including other wrappers directly is to reduce dependencies. But as on the other hand other wrappers depend on the functionality of the LIBSVM wrapper to avoid code duplication, LIBSVM is included into the package.

**Warning:** Currently we do not seek completeness. The **SVMBridge** is a classification-only package, so many options e.g. regarding regression are not supported at all. Some other options, although they can be used with classifications, might be missing too. In case you need something, you will need to adapt an existing classifier or simply write your own.

**Note:** For convenience, all the wrappers apart from LIBSVM are nonetheless installed along the package and can be accessed by specifying the correct path. See the introduction for examples on how to do this.

# LIBSVM

LIBSVM is practially the de-facto standard SVM solver. It can be found at http://

# LASVM

LASVM is very similar to LIBSVM wrapper, as both packages use the same parameters and the same model. LASVM has many more options than LIBSVM, as for our purpose we did not need them, for now these are not supported. Please feel free to enhance the wrapper and share the results with us, so we can include these in future versions of **SVMBridge**.

# LLSVM

Part of the BudgetedSVM package, at http://

# Budgeted SGD

Part of the BudgetedSVM package, at http://

# Pegasos

Part of the BudgetedSVM package, at http://

# SVMperf

Currently the package can be found at http://
Note: Please do not mix SVMperf with SVMlight or SVMRank.

SVMperf has many options, so instead of providing all of these as parameter, we decided instead to give the user a choise over several 'configurations'.

# CVM

libCVM, found at http://

# BVM

Though BVM is practially the same software as CVM, sharing models and parameters and executables with CVM, the philosophy of **SVMBridge** is to have a separate wrapper for each method. Thus, though BVM has its own wrapper, all the parameters etc are the same as CVM, so please read the CVM section.

There are several other packages, namely

### 2016-01-04 doc/tutorial/introduction.Rmd
snapshot

--- 2015-03-22 doc/tutorial/introduction.Rmd
@@
-Writing your own wrapper is also possible, and not too complicated.
-You must provide several callback functions.
-The most complex part will be provide functions to read and write the model.
-As a starter, you might want to look into the wrappers, **SVMBridge** provides.
+# Introduction
 
-# SVM Parameter
+**Info:** This guide gives you a short overview and an example of a typical call of some SVM Package with **SVMBridge**. For much more detailed information see [the next chapter](detailed.md).
 
-Every SVM has some parameter, tha usually must be tuned.
-While LIBSVM has (for RBF kernel) two prominent parameter, the regularization term $C$ and the kernel bandwidth $\gamma$, this might not be true for other methods.
-For example, Budgeted SVM operates with $\lambda$ instead of $C$, with $\lambda = \frac{1}{2 C}$, and the $\gamma$ is twice the gamma of LIBSVM.
-Because of this, **SVMBridge** does not have a prebuild set of parameters.
-Every wrapper is free to use the parameter it needs.
-This entails that to use a wrapper, you must first know which parameter it uses.
-You can use the help function to get a description of a wrapper.
+Here we provide a Quickstart example for you to make yourself familiar with the **SVMBridge**.
+We aim to use the LIBSVM package for a simple SVM classification.
+Remember, that a general machine learning problem usually consists of two parts: Training and Testing.
 
-LIBSVMHelp() { }
-
-# Training
-
-Training can be either performed from data in memory (probably the most useful case), from the file system by using the ```trainSVM``` function. 
-Note: In case you specify data from memory and from disc, **SVMBridge** will throw an error, as it will be confused about which source to use.
-
-# Testing
-
-Testing is done very similar to training by calling the ```testSVM``` function You have the option to use data from memory, from file system.
-Furthermore, predictions can be provided.
-By default, the predictions are written into a temporary file and are discarded after testing.
-Note: In case you specify data from memory and from disc, **SVMBridge** will throw an error, as it will be confused about which source to use.
-
-# Model file
-
-If you do not specify any model file for training, ```trainSVM``` will ask R for a temporary file and will provide this path to the SVM package.
-Internally, the ```writeModel``` routine of the SVM package will be used to create a model file from a model in memory.
-In case you have not specified a model file path, after training the model will be read from the temporary model file into the SVM Results Object.
-Contrary, if you did specify a model file path, then by default, the model is not read back into the SVM Results Object.
-To enforce reading the model, you can set the argument ```readModel``` to true.
-
-At testing time, you must provide the trained model.
-If you pass this as an in-memory model, ```testSVM``` will call the ```readModel``` routine of the SVM package first to dump the model to a temporary file on the filesystem.
-
-skipModel == TRUE
+As **SVMBridge** is just an easy way to call a precompiled SVM package (also called a 'wrapper'), we first need such an SVM package.
+In this introduction, we simply use the probably best known SVM package, LIBSVM.
+Download the source package from LIBSVM webpage (currently at http://www.csie.ntu.edu.tw/~cjlin/libsvm/).
+Proceed by downloading the package as zip/tar or directly from github.
+Put it into any folder you like.
+We here assume that you have a software directory and unzipped LIBSVM under 'software/LIBSVM'.
+Then go to this directory and compile LIBSVM, by calling 'make' from the command line (Make sure to have all the necessary c++ packages installed).
+After successful compilation you will have two binary executables: svm-train and svm-predict.
+Make sure you can execute them by just calling './svm-train' (an extensive help page should be shown).
+You also need some training and test data.
+The LIBSVM currently provides many different benchmark data sets (currently at 'http://www.csie.ntu.edu.tw/~cjlin/libsvmtools/datasets/').
+To keep things simply, we will use the same data set for testing and training.
+Download the scaled australian data set (currently at 'http://www.csie.ntu.edu.tw/~cjlin/libsvmtools/datasets/binary/australian_scale') and put it into a folder, we assume here that you download it to  'software/data/australian_scale'.
+Now you can write your first simple R-script:
 
@@
 library(SVMBridge)
+
+# let SVMBridge find the LIBSVM package. We assume here that you are in the './software' directory
+findSVMSoftware (method = 'LIBSVM', searchPath = "./LIBSVM")
+
+# use the SVMBridge to train LIBSVM
+svmObj =  callSVM(method = "LIBSVM", trainDataFile = './data/australian_scale', cost = 1, gamma = 1)
 ```
 
-# Predictions
+**Note:** In case something goes wrong, you can add the verbose = TRUE flag to the callSVM function.
+Most probably your pathes are wrong, e.g.
+**SVMBridge** cannot find the LIBSVM executable or the data set.
 
-After testing, the predictions will again be stored in the SVM Object and can be accessed by e.g.
-```SVMObj$predictions```.
-Depending on the purpose, the predictions are not always needed, for example when tuning the SVM parameters.
-To avoid extra I/O and memory consumption, reading back the predictions can be turned of by using the ```skipPredictions``` options.
-By default, ```skipPredictions``` is false, so ```callSVM``` will return the predictions in the SVM Object.
+The ```callSVM``` function will return an object that contains basically the training error, testing time (including I/O) and the final SVM model.
+See ...
+for more information about the SVM object, callSVM returns.
+We are going to print the time LIBSVM needed for training and then go ahead and test it on the same data set (for convenience, in  general testing on the training set will greatly underestimate the generalization error).
+For that we need to pass the trained model to the callSVM routine.
+Notice, that callSVM will automatically skip training, if no training file is given:
 
-skipPredictions == TRUE
+```splus
+# print the time needed for training
+print( paste("LIBSVM took", model$trainTime, "seconds to train."))
 
-# Writing your own wrapper
+# use the SVMBridge to test LIBSVM (here on the same data set)
+svmObj =  callSVM(method = "LIBSVM", testDataFile = './data/australian_scale', model = svmObj$model)
+```
 
-The method name will define all function names!
-This is very important, if you do not abide to this rule, callSVM will not find the corresponding callbacks!
-Suppose you want to write a wrapper for mySVM.
-In this case, you will need to create a file exactly called 'mySVM_wrapper.R'.
-Please note that the filename is case-sensitive.
-This is also true for all the callback functions.
+**SVMBridge** comes in handy, when one has the data already in memory. Let us generate some training and test data by using the build-in data set iris:
 
-Your mySVM software must be split into two parts: Training and Testing.
-This might be the same executable, but be aware, that testing must work without training data.
-In case your package needs testing data at training time (e.g.
-if the model only specifies the index of the support vector in the training file instead of copying them into the model), you need to perform quite heavy tricks to get this going.
-In general, using other packages and skip **SVMBridge** might be a good idea, as **SVMBridge** strictly follows the LIBSVM way of training and testing.
-
-## Adding the wrapper
-
-After you have written a wrapper, say 'mySVM_wrapper.R', you must add your wrapper to the **SVMBridge**.
-This you can either manually or automatically (see the next section for information about this).
-There are two parts of this procedure: First you must tell **SVMBridge** where to find your wrapper 'mySVM_wrapper.R'.
-After this you must specify where the corresponding binaries for mySVM are to be found.
-Remember that the **SVMBridge** acts mainly as a convenient tool to call a SVM solver via command line.
-
-To add your wrapper to the **SVMBridge** simply call ```addSVMPackage```, e.g.
-if mySVM resides in the mySVM subdirectory, you can call ```addSVMPackage (method = "mySVM", filePath = "./mySVM/mySVM_wrapper.R", softwarePath = "./mySVM/bin")```.
-Notice that filePath is mandatory, and the **SVMBridge** will source the specified wrapper.
-Though you are free to name your wrapper as you wish, it is a good idea to stick to the convention chosen and name the wrapper for your SVM package by adding '_wrapper.R', e.g.
-the wrapper for LASVM is called 'LASVM_wrapper.R' inside **SVMBridge**.
-The path for the binaries of mySVM is not mandotory.
-Instead of specifiying it, **SVMBridge** can automatically find it (based on the code in the mySVM_wrapper.R file).
-
-## Automatically finding the software
-
-After you added your wrapper, if you did not specified a softwarePath, **SVMBridge** can find the corresponding binary files for you.
-Remember that if you forget to specifiy the softwarePath,  **SVMBridge** will not understand where to find the binaries.
-
-The corresponding function is ```findSVMSoftware```.
-To tell the **SVMBridge** where to find the mySVM software, call it like this: ```findSVMSoftware ("mySVM", searchPath = "./mySVM", verbose = FALSE).
-
-For convience, there is also a findAllSVMSoftware function that will try to find the software of all registered SVM packages based on the given search path, e.g., if all your SVM software can be found inside the ./software directory, you can call ```findAllSVMSoftware(searchPath = "./software")```.
-
-## Automatically finding the wrapper
-
-Sometimes it is convenient to allow the user to specify the SVM software he needs.
-In this case it is no good to specify neither the wrapper nor the software by hand.
-Therefore **SVMBridge** allows you not only to automatically find the software, but also the wrapper.
-This only works, if the wrapper follows the naming convention, i.e.
-the wrapper for mySVM needs to be called mySVM_wrapper.R.
-To initiate the search, just call ```findSVMWrapper (method = "mySVM", searchPath = "./mySVM")```.
-The specified searchPath will, just as ```findSVMSoftware```, be searched recursively.
-Notice also that if multiple such wrappers exist inside a directory (in subdirectories), **SVMBridge** will stop searching after the very first hit.
-
-## The training parameters
-
-You will get a bunch of parameters.
-You will need to decide which ones you want to handle.
-E.g.
-the LIBSVM wrapper handles $C$, $\gamma$ and $\epsilon$, but also uses $kernelCacheSize$.
-On the other hand, the LIBSVM wrapper currently has no option to control the shrinking, although LIBSVM has a parameter for this.
-Similarly you will need to decide, whether you want the user to have control over all parameters or only a relevant subset.
-The callback will then gather all the parameter and create a command-line string.
-
-# SVM Result Object
-
-Actually we do enforce that.
-We do not enforce a structure on the returned SVM Return Object.
-Nonetheless, we encourage you to use the same structure we have provided in our example wrapper, if that is applicable to your case.
-The SVM Return Object contains the following fields:
- -execTime
- -error
-Depending on training or testing, this is the training or testing time, measured in wall time by microbenchmark.
-
-## Test parameters
-
-Very similar to training parameters, the test parameter callback will assemble a given model file and test data and possibly other options.
-
-extraParameters for testing??
-
-# Data Formats
-
-As nearly all SVM Packages work with the LIBSVM/SVMlight Sparse Format, this is the basic file format in SVMBridge.
-Notice that SVMBridge actually does neither enforce nor check this explicitly.
-You can support your own data format for your own SVM wrapper.
-In this case, keep in mind that other SVM packages might not be able to work with your data.
-
-To implement your own data format, recall that you have two ways to provide data to your SVM software.
-The first is by specifiying a path to a data file on the filesystem.
-In this case, SVMBridge will not explicitly work with the data, so you are free to use any format, as long as your SVM software can handle this.
-The second way is to pass data as a matrix to the SVMBridge.
-Here, SVMBridge needs to dump the given data first.
-By default, the SVMBridge will write the data in the LIBSVM Sparse Format.
-You are free to overwrite this behaviour, by rewriting the convertData function of the SVM Wrapper Object.
-
-Let us give an easy example.
-
-LABELS..
-MULTICLASS..
-
-# Vocabulary
-
-The SVM software is the compiled object on the file system, e.g.
-when you download and compile LIBSVM you will end  up with two binaries, one for training, one for testing.
-These files is an SVM software for us.
-
-Glue code that handles controlling the SVM software is called an SVM wrapper. This contains e.g. routines for reading and writing an SVM model.
-
-Finally, an SVM package is for us the internal object in the SVMBridge, that has basically consists of the wrapper and the software (e.g. a path to the binary).
-
-## Speed
-
-It is possible to bypass any extra-I/O and directly communicate with the SVM Package by specifying the training and test data as well as the model and prediction as a file.
-If SVMBridge gets file pathes, it will not try to reread the written model nor the predictions.
-Sometimes you do not need the predictions, neither on memory nor on disk.
-In this case you can pass "/dev/null" as prediction file.
-This will make the SVM Test write to null, so it will not take any I/O time,  and therefore the SVMBridge will not reread the predictions, as it was already written on disk.
+**Warning:** The findSVMSoftware functionality is made for convenience, not for performance.
+As all given directories has to be searched, this can slow down all I/O operations.
+Prevent running this function on a high performance cluster often, as might very well slowdown the whole cluster!

### 2016-01-04 R/BSGD_wrapper.R

@@ -316,29 +316,48 @@ writeModel.BSGD = function (x, model = NA, modelFile = "./model", verbose = FALS
 	if (verbose == TRUE) {
 		cat ("Writing BSGD Model..\n")
 	}
-
-	# BROKEN, FIXME
+ 
+	dim = ncol (model$SV)
+	nClasses = 2
+	nWeights = nrow (model$SV)
+	gamma = model$gamma * 2
+	degree = 1
+	coef = 1
 	
-	# FIXME: label order
-	# TODO: support multiclass
-	model$nrclass = 2
-	posSV = sum(model$a > 0)
-	negSV = sum(model$a < 0)
+	print (as.character (paste (model$label, sep = " ", collapse = " ")))
+	
     # open connection
     modelFileHandle <- file(modelFile, open = "w+")
-	writeLines(paste ("svm_type c_svc", sep = ""), modelFileHandle )
-	writeLines(paste ("kernel_type", "rbf", sep = " "), modelFileHandle )
-	writeLines(paste ("gamma", model$gamma, sep = " "), modelFileHandle )
-	writeLines(paste ("nr_class", model$nrclass, sep = " "), modelFileHandle )
-	writeLines(paste ("total_sv", length(model$a), sep = " "), modelFileHandle )
-	writeLines(paste ("rho", model$bias, sep = " "), modelFileHandle )
-	writeLines(paste ("label 1 -1", sep = " "), modelFileHandle )
-	writeLines(paste ("nr_sv", posSV, negSV, sep = " "), modelFileHandle )
-	writeLines(paste ("SV", sep = ""), modelFileHandle )
+	writeLines(paste ("ALGORITHM: 4", sep = ""), modelFileHandle )
+	writeLines(paste ("DIMENSION:", dim, sep = " "), modelFileHandle )
+	writeLines(paste ("NUMBER_OF_CLASSES:", nClasses, sep = " "), modelFileHandle )
+	writeLines(paste ("LABELS:", as.character (paste (model$label, sep = " ", collapse = " ")), sep = " "), modelFileHandle )
+	writeLines(paste ("NUMBER_OF_WEIGHTS:", nWeights, sep = " "), modelFileHandle )
+	writeLines(paste ("BIAS_TERM: ", model$bias, sep = " "), modelFileHandle )
+	writeLines(paste ("KERNEL_FUNCTION: 0", sep = " "), modelFileHandle )
[…]

### 2016-01-04 R/LIBSVM_wrapper.R

@@ -294,68 +294,71 @@ readPredictions.LIBSVM = function (x, predictionsFile = "", verbose = FALSE) {
 	return (predictions)
 }
 
-	
+
+
 #' findSoftware.LIBSVM
+#' Will find the LIBSVM binaries.
 #'
-#' @param	x			svm object
-#' @param	searchPath		path to search for software
-#' @param	verbose			be verbose?
+#' @param	x	svm object
+#' @param	searchPath	path to search for software
+#' @param	execute		shall the binaries  be executed to make sure they are the right one?
+#' @param	verbose		be verbose?
 #'
-#' @return	x			svm object
+#' @return	x	svm object with  x$testBinaryPath and x$trainBinaryPath set.
+#' 
+#' @note 		will produce an error if the executable could not be found.
 #'	
-findSoftware.LIBSVM = function (
-	x,
-	searchPath = "./",
-	verbose = FALSE)
-	{
-		if (verbose == TRUE) {
-			BBmisc::messagef("    LIBSVM Object: Executing search for software for %s", x$method)
-		}
-		
-		if(.Platform$OS.type == "unix")
-			trainBinaryPattern = "^svm-train$"
-		else
-			trainBinaryPattern = "^svm-train.exe"
-			
-		trainBinaryOutputPattern = c('saveExponential : set exponential',
-			'.q : quiet mode .no outputs')
+findSoftware.LIBSVM = function (x, searchPath = "./", execute = TRUE, verbose = FALSE) {
 
-		binaryPath = findBinary (searchPath, trainBinaryPattern, trainBinaryOutputPattern, verbose = verbose)
+	if (verbose == TRUE) {
+		cat("    LIBSVM Object: Executing search for software for %s", x$method)
+	}
 		
-		if (verbose == TRUE) {
-			BBmisc::messagef("Executing search for binaries in:  %s", searchPath) 
-		}
-		
-		if (verbose == TRUE) {
-			BBmisc::messagef("--> Found train binary at %s", binaryPath) 
-		}
-		x$trainBinaryPath = binaryPath
+	if(.Platform$OS.type == "unix") {
+		trainBinaryPattern = "^svm-train$"
+		testBinaryPattern = "^svm-predict$"
+	} else {
+		trainBinaryPattern = "^svm-train.exe"
+		testBinaryPattern = "^svm-predict.exe"
+	}
 
-		if(.Platform$OS.type == "unix")
[…]

### 2016-01-04 R/SVMWrapper.R

@@ -32,92 +32,90 @@
 #' @param b Description
 #' ...
 
-	createSVMWrapper = function(method, ...) {
-		
-		# check that trainingBinary exists, if given
-		# ...    
-		
-		# register the newly created object in our environment
-		
-		# create S3 Object
-		svmPackage = BBmisc::makeS3Obj(c(method, "SVMWrapper"),
-			method = method
-		)
-
-		# TODO: replace existing object in package list
-		
-		# add object to our internal list of created wrappers
-	#	SVMBridgeEnv$packages = c(SVMBridgeEnv$packages, svmPackage)
-		
-		return (svmPackage)
-	}
+createSVMWrapper = function(method, ...) {
+	# check that trainingBinary exists, if given
+	# ...    
 	
+	# register the newly created object in our environment
+	
+	# create S3 Object
+	svmPackage = BBmisc::makeS3Obj(c(method, "SVMWrapper"),
+		method = method
+	)
+
+	# TODO: replace existing object in package list
+	
+	# add object to our internal list of created wrappers
+#	SVMBridgeEnv$packages = c(SVMBridgeEnv$packages, svmPackage)
+	
+	return (svmPackage)
+}
+
+
+
+
+createTrainingArguments = function (x,
+	trainDataFile = "",
+	modelFile = "",
+	extraParameter = "",
+	kernelCacheSize = 1024,
+	cost = 1, 
+	useBias = FALSE,
+	gamma = 1,
+	epsilon = 0.001, 
+	...) 
+{
+	UseMethod("createTrainingArguments")
+}
+
+
+
+createTestArguments = function (x,
+	testDataFile = "",
+	modelFile = "", 
+	predictionsFile = "",
+	...) 
+{
+	UseMethod("createTestArguments")
+}
+
+
+
+extractTrainInfo = function (x, output) {
+	UseMethod ("extractTrainInfo")
+}
+
+
+
+extractTestInfo = function (x, output) {
+	UseMethod ("extractTestInfo")
+}
+
+
+
+readModel = function (x, modelFile = './model', verbose = FALSE) {
+	UseMethod("readModel")
+}
+
+
+
+writeModel = function (x, model = NA, modelFile = "./model", verbose = FALSE) {
+	UseMethod("writeModel")
+}
+
+
+detectModel = function (x, modelFile = "./model", verbose = FALSE) {
+	UseMethod("detectModel")
[…]

### 2016-01-04 R/addSVMPackage.R

@@ -85,7 +85,7 @@ addSVMPackage <- function (method = NA, filePath = NA, softwarePath = NA,
 				}
 			} else {
 				if (verbose == TRUE)
-					cat ("Did not find training binary.\n")
+					cat ("Did not find training binary: ", trainBinaryPath, "\n")
 			}
 		} else {
 			if (verbose == TRUE)
@@ -112,7 +112,7 @@ addSVMPackage <- function (method = NA, filePath = NA, softwarePath = NA,
 				}
 			} else {
 				if (verbose == TRUE)
-					cat ("Did not find testing binary.\n")
+					cat ("Did not find testing binary: ", testBinaryPath, "\n")
 			}
 		} else {
 			if (verbose == TRUE)

### 2016-01-04 R/expandTilde.R

@@ -1,4 +1,3 @@
-#!/usr/bin/Rscript  --vanilla 
 #
 # SVMBridge 
 #		(C) 2015, by Aydin Demircioglu
@@ -22,14 +21,14 @@
  
 #' Expand path with tilde 
 #'	
-#' Given a path, try to expand the path, crossplatform independent.
+#' Given a path, try to expand the path in a crossplatform independent way.
 #'
 #' @param 	path		path to expand
 #' @param	verbose			print messages while searching?
 #'
 #' @export
 expandTilde <- function(path = NULL, verbose = FALSE){
-	#look for tilde characters and expand them
+	# look for tilde characters and expand them
 	if(grepl("~", path) == TRUE){
 		if(.Platform$OS.type == "windows"){
 			firstPart = path.expand("~")
@@ -42,7 +41,7 @@ expandTilde <- function(path = NULL, verbose = FALSE){
 		path = gsub("[\\]", "/", path)
 		
 		if(verbose == TRUE){
-			BBmisc::messagef("  Expanded path: %s", path)
+			cat("    Expanded path: %s", path)
 		}
 	}
 	return (path)

### 2016-01-04 R/findAllSVMSoftware.R

@@ -1,0 +1,44 @@
+#
+# SVMBridge 
+#		(C) 2015, by Aydin Demircioglu
+#
+#		findAllSVMSoftware.R
+# 
+# SVMBridge is free software: you can redistribute it and/or modify
+# it under the terms of the GNU Lesser General Public License as published
+# by the Free Software Foundation, either version 3 of the License, or
+# (at your option) any later version.
+#
+# SVMBridge is distributed in the hope that it will be useful,
+# but WITHOUT ANY WARRANTY; without even the implied warranty of
+# MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
+# GNU Lesser General Public License for more details.
+#
+# Please do not use this software to destroy or spy on people, environment or things.
+# All negative use is prohibited.
+#
+
+
+
+#' findAllSVMSoftware 
+#'		Given a search path, it will try to find the software packages
+#'		for all registered SVM package by calling the 'findSoftware' routine of each known wrapper.
+#'
+#' @param 	searchPath	 	search the given path for the SVM binaries of all known SVM packages.
+#' @param	verbose			print messages while searching?
+#'
+#' @note	To make sure that the binary is correct, it will be executed! (see findBinary for more infos)
+#' @note	If multiple binaries are found, the last one will be taken. Overwrite by hand, if necessary.
+#'
+#' @export
+findAllSVMSoftware <- function (searchPath = NA, verbose = FALSE) {
+	if (verbose == TRUE) {
+		cat("-Finding all SVM software packages:")
+	}
+	
+	for (i in seq(1, length(SVMBridgeEnv$packages))) {
+		method = SVMBridgeEnv$packages[[i]]$method
+		findSVMSoftware (method = method, searchPath = searchPath, verbose = verbose)
+	}
+}
+

### 2016-01-04 R/findBinary.R

@@ -35,6 +35,7 @@
 #					but it might be hopeless. With luck, the found binary will be left untouched,
 #					and thus work, if not, you must set the path by hand.
 #' @note		If multiple binaries are found, the last one will be taken. Overwrite by hand, if necessary.
+#' @export
 
 findBinary <- function (searchPath, pattern, outputPattern, verbose = FALSE) {
 	if (verbose == TRUE) { BBmisc::messagef("  Checking for pattern %s", pattern) }

### 2016-01-04 R/findSVMSoftware.R

@@ -1,4 +1,3 @@
-#!/usr/bin/Rscript  --vanilla 
 #
 # SVMBridge 
 #		(C) 2015, by Aydin Demircioglu
@@ -20,67 +19,32 @@
 #
 
 
-#' findAllSVMSoftware 
-#'		given a search path, it will try to find the corresponding software packages
-#'		for all registered SVM packages.
-#'
-#' @param 	searchPath	 	search the given path for the SVM binaries of all known SVM packages.
-#' @param	verbose			print messages while searching?
-#'
-#' @note		To make sure that the binary is correct, it will be executed! (see findBinary for more infos)
-#' @note		If multiple binaries are found, the last one will be taken. Overwrite by hand, if necessary.
-#'
-#' @export
-findAllSVMSoftware <- function (searchPath = NA, verbose = FALSE) {
-	if (verbose == TRUE) {
-		BBmisc::messagef("API: Searching for all software packages:")
-	}
-	
-	for (i in seq(1, length(SVMBridgeEnv$packages))) {
-		method = SVMBridgeEnv$packages[[i]]$method
-		if (verbose == TRUE) {
-			BBmisc::messagef("  Searching for software for SVM package %s:", method)
-		}
-		findSVMSoftware (method = method, searchPath = searchPath, verbose = verbose)
-	}
-}
-
-
 
 #' findSVMSoftware 
 #'		given a search path, it will try to find the corresponding software packages
 #'		for the given method.
 #'
-#' @param	method		name of the SVM method
-#' @param 	searchPath	 	search the given path for the SVM binaries of the given SVM method.
+#' @param	method	   name of the SVM method
+#' @param   searchPath   search the given path for the SVM binaries of the given SVM method.
 #' @param	verbose		print messages while searching?
+#' 
+#' @note    calls the findSoftware routine of the corresponding wrapper.
+#'
 #' @export	findSVMSoftware
 findSVMSoftware <- function (method = NA, searchPath = NA, verbose = FALSE) {
-	if (verbose == TRUE) {
-		BBmisc::messagef("API: Finding software for %s", method)
-	}
-	
-	if (is.na(searchPath)) {
-		BBmisc::stopf("No search path is given!")
-	}
-	
-	if (is.na(method)) {
-		BBmisc::stopf ("No method name is given")
-	}
-	
-	if (verbose == TRUE) {
-		BBmisc::messagef("  Try to find binaries for %s", method) 
-	}
-	
-	#look for tilde characters and expand them
[…]

### 2016-01-04 R/registerSVMPackage.R

@@ -1,8 +1,0 @@
-
-registerSVMPackage = function (package) {
-	# check for inheritance?
-	# check for all functions
-	
-	# add package to our list
-
-}

### 2016-01-04 compile.R

@@ -1,16 +1,23 @@
+#!/usr/bin/Rscript  --vanilla 
+
+# work around
+    system2(command="./cleanup", args="")
+    
 library(methods)
 library(devtools)
 library(roxygen2)
+library(Rcpp)
+compileAttributes ()
 
 roxygenize(".")
 load_all (".")
 build_vignettes(".")
 document(".")
-#devtools::build_win() 
-
-# is done by document/loadall anyway?
-library(Rcpp)
-compileAttributes()
 
 devtools::test()
 devtools::check()
+
+
+# work around
+system2(command="R", args="CMD build .")
+system2(command="R", args="CMD check --as-cran ./LiblineaR.ACF_1.94-2.tar.gz")

### 2016-01-04 inst/wrapper/BSGD_wrapper.R

@@ -316,34 +316,82 @@ writeModel.BSGD = function (x, model = NA, modelFile = "./model", verbose = FALS
 	if (verbose == TRUE) {
 		cat ("Writing BSGD Model..\n")
 	}
-
-	# BROKEN, FIXME
+ 
+	dim = ncol (model$SV)
+	nClasses = 2
+	nWeights = nrow (model$SV)
+	gamma = model$gamma * 2
+	degree = 1
+	coef = 1
 	
-	# FIXME: label order
-	# TODO: support multiclass
-	model$nrclass = 2
-	posSV = sum(model$a > 0)
-	negSV = sum(model$a < 0)
+	print (as.character (paste (model$label, sep = " ", collapse = " ")))
+	
     # open connection
     modelFileHandle <- file(modelFile, open = "w+")
-	writeLines(paste ("svm_type c_svc", sep = ""), modelFileHandle )
-	writeLines(paste ("kernel_type", "rbf", sep = " "), modelFileHandle )
-	writeLines(paste ("gamma", model$gamma, sep = " "), modelFileHandle )
-	writeLines(paste ("nr_class", model$nrclass, sep = " "), modelFileHandle )
-	writeLines(paste ("total_sv", length(model$a), sep = " "), modelFileHandle )
-	writeLines(paste ("rho", model$bias, sep = " "), modelFileHandle )
-	writeLines(paste ("label 1 -1", sep = " "), modelFileHandle )
-	writeLines(paste ("nr_sv", posSV, negSV, sep = " "), modelFileHandle )
-	writeLines(paste ("SV", sep = ""), modelFileHandle )
+	writeLines(paste ("ALGORITHM: 4", sep = ""), modelFileHandle )
+	writeLines(paste ("DIMENSION:", dim, sep = " "), modelFileHandle )
+	writeLines(paste ("NUMBER_OF_CLASSES:", nClasses, sep = " "), modelFileHandle )
+	writeLines(paste ("LABELS:", as.character (paste (model$label, sep = " ", collapse = " ")), sep = " "), modelFileHandle )
+	writeLines(paste ("NUMBER_OF_WEIGHTS:", nWeights, sep = " "), modelFileHandle )
+	writeLines(paste ("BIAS_TERM: ", model$bias, sep = " "), modelFileHandle )
+	writeLines(paste ("KERNEL_FUNCTION: 0", sep = " "), modelFileHandle )
[…]

### 2016-01-04 inst/wrapper/BVM_wrapper.R

@@ -93,6 +93,37 @@ extractTestInfo.BVM = function (x, output) {
 		return (ret)
 	}
  
+ 
+
+#' Detect whether a file is a model for LIBSVM.
+#'
+#' @param	x		Object
+#' @param	modelFile		File to check 
+#' @param	verbose		Verbose output?
+#'
+#' @return	TRUE if the given modelFile exists and fits the LIBSVM model, or FALSE if not.
+#'
+#' @note	This is a very basic check, enough to distinguish the wrappers provided within the SVMBridge
+
+detectModel.BVM = function (x, modelFile = NULL, verbose = FALSE) {
+	checkmate::checkFlag (verbose)
+	if (is.null (modelFile) == TRUE) 
+		return (FALSE)
+	
+	# read first lines and detect magic marker
+	if (file.exists (modelFile) == FALSE) 
+		return (FALSE)
+		
+	line = readLines(modelFile, n = 1)
+	if (line == "svm_type bvm") {
+		return (TRUE)
+	}
+	
+	return (FALSE)
+}
+
+ 
+ 
 #
 # @param[in]	predictionsFile		file to read predictions from
 # @return		array consisting of predictions

### 2016-01-04 inst/wrapper/LASVM_wrapper.R

@@ -103,6 +103,34 @@
 		return (ret)
 	}
 	
+
+#' Detect whether a file is a model for LASVM.
+#'
+#' @param	x		Object
+#' @param	modelFile		File to check 
+#' @param	verbose		Verbose output?
+#'
+#' @return	TRUE if the given modelFile exists and fits the LASVM model, or FALSE if not.
+#'
+#' @note	This is a very basic check, enough to distinguish the wrappers provided within the SVMBridge
+
+detectModel.LASVM = function (x, modelFile = NULL, verbose = FALSE) {
+	checkmate::checkFlag (verbose)
+	if (is.null (modelFile) == TRUE) 
+		return (FALSE)
+	
+	# read first lines and detect magic marker
+	if (file.exists (modelFile) == FALSE) 
+		return (FALSE)
+
+	line = readLines(modelFile, n = 1)
+	if (line == "svm_type c_svc") {
+		return (TRUE)
+	} 
+	
+	return (FALSE)
+}
+
 	
 	
 	#

### 2016-01-04 inst/wrapper/LLSVM_wrapper.R

@@ -4,7 +4,6 @@
 
 
 
-# evalLLSVM
 # @param[in]    trainDataFile       file to read training data from
 # @param[in]    testDataFile        file to read test data from
 # @param[in]    cost            cost parameter C
@@ -12,8 +11,6 @@
 # @param[in]    rank            number of landmarks
 # @param[in]    bindir          relativ path to the binaries, defaults to default.
 # @param[in]    modelFile       path to model, defaults to a temporary file (given by R)
-
-
 
 createTrainingArguments.LLSVM = function (x,
 								trainDataFile = "",
@@ -75,6 +72,7 @@ extractTrainingInfo.LLSVM = function (x, output) {
 }
 
 
+
 extractTestInfo.LLSVM = function (x, output) {
     
     # ---- grep the error rate
@@ -84,10 +82,12 @@ extractTestInfo.LLSVM = function (x, output) {
     return (err)
 }
 
+
+
 readModel.LLSVM = function (x, modelFile = "./model", verbose = FALSE)
 {
 	if (verbose == TRUE) {
-		BBmisc::messagef("Reading LLSVM model from %s", modelFile)
+		cat("Reading LLSVM model from ", modelFile, "\n")
 	}
 
     # open connection
@@ -107,7 +107,8 @@ readModel.LLSVM = function (x, modelFile = "./model", verbose = FALSE)
             pattern <- "KERNEL_GAMMA_PARAM: (.*)"
             gamma = as.numeric(sub(pattern, '\\1', oneLine[grepl(pattern, oneLine)])) 
         }  
-      
+
+        
         # bias
         if (grepl("BIAS_TERM", oneLine) == TRUE) 
         {
@@ -136,40 +137,146 @@ readModel.LLSVM = function (x, modelFile = "./model", verbose = FALSE)
     }
   
   
-	# read and interprete data 
-	# basically all data is sparse data format, but the data around this differs
-	svmatrix = readSparseFormat(con)
-
-	# add header information
-	svmatrix$gamma = gamma
-	svmatrix$bias = bias
-	svmatrix$modelType = "LLSVM"
+	# these will contain the coefficients and the  svs.
+	supportvectors <- matrix()
+	coefficients <- matrix()
+	weights <- matrix()
 	
+	while (length(oneLine <- readLines(con, n = 1, warn = FALSE)) > 0) {
+	
+		# remove comment if necesary
[…]

### 2016-01-04 inst/wrapper/SVMperf_wrapper.R

@@ -306,8 +306,8 @@ extractTestInfo.SVMperf = function (x, output) {
 readModel.SVMperf <- function (x, modelFile = "./model", verbose = FALSE)
 {
 	if (verbose == TRUE) {
-			BBmisc::messagef ("Reading SVMperf model from %s.", modelFile)
-		}
+		cat ("Reading SVMperf model from ", modelFile, "\n")
+	}
     # open connection
     con  <- file(modelFile, open = "r")
 
@@ -349,6 +349,23 @@ readModel.SVMperf <- function (x, modelFile = "./model", verbose = FALSE)
 	
 	# close connection
 	close(con)
+
+	# this is only true for multiclass, but SVMperf does not support multiclass--
+	# we must sort the alphas, else computation of optimize values will not work
+	# as LIBSVMs alphas are sorted and prediction depend on this (because of multiclass
+	# and the special format LIBSVM saves the alphas)
+	sortSVs = FALSE
+	if (sortSVs == TRUE) {
+		s = sort (model$alpha, index.return = TRUE, decreasing = TRUE)
+		model$alpha[,1] = s$x
+		model$SV = model$SV[s$ix,]
+	}
+	
+	# invert the alphas...
+	#model$alpha[1:model$nSV[1],] = - model$alpha[1:model$nSV[1],]
+
+	#model$alpha[(1+model$nSV[1]):nrow(model$alpha),] = - model$alpha[(1+model$nSV[1]):nrow(model$alpha),]
+
 	
 	# return
 	return (model)
@@ -358,10 +375,84 @@ readModel.SVMperf <- function (x, modelFile = "./model", verbose = FALSE)
 
 # DUMMY
 writeModel.SVMperf <- function (x, model = NA, modelFile = "./model", verbose = FALSE) {
-	ret = writeModel.LIBSVM (model = model, modelFile = modelFile, verbose = verbose)
-		return (ret)
+	
+	if (verbose == TRUE) {
+		cat ("Writing SVMperf Model to ", modelFile, "\n")
+	}
+	
+	# check of S3 object here.
+	
+	model$nrclass = length(model$label)
+	posSV = sum(model$alpha > 0)
+	negSV = sum(model$alpha < 0)
+
+	# open connection
+	if (verbose == TRUE)
+		cat ("  Writing Header\n")
+	modelFileHandle <- file(modelFile, open = "w+")
+	writeLines ("SVM-light Version V6.20", modelFileHandle)
+	writeLines("2 # kernel type", modelFileHandle) #kernel type
[…]

### 2016-01-04 tests/testthat/test_readModelFromFile.R

@@ -19,6 +19,8 @@ test_that("reading a model is working.", {
 		# detect model type
 		modelFile = paste0 ("../data/", solver, ".australian.model")
 		model = readModelFromFile (modelFile)
+		if (solver == "BSGD") 
+			writeModelToFile (model, modelFile = "./TMPBSGD.bsgd", verbose = TRUE)
 
 		# check only information that is there in all models
 		if ((solver == "BSGD") || (solver == "LLSVM"))

### 2016-01-04 doc/tutorial/detailed.Rmd
restructuring

@@
 
+# Advanced
+
+In some cases you might to want to modify the internal object corresponding to a method.
+**SVMBridge** stores these in its own environment.
+You can access these Conveniently by using ```getSVMObject()```.
+After modifying them, you can store the object back by overwriting ```setSVMObject()``, so that the SVMBridge can work with it.
+Notice that broken objects might lead to unforseen behaviour and crashes.
+
 # Vocabulary

### 2016-01-04 R/addSVMPackage.R

@@ -20,172 +20,94 @@
 #
 
 
-#' add package to the bridge
+#' Make a new SVM package known to the bridge.
+#'
+#' This function will create an object corresponding to the SVM object.
+#' By using the findSVMSoftware/findSVMWrapper functions., one can add 
+#' search for the corresponding wrappe and the software.
+#' To avoid lengthy searches,  this function can be used to directly 'shortcut'
+#' the search, by specifiying the directories where the wrapper/software lies.
+#' If this is successful, later findSVM.. calls are unneccessary.
 #'
 #' @param 	method		name of solver
-#' @param	filePath		path to wrapper
-#' @param	softwarePath		path where to find the solver.
-#' @param 	trainBinaryPath		if not NULL this will override any search option
-#' @param 	testBinaryPath		if not NULL this will override any search option
-#' @param	wrapperPath		if not NULL this will override any search option
+#' @param	wrapperName		name of the wrapper (as filename). if none given, method_wrapper.R will be used.
+#' @param	wrapperPath		file path to the wrapper. if none is given, findSVMWrapper needs to be called
+#' @param	softwarePath		path where to find the solver (software), if none is given, findSVMSoftware has to be called
 #' @param	verbose		be verbose?
 #'
 #' @note	 	first the given train and testBinaryPaths will be directly checked.
 #' if the binary does not exist there, the softwarePath will be added and rechecked
 #' and only if this does not work, the software will be searched via softwarePath.
 #' so one can override the search by specifiying train-/testBinaryPath.
-#'
+#' @note		If the wrapper searches for software, it will NOT execute it, existance is enough at this point.
+#' 
 #' @export	addSVMPackage
-addSVMPackage <- function (method = NA, filePath = NA, softwarePath = NA, 
-	trainBinaryPath = NA, testBinaryPath = NA, wrapperPath = NA, verbose = FALSE)
+addSVMPackage = function (method = NA, 
+	wrapperName = NA,
+	wrapperPath = NA,
+	softwarePath = NA, 
+	verbose = FALSE)
 {
-	# we always need a method
-	if (is.na(method) == TRUE) {
[…]

### 2016-01-04 R/checkModel.R

@@ -1,0 +1,48 @@
+#
+# SVMBridge 
+#		(C) 2015, by Aydin Demircioglu
+#
+#		writeLIBSVMModel.R
+# 
+# SVMBridge is free software: you can redistribute it and/or modify
+# it under the terms of the GNU Lesser General Public License as published
+# by the Free Software Foundation, either version 3 of the License, or
+# (at your option) any later version.
+#
+# SVMBridge is distributed in the hope that it will be useful,
+# but WITHOUT ANY WARRANTY; without even the implied warranty of
+# MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
+# GNU Lesser General Public License for more details.
+#
+# Please do not use this software to destroy or spy on people, environment or things.
+# All negative use is prohibited.
+#
+ 
+
+	
+#' Write LIBSVM model.
+#'
+#' Given a model in memory, write it out in LIBSVM model format, so that e.g. LIBSVM can predict on it directly.
+#' 
+#' @param	model		model object to write
+#' @param	modelFile		path where to write the model
+#' @param	verbose		be verbose?
+#'
+#' @note 		As this is a basic for all other model readers, we export it.
+#'
+#' @export	writeLIBSVMModel
+checkModel = function (model = NA, verbose = FALSE) {
+
+	if (verbose == TRUE)
+		cat ("Checking model for completeness.\n")
+	
+	checkmate::assertNumeric (model$gamma)
+	checkmate::assertMatrix (model$alpha)
+
+	if (verbose == TRUE)
+		cat ("    Done.\n")
+	
+	# close connection
+	close(modelFileHandle)
+}
+ 

### 2016-01-04 R/getSVMMethodsAsList.R

@@ -1,0 +1,50 @@
+#
+# SVMBridge 
+#		(C) 2015, by Aydin Demircioglu
+#
+#		getSVMMethodsAsList.R
+# 
+# SVMBridge is free software: you can redistribute it and/or modify
+# it under the terms of the GNU Lesser General Public License as published
+# by the Free Software Foundation, either version 3 of the License, or
+# (at your option) any later version.
+#
+# SVMBridge is distributed in the hope that it will be useful,
+# but WITHOUT ANY WARRANTY; without even the implied warranty of
+# MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
+# GNU Lesser General Public License for more details.
+#
+# Please do not use this software to destroy or spy on people, environment or things.
+# All negative use is prohibited.
+#
+
+
+
+#' getSVMMethodsAsList
+#' 
+#' Retrieve the names all known packages as a list of strings.
+#'
+#' @param	verbose		be verbose in output?
+#' 
+#' @return	A vector of all known SVM solvers as string.
+#'
+#' @export	outputAllSVMSoftwarePackages
+
+getSVMMethodsAsList = function (verbose = TRUE) {
+
+	knownPackages = c()
+	if (verbose == TRUE) {
+		cat("Currently known solver:\n")
+	}
+	
+	for (package in SVMBridgeEnv$packages) {
+		if (verbose == TRUE) {
+			cat ("    -", package, "\n")
+		}
+
+		knownPackages = c(knownPackages, package)
+	}
+	
+	return (knownPackages)
+}
+

### 2016-01-04 R/getSVMObject.R

@@ -1,0 +1,34 @@
+#
+# SVMBridge 
+#		(C) 2015, by Aydin Demircioglu
+#
+#		getSVMObject.R
+# 
+# SVMBridge is free software: you can redistribute it and/or modify
+# it under the terms of the GNU Lesser General Public License as published
+# by the Free Software Foundation, either version 3 of the License, or
+# (at your option) any later version.
+#
+# SVMBridge is distributed in the hope that it will be useful,
+# but WITHOUT ANY WARRANTY; without even the implied warranty of
+# MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
+# GNU Lesser General Public License for more details.
+#
+# Please do not use this software to destroy or spy on people, environment or things.
+# All negative use is prohibited.
+#
+
+
+
+#' getSVMObject
+#'
+#' Retrieve the object for a given method.
+#'
+#' @param	method		name of method
+#' @return	SVM object for the given method (or NULL if it does not exist)
+#'
+#' @export
+getSVMObject = function (method = method) {
+	return (SVMBridgeEnv$packages[[method]])
+}
+

### 2016-01-04 R/setSVMObject.R

@@ -1,0 +1,36 @@
+#
+# SVMBridge 
+#		(C) 2015, by Aydin Demircioglu
+#
+#		getSVMObject.R
+# 
+# SVMBridge is free software: you can redistribute it and/or modify
+# it under the terms of the GNU Lesser General Public License as published
+# by the Free Software Foundation, either version 3 of the License, or
+# (at your option) any later version.
+#
+# SVMBridge is distributed in the hope that it will be useful,
+# but WITHOUT ANY WARRANTY; without even the implied warranty of
+# MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
+# GNU Lesser General Public License for more details.
+#
+# Please do not use this software to destroy or spy on people, environment or things.
+# All negative use is prohibited.
+#
+
+
+
+#' getSVMObject
+#'
+#' Retrieve the object for a given method.
+#'
+#' @param	method		name of method
+#' @return	SVM object for the given method (or NULL if it does not exist)
+#'
+#' @export
+setSVMObject = function (method = method, object = NULL) {
+	if (checkmate::testNull (object) == TRUE) {
+		warn ("Setting a NULL object, i.e. object for method ", method, " will be erased.")
+	SVMBridgeEnv$packages[[method]] = object
+}
+

### 2016-01-04 R/writeLIBSVMModel.R

@@ -1,10 +1,8 @@
-#!/usr/bin/Rscript  --vanilla 
-
 #
 # SVMBridge 
 #		(C) 2015, by Aydin Demircioglu
 #
-#	.LASVM_walltime_wrapper.R
+#		writeLIBSVMModel.R
 # 
 # SVMBridge is free software: you can redistribute it and/or modify
 # it under the terms of the GNU Lesser General Public License as published
@@ -22,18 +20,27 @@
  
 
 	
-#' Write LIBSVM model
+#' Write LIBSVM model.
 #'
-#' As this is a basic for all other model readers, we export it.
+#' Given a model in memory, write it out in LIBSVM model format, so that e.g. LIBSVM can predict on it directly.
 #' 
 #' @param	model		model object to write
 #' @param	modelFile		path where to write the model
 #' @param	verbose		be verbose?
 #'
+#' @note 		As this is a basic for all other model readers, we export it.
+#'
 #' @export	writeLIBSVMModel
 writeLIBSVMModel = function (model = NA, modelFile = "./model", verbose = FALSE) {
+
+	checkmate::assertFlag (verbose)
+	checkmate::assertString (modelFile)
+
+	# check the model if everything is as we expect it
+	# here: alpha, label... gamma,..
+	
 	if (verbose == TRUE) {
-		BBmisc::messagef ("Writing SVM Model to %s", modelFile)
+		cat("Writing SVM Model to ", modelFile, "\n")
 	}
 	
 	# check of S3 object here.
@@ -44,9 +51,10 @@ writeLIBSVMModel = function (model = NA, modelFile = "./model", verbose = FALSE)
 
 	# open connection
 	if (verbose == TRUE)
-		cat ("  Writing Header\n")
-	modelFileHandle <- file(modelFile, open = "w+")
+		cat ("    Writing Header.\n")
+	modelFileHandle = file(modelFile, open = "w+")
 
+	# TODO: allow for other kernels probably?
 	writeLines(paste ("svm_type c_svc", sep = ""), modelFileHandle )
 	writeLines(paste ("kernel_type", "rbf", sep = " "), modelFileHandle )
 	
@@ -71,11 +79,13 @@ writeLIBSVMModel = function (model = NA, modelFile = "./model", verbose = FALSE)
 	#svmatrix = dumpSparseFormat(model$alpha, model$X)
 	#writeLines(svmatrix, modelFileHandle, sep = "" )
[…]

### 2016-01-04 tests/other/usecases_addSVMPackage.R

@@ -1,0 +1,67 @@
+
+have another function like 
+changeSVMPackage ( , key=value)?
+
+# expect to simply fail
+addSVMPackage ()
+
+# FIRST BATCH
+# will just create an object with empty wrapper and empty software.
+# both can be added later via findSVM....R
+
+# add a new package
+addSVMPackage (method = "LIBSVM")
+
+
+# SECOND BATCH
+# empty package + wrapper gets specified
+
+# + will locate the wrapper at ./wrapperPath/LIBSVM_wrapper.R
+addSVMPackage (method = "LIBSVM", wrapperPath = "./wrapperPath")
+
+# + will locate the wrapper at ./wrapperPath/specified.R
+addSVMPackage (method = "LIBSVM", wrapperPath = "./wrapperPath/specified.R")
+
+# + will search the wrapper below './wrapperPath'
+addSVMPackage (method = "LIBSVM", wrapperSearchPath = "./wrapperPath")
+
+# expect to fail, cannot work with both? OR: will try to locate first, if not, will search for it.
+addSVMPackage (method = "LIBSVM", wrapperPath = "./wrapperPath", wrapperSearchPath = "./wrapperPath")
+
+# override default wrapper name, so you can use findWrapper and find a differently named wrapper.
+addSVMPackage (method = "LIBSVM", wrapperName = "my_wrapper.R")
+
+# expect to fail, cannot work with both? OR: will try to locate first, if not, will search for it.
+addSVMPackage (method = "LIBSVM", wrapperName = "my_wrapper.R", wrapperSearchPath = "./wrapperPath")
+addSVMPackage (method = "LIBSVM", wrapperName = "my_wrapper.R", wrapperPath = "./wrapperPath")
+
+
+
+
+#THIRD BATCH
+# empty package + wrapper gets specified + software gets specified
+
+# + will locate the wrapper at ./wrapperPath/LIBSVM_wrapper.R
+addSVMPackage (method = "LIBSVM", wrapperPath = "./wrapperPath")
+
+# + will locate the wrapper at ./wrapperPath/specified.R
+addSVMPackage (method = "LIBSVM", wrapperPath = "./wrapperPath/specified.R")
+
+
+# NOTE: there is no case where we have no wrapper, but software!
+
+# FOURTH BATCH: mutants
+
+# this will try to find the wrapper at ./softwarePath/LIBSVM_wrapper.R
+addSVMPackage (method = "LIBSVM", softwarePath = "./softwarePath")
+
[…]

### 2016-01-05 doc/tutorial/detailed.Rmd
add some more unit tests

@@
 
+## Calling findSoftware (or other methods) result in 'no applicable method for 'findSoftware' applied to an object of class...'
+
+This means that the wrapper for your SVM was not found, therefore no object was created, and the findSoftware could not be executed, because it is simply not there.
+Double check that adding the SVM package works for you, e.g.
+by enabling verbose mode.
+'
+
 # Speed

### 2016-01-05 R/SVMWrapper.R

@@ -33,15 +33,22 @@
 #' ...
 
 createSVMWrapper = function(method, ...) {
-	# check that trainingBinary exists, if given
-	# ...    
-	
-	# register the newly created object in our environment
-	
+
+	# remove first all old stuff lying around
+	# TODO: one must be able to do this better, but how? can anyone fix this :/
+	suppressWarnings( rm (envir = globalenv(), list = c(paste0("createTestArguments", ".", method), 
+		paste0("createTrainingArguments", ".", method), 
+		paste0("detectModel", ".", method), 
+		paste0("extractTestInfo", ".", method), 
+		paste0("extractTrainingInfo", ".", method), 
+		paste0("findSoftware", ".", method), 
+		paste0("print", ".", method), 
+		paste0("readModel", ".", method), 
+		paste0("readPredictions", ".", method), 
+		paste0("writeModel", ".", method) )))
+
 	# create S3 Object
-	svmPackage = BBmisc::makeS3Obj(c(method, "SVMWrapper"),
-		method = method
-	)
+	svmPackage = BBmisc::makeS3Obj(c(method, "SVMWrapper"), method = method)
 
 	# TODO: replace existing object in package list
 	

### 2016-01-05 R/findSVMSoftware.R

@@ -38,13 +38,33 @@ findSVMSoftware <- function (method = NA, searchPath = NA, verbose = FALSE) {
 	checkmate::assertFlag (verbose)
 
 	if (verbose == TRUE) {
-		cat("-Finding software for %s", method)
+		cat("Finding software for ", method, "\n")
 	}
 	
 	# replace tilde characters by expanding them
 	searchPath = expandTilde(path = searchPath, verbose = verbose)
 	
-	# call the find software method of the solver
-	SVMBridgeEnv$packages[[method]] = findSoftware (SVMBridgeEnv$packages[[method]], searchPath = searchPath, verbose = verbose)
+	# iterate over all search paths until we hit the target
+	dirList = list.dirs(searchPath, recursive = TRUE)
+    
+    SVMObject = getSVMObject (method)
+    for (dir in dirList) {
+		# call the find software method of the solver
+		if (verbose == TRUE) {
+			cat ("    ", dir, "\n")
+		}
+		SVMObject = findSoftware (SVMObject, searchPath = dir, verbose = verbose)
+		if (is.null (SVMObject$trainBinaryPath) == FALSE) {
+			if (verbose == TRUE) {
+				cat ("    Found binaries in", dir, "\n")
+				cat ("       ", SVMObject$trainBinaryPath, "\n")
+				cat ("       ", SVMObject$testBinaryPath, "\n")
+			}
+			break
+		}
+	}
+	
+	setSVMObject (method, SVMObject)
+	return (is.null (SVMObject$trainBinaryPath) == FALSE)
 }
 

### 2016-01-05 tests/testthat/test_findSVMSoftware.R

@@ -1,0 +1,53 @@
+context("Find SVM Software")
+
+
+test_that("finding works in general", {
+	addSVMPackage ("oneSVM", wrapperPath = "../dummy/oneSVM")
+	findSVMSoftware ("oneSVM", searchPath = "../dummy/oneSVM")
+
+	svmObject = getSVMObject ("oneSVM")
+	expect_equal (svmObject$method, "oneSVM")
+	expect_equal (svmObject$wrapperName, "oneSVM_wrapper.R")
+	expect_equal (class(svmObject)[2], "SVMWrapper")
+	expect_equal (svmObject$trainBinaryPath, "../dummy/oneSVM/oneSVM-learn")
+	expect_equal (svmObject$testBinaryPath, "../dummy/oneSVM/oneSVM-predict")
+}	)
+
+
+
+test_that("finding non existent software returns error code", {
+	addSVMPackage ("oneSVM", wrapperPath = "../dummy/twoSVM")
+	retValue = findSVMSoftware ("oneSVM", searchPath = "../dummy/twoSVM")
+
+	svmObject = getSVMObject ("oneSVM")
+	expect_equal (retValue, FALSE)
+	expect_equal (svmObject$method, "oneSVM")
+	expect_equal (svmObject$wrapperName, "oneSVM_wrapper.R")
+	expect_equal (class(svmObject)[2], "SVMWrapper")
+	expect_null (svmObject$trainBinaryPath)
+	expect_null (svmObject$testBinaryPath)
+}	)
+	
+
+
+test_that("finding software without wrapper does not work", {
+	addSVMPackage ("oneSVM")
+	expect_error (findSVMSoftware ("oneSVM", searchPath = "../dummy/fourSVM"), "no applicable method")
+}	)
+
+
+	
+test_that("finding software in a sub-sub-directory works", {
+	addSVMPackage ("oneSVM", wrapperPath = "../dummy/fourSVM")
+	retValue = findSVMSoftware ("oneSVM", searchPath = "../dummy/fourSVM")
+
+	svmObject = getSVMObject ("oneSVM")
+	expect_equal (retValue, TRUE)
+	expect_equal (svmObject$method, "oneSVM")
+	expect_equal (svmObject$wrapperName, "oneSVM_wrapper.R")
+	expect_equal (class(svmObject)[2], "SVMWrapper")
[…]

### 2016-01-05 tests/testthat/test_findSVMWrapper.R

@@ -1,0 +1,11 @@
+context("Find SVM Wrapper")
+
+
+test_that("finding non existent wrapper returns error code", {
+
+}	)
+	
+
+test_that("finding wrapper in a sub-sub-directory works", {
+	
+} )

### 2016-01-05 wrapper/LASVM_wrapper.R

@@ -192,8 +192,8 @@ detectModel.LASVM = function (x, modelFile = NULL, verbose = FALSE) {
 	
  
 	print.LASVM_walltime = function(x) {
-		BBmisc::messagef("--- Object: %s", x$method)
-		BBmisc::messagef("       Training Binary at %s", x$trainBinaryPath)
-		BBmisc::messagef("       Test Binary at %s", x$testBinaryPath)
+		cat("--- Object: ", x$method)
+		cat("       Training Binary at ", x$trainBinaryPath)
+		cat("       Test Binary at ", x$testBinaryPath)
 	}
 	

### 2016-01-06 R/getSVMMethodsAsList.R
snapshot

@@ -28,7 +28,7 @@
 #' 
 #' @return	A vector of all known SVM solvers as string.
 #'
-#' @export	outputAllSVMSoftwarePackages
+#' @export
 
 getSVMMethodsAsList = function (verbose = TRUE) {
 

### 2016-01-06 R/testSVM.R

@@ -1,4 +1,3 @@
-#!/usr/bin/Rscript  --vanilla 
 #
 # SVMBridge 
 #		(C) 2015, by Aydin Demircioglu
@@ -20,11 +19,13 @@
 #
  
 
-#' testSVM
+#' test a trained SVM.
+#'
+#' This is the main routine for testing/prediction an SVM.
 #'
 #' @param	testDataFile		file to read training data from 
 #' @param	testDataX		matrix to read training data from 
-#' @param	testDatay		matrix to read training label from 
+#' @param	testDataY		matrix to read training label from 
 #'
 #' @param	model		list containing a trained SVM model
 #' @param	modelFile       path to a model file
@@ -35,16 +36,12 @@
 #' @param	extraParameter		extra parameters for solver
 #' @param	verbose		be verbose?
 #'
-#' @note		exclusive parameters, i.e. you cannot specify both:
+#' @note		testDataFile and testDataX,testDataY are exclusive parameters, i.e. you cannot specify both.
+#' @note		Make sure the modelFile is a character string and not a factor! Else you might obtain strange modelfile arguments.
 #'
 #' @return		SVM Object
-#' @export	testSVM
-# #examples
-# 	testSVM(method = 'LIBSVM', testDataFile = './data/australian.test') 
 #'
-#
-# NOTE: make sure the modelFile is a character string and not a factor! else you might strange modelfile arguments.
-# If no
+#' @export
 
 testSVM = function(
 	method = NULL,
@@ -52,7 +49,7 @@ testSVM = function(
 	# data
 	testDataFile = NULL,
 	testDataX = NULL, 
-	testDatay = NULL, 
+	testDataY = NULL, 
 
 	# model 
 	model = NULL,
@@ -128,7 +125,7 @@ testSVM = function(
 		testDataFile = tempfile()
 		if (verbose == TRUE)
 			BBmisc::messagef("  Writing given test data as %s", testDataFile)
-		e1071::write.matrix.csr(testDataX, testDataFile, testDatay)
+		e1071::write.matrix.csr(testDataX, testDataFile, testDataY)
 	} 
 
 	if (verbose == TRUE) 
@@ -183,5 +180,4 @@ testSVM = function(
 	
 	return (results)   
 }
-
 

### 2016-01-06 R/trainSVM.R

@@ -1,4 +1,3 @@
-#!/usr/bin/Rscript  --vanilla 
 #
 # SVMBridge 
 #		(C) 2015, by Aydin Demircioglu
@@ -20,13 +19,15 @@
 #
  
 
-#' trainSVM
+#' Train an SVM.
+#'
+#' This is the main routine that trains an SVM.
 #'
 #' @param	method		name of the SVM method/solver
 #'
 #' @param	trainDataFile		file to read training data from 
 #' @param	trainDataX		matrix to read training data from 
-#' @param	trainDatay		matrix to read training label from 
+#' @param	trainDataY		matrix to read training label from 
 #'
 #' @param	trainBinaryPath		full path to the training binary to call
 #'
@@ -42,10 +43,8 @@
 #' @note		exclusive parameters, i.e. you cannot specify both:
 #'
 #' @return		SVM Object
-#' @export		trainSVM
-# #examples
-# #	trainSVM(model = 'LIBSVM', trainDataFile = './data/australian.train') 
 #'
+#' @export
 
 trainSVM = function(
 	# method
@@ -54,7 +53,7 @@ trainSVM = function(
 	# data
 	trainDataFile = NULL,
 	trainDataX = NULL, 
-	trainDatay = NULL, 
+	trainDataY = NULL, 
 
 	# full path to executable
 	trainBinaryPath = NULL,
@@ -130,7 +129,7 @@ trainSVM = function(
 		trainDataFile = tempfile()
 		if (verbose == TRUE)
 			BBmisc::messagef("  Writing given data as %s", trainDataFile)
-		e1071::write.matrix.csr(trainDataX, trainDataFile, trainDatay)
+		e1071::write.matrix.csr(trainDataX, trainDataFile, trainDataY)
 	} 
 
 	if (verbose == TRUE) 

### 2016-01-06 inst/wrapper/LASVM_wrapper.R

@@ -146,44 +146,24 @@ detectModel.LASVM = function (x, modelFile = NULL, verbose = FALSE) {
 
 	
 
-	findSoftware.LASVM = function (x, searchPath = "./", verbose = FALSE) {
+	findSoftware.LASVM = function (x, searchPath = "./", execute = FALSE, verbose = FALSE) {
 
 		if (verbose == TRUE) {
-			BBmisc::messagef("    LASVM Object: Executing search for software for %s", x$method)
+			cat ("    LASVM Object: Executing search for software for ", x$method)
 		}
-		
-		trainBinaryPattern = "^la_svm$"
-		trainBinaryOutputPattern = 'la_svm .options. training_set_file .model_file.'
-		binaryPath = findBinary (searchPath, trainBinaryPattern, trainBinaryOutputPattern, verbose = verbose)
 
-		# TODO: check for empty path+handling
-		
-		if (verbose == TRUE) {
-			BBmisc::messagef("--> Found train binary at %s", binaryPath) 
-		}
-		x$trainBinaryPath = binaryPath
-
-
-		testBinaryPattern = "^la_test$"
-		testBinaryOutputPattern = 'la_test .options. test_set_file model_file output_file'
-
-		binaryPath = findBinary (searchPath, testBinaryPattern, testBinaryOutputPattern, verbose = verbose)
-		
-		# TODO: check for empty path+handling
-
-		if (verbose == TRUE) {
-			BBmisc::messagef("--> Found test binary at %s", binaryPath) 
-		}
-		x$testBinaryPath = binaryPath
+		# can do now OS specific stuff here
+		x$trainBinaryPath = findBinaryInDirectory ("la_svm", dir = searchPath, patterns = list ('la_svm .options. training_set_file .model_file.'))
+		x$testBinaryPath = findBinaryInDirectory ("la_test", dir = searchPath, patterns = list ('la_test .options. test_set_file model_file output_file'))
 
 		return(x)
 	}
-
- 
+	
+	
  
 	print.LASVM_walltime = function(x) {
-		BBmisc::messagef("--- Object: %s", x$method)
-		BBmisc::messagef("       Training Binary at %s", x$trainBinaryPath)
-		BBmisc::messagef("       Test Binary at %s", x$testBinaryPath)
+		cat("--- Object: ", x$method)
+		cat("       Training Binary at ", x$trainBinaryPath)
+		cat("       Test Binary at ", x$testBinaryPath)
 	}
 	

### 2016-01-06 tests/testthat/test_expandTilde.R

@@ -1,0 +1,20 @@
+context("Expand Tilde")
+
+
+test_that("something without a tilde does not change", {
+
+	path = "/this/is\\a\\strange/thing#we/do\\not&^$#@!anything}{L:\"about\n\n"
+	expPath = expandTilde (path = path)
+	expect_equal (path, expPath)
+}	)
+	
+
+test_that("tilde gets expanded", {
+	# problem is a little bit: we do not know the home directory.
+	# so we only test here if the tilde is finally gone or not
+	path = "~/testdir"
+	expPath = expandTilde (path = path)
+	expect_equal (grepl("~", expPath), FALSE)
+}	)
+	
+

### 2016-01-14 doc/tutorial/introduction.Rmd
small up

@@
 **Warning:** The findSVMSoftware functionality is made for convenience, not for performance.
-As all given directories has to be searched, this can slow down all I/O operations.
-Prevent running this function on a high performance cluster often, as might very well slowdown the whole cluster!
+As all given (sub)directories have to be searched, this can slow down all I/O operations.
+Restrain from running this function on a high performance cluster, as it might very well slowdown the whole cluster!

### 2016-01-14 doc/tutorial/wrappers.Rmd
small up

# Wrappers

**SVMBridge** comes with a couple of wrappers for different SVM packages. Here we list these and provide a few details on how to use them.

**Warning:** Currently we do not seek completeness. The **SVMBridge** is a classification-only package, so many options e.g. regarding regression etc are not supported at all. Some other options, although they can be used with classifications, might be missing too. In case you need something, you will need to adapt an existing classifier or simply write your own from scratch. If you do so, please share your results with us, so we can include these in future versions of the **SVMBridge**.

**Note:** For convenience, all the wrappers are installed along the package and can be accessed by specifying the correct path. See the introduction for examples on how to do this.

## LIBSVM

LIBSVM is practically the de-facto standard SVM solver.
The wrapper supports for now classification and RBF kernel only.

LIBSVM can be found at https://www.csie.ntu.edu.tw/~cjlin/libsvm/.

## LASVM

LASVM is very similar to LIBSVM wrapper, as both packages use the same parameters and the same model. LASVM has many more options than LIBSVM, as for our purpose we did not need them, for now these are not supported.

LASVM can be found at http://leon.bottou.org/projects/lasvm/.

## LLSVM

LLSVM is part of the BudgetedSVM package, see http://www.dabi.temple.edu/budgetedsvm/.

**Note**: The models saved by LLSVM are quite non-standard, as the transformation matrices
have to be saved instead of a list of support vectors. The wrapper includes a simple
reader, which is written in pure R. Therefore, the performance of this wrapper is rather
reduced. In case you need faster model readers/writers, you need to handle it by
yourself.

## Budgeted SGD

BSGD is part of the BudgetedSVM package, see http://www.dabi.temple.edu/budgetedsvm/.

**Note**: The models saved by LLSVM are quite non-standard, as the transformation matrices
have to be saved instead of a list of support vectors. The wrapper includes a simple
reader, which is written in pure R. Therefore, the performance of this wrapper is rather
reduced. In case you need faster model readers/writers, you need to handle it by
yourself.

## SVMperf

SVMperf has a great deal of options, so instead of providing all of these as parameter, we decided instead to give the user a choice over several 'configurations'.

SVMperf can be found at https://www.cs.cornell.edu/people/tj/svm_light/svm_perf.html.

**Note**: Please do not mix up SVMperf with SVMlight or SVMRank.

## CVM

**Note**: The source code for CVM cannot be found in the internet anymore.
Therefore we provide a copy of it in our github repository.

## BVM

Though BVM is practially the same software as CVM, sharing models and parameters and executables with CVM, the philosophy of **SVMBridge** is to have a separate wrapper for each method. Thus, though BVM has its own wrapper, all the parameters etc are the same as CVM, so please read the CVM section.

**Note**: The source code for BVM cannot be found in the internet anymore.
Therefore we provide a copy of it in our github repository.

### 2016-01-14 src/OptimizationValues.cpp

@@ -398,7 +398,7 @@ Rcpp::List computeOptimizationValues (NumericMatrix X, NumericVector Y, double C
 		if (SV.ncol() > X.ncol()) {
 			NumericMatrix tmpX (X.nrow(), SV.ncol());
 			// better copy possible?
-			for (size_t i = 0; i < X.nrow(); i++) {
+			for (unsigned long i = 0; i < X.nrow(); i++) {
 				NumericMatrix::Row zzrow = X( i, _);
 				NumericMatrix::Row yyrow = tmpX( i, _);
 				std::copy (zzrow.begin(), zzrow.end(), yyrow.begin());
@@ -406,7 +406,7 @@ Rcpp::List computeOptimizationValues (NumericMatrix X, NumericVector Y, double C
 			X = tmpX;
 		} else {
 			NumericMatrix tmpSV (SV.nrow(), X.ncol());
-			for (size_t i = 0; i < SV.nrow(); i++) {
+			for (unsigned long i = 0; i < SV.nrow(); i++) {
 				NumericMatrix::Row zzrow = SV( i, _);
 				NumericMatrix::Row yyrow = tmpSV( i, _);
 				std::copy (zzrow.begin(), zzrow.end(), yyrow.begin());

### 2016-01-14 src/RcppExports.cpp

@@ -26,13 +26,13 @@ BEGIN_RCPP
 END_RCPP
 }
 // readSparseData
-List readSparseData(std::string filename, size_t skipBytes, bool verbose, bool zeroBased);
+List readSparseData(std::string filename, unsigned long skipBytes, bool verbose, bool zeroBased);
 RcppExport SEXP SVMBridge_readSparseData(SEXP filenameSEXP, SEXP skipBytesSEXP, SEXP verboseSEXP, SEXP zeroBasedSEXP) {
 BEGIN_RCPP
     Rcpp::RObject __result;
     Rcpp::RNGScope __rngScope;
     Rcpp::traits::input_parameter< std::string >::type filename(filenameSEXP);
-    Rcpp::traits::input_parameter< size_t >::type skipBytes(skipBytesSEXP);
+    Rcpp::traits::input_parameter< unsigned long >::type skipBytes(skipBytesSEXP);
     Rcpp::traits::input_parameter< bool >::type verbose(verboseSEXP);
     Rcpp::traits::input_parameter< bool >::type zeroBased(zeroBasedSEXP);
     __result = Rcpp::wrap(readSparseData(filename, skipBytes, verbose, zeroBased));
@@ -40,7 +40,7 @@ BEGIN_RCPP
 END_RCPP
 }
 // writeSparseData
-List writeSparseData(std::string filename, NumericMatrix X, NumericMatrix Y, size_t skipBytes, bool verbose, bool zeroBased);
+List writeSparseData(std::string filename, NumericMatrix X, NumericMatrix Y, unsigned long skipBytes, bool verbose, bool zeroBased);
 RcppExport SEXP SVMBridge_writeSparseData(SEXP filenameSEXP, SEXP XSEXP, SEXP YSEXP, SEXP skipBytesSEXP, SEXP verboseSEXP, SEXP zeroBasedSEXP) {
 BEGIN_RCPP
     Rcpp::RObject __result;
@@ -48,7 +48,7 @@ BEGIN_RCPP
     Rcpp::traits::input_parameter< std::string >::type filename(filenameSEXP);
     Rcpp::traits::input_parameter< NumericMatrix >::type X(XSEXP);
     Rcpp::traits::input_parameter< NumericMatrix >::type Y(YSEXP);
-    Rcpp::traits::input_parameter< size_t >::type skipBytes(skipBytesSEXP);
[…]

### 2016-01-14 src/SparseData.cpp

@@ -102,7 +102,7 @@ static char* readline(FILE *input)
 //'	}
 //' @export
 // [[Rcpp::export]] 
-List readSparseData (std::string filename, size_t skipBytes = 0, bool verbose = false, bool zeroBased = false) {
+List readSparseData (std::string filename, unsigned long  skipBytes = 0, bool verbose = false, bool zeroBased = false) {
 	
 	std::setprecision(16);
   
@@ -444,7 +444,7 @@ List readSparseData (std::string filename, size_t skipBytes = 0, bool verbose = 
 //'		writeSparseData ("./australian.data", X, Y)
 //' @export
 // [[Rcpp::export]] 
-List writeSparseData (std::string filename, NumericMatrix X, NumericMatrix Y, size_t skipBytes = 0, bool verbose = false, bool zeroBased = false) {
+List writeSparseData (std::string filename, NumericMatrix X, NumericMatrix Y, unsigned long  skipBytes = 0, bool verbose = false, bool zeroBased = false) {
 	
 	try
 	{

### 2016-01-14 doc/tutorial/windowsHowTo.Rmd
cleanup, add LLSVM writer

@@
-***SVMBridge Windows Tutorial***
+# SVMBridge Windows Quick Tutorial
 
-1) Download R 3.2.x or later for Windows (32/64 bit) (https://cran.r-project.org/bin/windows/base/)
+- Download latest R for Windows (32/64 bit) (https://cran.r-project.org/bin/windows/base/)
+- Download Rtools for R 3.2.x and later (https://cran.r-project.org/bin/windows/Rtools/)
+- Install R and use the 32 bit option only
+- Install Rtools and use the 32 bit option only
+- Install SVMBridge by starting R and typing ```install.packages ("SVMBridge")```
+- If SVMBridge was successfully installed, you can now execute the OS-specific tests. For this
+you need testthat. To install testthat, ```install.packages ("testthat")```
+- Then execute the tests by .....
+- Now if the tests were succesful, you end up with a bunch of pre-compiled SVM software.
+Copy these over to a directory you like, e.g.
+C:\Programs\SVMSolver\.
+Alternatively, especially for those SVM solvers for which the **SVMBridge** has no support, download and compile the SVM software by yourself and also put it into a directory of your choice.
 
-2) Download Rtools for R 3.2.x and later (https://cran.r-project.org/bin/windows/Rtools/)
+You're ready to test the SVMBridge.
 
-3) Download Git for Windows (https://git-scm.com/downloads)
-
-3) Install R and use the 32 bit option only
-
-4) Install Rtools and use the 32 bit option only
-
-5) Install Git
-
-6) Start up the Git Bash and navigate into a directory in which you want to download SVMbridge to. For best compatibility,
-   install SVMBridge into your home directory. Some tests may fail if you put it elsewhere.
-
-7) Download the SVMBridge repository by using git clone in the git console 
-   (git clone https://github.com/aydindemircioglu/SVMBridge.git)
-
-8) Delete the makefile.win file, which is located in the src directory in SVMBridge.
+- ...
 
@@
    (For the example with LIBSVM you don't need to change the names of the binaries)
-
-9) Start up R Gui and install following packages by typing "install.packages("package-name")" into the R Gui: BBmisc, stringr, 
-    plyr, Rcpp, microbenchmark, R.utils, e1071, SparseM, testthat, knitr, devtools
 

### 2016-01-14 R/sparseFormat.R

@@ -127,11 +127,11 @@ readSparseFormat <- function (con)
 #'
 #' @return	sparseLine		string with row data 
 #'
-dumpSparseLine = function (row) {
+dumpSparseLine = function (row, dense = FALSE) {
 	sparseLine = ''
 
 	for (x in seq(1, length(row))) {
-		if (row[x] != 0) {
+		if ((row[x] != 0) | dense == TRUE) {
 			sparseLine = paste(sparseLine, paste(x, row[x], sep =":"), sep = " ")
 		}
 	}
@@ -195,7 +195,7 @@ readSparseDataFromConnection = function (con, verbose = FALSE, zeroBased = FALSE
 writeSparseDataToConnection = function (con, X, Y, verbose = FALSE, zeroBased = FALSE) {
 	currentPosition = seek(con)
 	filename = summary(con)$description
-	writeSparseData(filename, X, Y, skipBytes = currentPosition, verbose = verbose, zeroBased = zeroBased)
+	writeSparseData(filename, X, Y, skipBytes = currentPosition, verbose = verbose, zeroBased = zeroBased, dense = FALSE)
 }
 
 

### 2016-01-14 inst/tests/speedtest.R

@@ -1,25 +1,0 @@
-	
-	
-# 08/15 svm
-#	s = ssvm (x, y, gamma = 0.585568587668891, epsilon = 1e-6)
-
-# 	writeSparseData (x,y, filename = "./tmp/testOutput.sparse")
-# 	e = readSparseData (filename = "./tmp/testOutput.sparse", normalizeLabels = FALSE, verbose = TRUE)
-# 	
-# 	messagef("Testing Reading, e1071 vs swarmsvm, 5 times")
-# 	t = microbenchmark( read.matrix.csr("../lab/data/poker/poker.combined.scaled"), times = 5)
-# 	messagef("e1071:")	
-# 	print (t)
-# 	t = microbenchmark( readSparseData (filename = "../lab/data/poker/poker.combined.scaled"), times = 5)
-# 	messagef("swarmsvm:")	
-# 	print(t)
-# 
-# 	messagef("Testing Writing, e1071 vs swarmsv, 3 times")
-# 	t = microbenchmark( write.matrix.csr(x, "tmp/poker.combined.scaled", y), times = 30)
-# 	messagef("e1071:")	
-# 	print (t)
-# 	t = microbenchmark( writeSparseData (x, y, filename = "tmp/poker.combined.scaled"), times = 30)
-# 	messagef("swarmsvm:")	
-# 	print(t)
-
-

### 2016-01-14 inst/tests/test_BSGD_wrapper.R

@@ -1,28 +1,0 @@
-context("BSGD")
-
-test_that("BSGD wrapper works", {
-	solver = "BSGD"
-	addSVMPackage (method = solver, verbose = FALSE)
-	findSVMSoftware (solver, searchPath = "~/shark/svm_large_data/software/", verbose = TRUE)
-	verbose = TRUE
-	C = 0.71
-	gamma = 0.41
-	dataset = ("../data/mnist.data")
-	
-	trainObj =  trainSVM(
-		method = solver,
-		trainDataFile = dataset,
-		cost = C, 
-		gamma = gamma, 
-		readModelFile = FALSE,
-		modelFile = "~../data/mnist.BSGD.model",
-		verbose = verbose
-	)  
-	
-	testObj =  testSVM(
-			method = solver, 
-			testDataFile = dataset,
-			modelFile = "../data/mnist.BSGD.model",
-			verbose = verbose
-		) 
-})

### 2016-01-14 inst/tests/test_expandTilde.R

@@ -1,28 +1,0 @@
-context("expandTilde")
-
-test_that(" tilde characters are expanded correctly.", {
-	solver = "LIBSVM"
-	verbose = TRUE
-	cost = runif(1)
-	gamma = runif(1)
-	addSVMPackage (method = solver, verbose = FALSE)
-	findSVMSoftware (solver, searchPath = "~/shark/svm_large_data/software/", verbose = FALSE)
-	trainFile = ("~/SVMBridge/tests/data/sparse.data")
-	obj1 =  trainSVM(
-			method = solver,
-			trainDataFile = trainFile, 
-			cost = cost, 
-			gamma = gamma, 
-			readModelFile = TRUE,
-			verbose = verbose
-		)  
-	obj2 =  testSVM(
-			method = solver,
-			testDataFile = trainFile,
-			model = obj1$model,
-			verbose = verbose
-		) 
-	obj3 = readSparseData(filename = trainFile,
-			      verbose = verbose
-		)
-})

### 2016-01-14 wrapper/BSGD_wrapper.R

@@ -324,7 +324,7 @@ writeModel.BSGD = function (x, model = NA, modelFile = "./model", verbose = FALS
 	degree = 1
 	coef = 1
 
-	print (as.character (paste (model$label, sep = " ", collapse = " ")))
+#	print (as.character (paste (model$label, sep = " ", collapse = " ")))
 
     # open connection
     modelFileHandle <- file(modelFile, open = "w+")

### 2016-01-14 wrapper/LLSVM_wrapper.R

@@ -243,12 +243,66 @@ readModel.LLSVM = function (x, modelFile = "./model", verbose = FALSE)
 
 
 # dummy for now
-writeModel.LLSVM = function (x, model = NA, modelFile = "./model", verbose = FALSE) {
-		ret = writeLIBSVMModel (model = model, modelFile = modelFile, verbose = verbose)
-		return (ret)
+writeModel.LLSVM = function (x, model = NA, modelFile = "./model", verbose = FALSE)
+{
+	#FIXME: does this change globally?
+	options(scipen=999)
+
+	# check for weightvector, coefficients and SV
+	
+	# check for size 
+	if ((nrow(model$SV) != nrow(model$alpha)) || (nrow(model$alpha) != length(model$w))) {
+		warning ("Cannot write model, as model is not consistent!")
+		return (FALSE)
 	}
 
+	# get things
+	dim = ncol (model$SV)
+	nClasses = 2
+	nWeights = nrow (model$SV)
+	gamma = model$gamma * 2
+	degree = 1
+	coef = 1
 
+	
+	# write header
+	modelFileHandle <- file(modelFile, open = "w+")
+	writeLines(paste ("ALGORITHM: 3", sep = ""), modelFileHandle )
+	writeLines(paste ("DIMENSION:", dim, sep = " "), modelFileHandle )
+	writeLines(paste ("NUMBER_OF_CLASSES:", nClasses, sep = " "), modelFileHandle )
+	writeLines(paste ("LABELS:", as.character (paste (model$label, sep = " ", collapse = " ")), sep = " "), modelFileHandle )
+	writeLines(paste ("NUMBER_OF_WEIGHTS:", nWeights, sep = " "), modelFileHandle )
+	writeLines(paste ("BIAS_TERM: ", model$bias, sep = " "), modelFileHandle )
+	writeLines(paste ("KERNEL_FUNCTION: 0", sep = " "), modelFileHandle )
+	writeLines(paste ("KERNEL_GAMMA_PARAM:", gamma, sep = " "), modelFileHandle )
+	writeLines(paste ("KERNEL_DEGREE_PARAM:", degree, sep = " "), modelFileHandle )
+	writeLines(paste ("KERNEL_COEF_PARAM:", coef, sep = " "), modelFileHandle )
+	writeLines(paste ("MODEL:", sep = ""), modelFileHandle )
+	
+	
[…]

### 2016-02-03 doc/tutorial/detailed.Rmd
snapshot
fix bvm wrapper

@@
 by enabling verbose mode.
+
+## Something like basename(trainBinaryPath) : a character vector argument expected happened.
+
+(FIXME actually). The software was not found, so there is no trainBinaryPath in the SVM object.
 '
+## Reading my SVM model (or data file) yields a 0x0 matrix?
+
+Make sure the model has the correct line endings.
+On windows you need 0x0d0a, else, if the model/data file has e.g.
+linux endings (0x0a), these will not be correctly read, that is the whole file will be read as one line, confusing the reader.
 

### 2016-02-03 tests/wrapper/BVM_wrapper.R

@@ -134,6 +134,8 @@ findSoftware.BVM = function (x, searchPath = "./", execute = FALSE, verbose = FA
 	# can do now OS specific stuff here
 	x$trainBinaryPath = findBinaryInDirectory (trainBinaryPattern, dir = searchPath, patterns = list ('bvm-train .options. training_set_file .model_file.'))
 	x$testBinaryPath = findBinaryInDirectory (testBinaryPattern , dir = searchPath, patterns = list ('bvm-predict .options. test_file model_file output_file'))
+	
+	return (x)
 }
 
 

### 2016-02-04 compile.R
some package reordering

@@ -7,17 +7,29 @@ library(methods)
 library(devtools)
 library(roxygen2)
 library(Rcpp)
+
+cat ("## Super cleanup\n")
+system2(command="rm", args="-rf inst")
+system2(command="mkdir", args="inst")
+system2(command="mkdir", args="inst/doc")
+
+cat ("## Compiling Rcpp attributes\n")
 compileAttributes ()
+cat ("## Roxygenizing\n")
+roxygenize(clean = TRUE)
+cat ("## Documenting\n")
+document()
+cat ("## Building Vignettes\n")
+build_vignettes()
 
-roxygenize(".")
-load_all (".")
-build_vignettes(".")
-document(".")
-
-devtools::test()
+cat ("## Checking package\n")
 devtools::check()
+cat ("## Testing package\n")
+devtools::test()
 
 
 # work around
+cat ("## Building package\n")
 system2(command="R", args="CMD build .")
+cat ("## Checking package for CRAN\n")
 system2(command="R", args="CMD check --as-cran ./SVMBridge_0.1.tar.gz")

### 2016-02-04 inst/examples/SparseData-Example.R

@@ -1,86 +1,0 @@
-#!/usr/bin/Rscript  --vanilla 
-
-#
-# SVMBridge 
-#
-#		(C) 2015, by Aydin Demircioglu
-# 
-# SVMBridge is free software: you can redistribute it and/or modify
-# it under the terms of the GNU Lesser General Public License as published
-# by the Free Software Foundation, either version 3 of the License, or
-# (at your option) any later version.
-#
-# SVMBridge is distributed in the hope that it will be useful,
-# but WITHOUT ANY WARRANTY; without even the implied warranty of
-# MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
-# GNU Lesser General Public License for more details.
-#
-# Please do not use this software to destroy or spy on people, environment or things.
-# All negative use is prohibited.
-#
-
-	library(methods)
-	library(devtools)
-	load_all ("../..")
-
-
-#In the following we are going to do a few test runs with SparseData.cpp to simulate certain scenarios
-
-# --- perform a simple reading task
-cat("Test #1 Start --- perform a simple reading task\nReading tests/data/australian.train\n")
-S = readSparseData ("sparse.data")
-print (paste("Data has", nrow(S$X), "points."))
-print (paste("Labels are", unique(S$Y), "."))
-
-
-# --- peform a simple writing task
-cat("\n\nTest #2 Start --- peform a simple writing task\nGetting Iris Matrix\n")
-X = as.matrix(iris[,1:4])
-Y = as.matrix(as.numeric(iris[,5]))
-cat("Write the matrix to file ./australian.data\n")
-t = tempfile()
-writeSparseData (t, X, Y)
-
-
-
-# --- perform a reading/writing task with faulty parameter zeroBased = FALSE
-#		The code should work without any error even with faulty zeroBased parameter, but it should give out a certain warning message with 			 		informations about the underlying problem. 
-cat("\n\nTest #3 Start --- perform a reading/writing task with faulty parameter zeroBased = FALSE\n")
-S = readSparseData ("../../tests/data/a0", zeroBased = FALSE, verbose = TRUE) #zeroBased should be TRUE here
-writeSparseData ( "../../tmp/test.sparse.data", S$X,  S$Y, verbose = TRUE, zeroBased = FALSE)
-
-
-# --- perform a reading/writing task with faulty parameter zeroBased = TRUE
[…]

### 2016-02-04 inst/examples/Wrapper-Example.R

@@ -1,352 +1,0 @@
-#!/usr/bin/Rscript  --vanilla 
-
-#
-# SVMBridge 
-#		(C) 2015, by Aydin Demircioglu
-#
-#		mySVM_wrapper.R
-# 
-# SVMBridge is free software: you can redistribute it and/or modify
-# it under the terms of the GNU Lesser General Public License as published
-# by the Free Software Foundation, either version 3 of the License, or
-# (at your option) any later version.
-#
-# SVMBridge is distributed in the hope that it will be useful,
-# but WITHOUT ANY WARRANTY; without even the implied warranty of
-# MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
-# GNU Lesser General Public License for more details.
-#
-# Please do not use this software to destroy or spy on people, environment or things.
-# All negative use is prohibited.
-#
-
-
-#' Create training arguments for mySVM
-#'
-#' @param	x			SVM object
-#' @param	trainDataFile		file to read training data from.
-#' @param	modelFile		path to model, defaults to a temporary file (given by R).
-#' @param	kernelCacheSize		kernel cache parameter
-#' @param	svmType			type of svm
-#' @param	cost			cost parameter C.
-#' @param	gamma			gamma parameter, note: RBF kernel used by pegasos is exp(-0.5 ...).
-#'
-#' @return	args			arguments for training
-#'
-createTrainingArguments.mySVM = function (
-	x,
-	...,
-	trainDataFile = "",
-	modelFile = "",
-	extraParameter = "",
-	kernelCacheSize = 1024,
-	cost = 1,
-	svmType = "-1",
-	useBias = FALSE,
-	gamma = 1,
-	epsilon = 0.001, 
-	degree = -1,
-	coef0 = -1,
-	nu = -1,
-	shrinking = -1,
-	probabilityEstimates = -1,
-	weight = -1,
-	n = -1,
-	kernelType = "rbf",
-	quietMode = FALSE)
-{
-		svmTypeParameter = ""
-		if (svmType == "CSVC" || svmType == "C-SVC")
-			svmTypeParameter = "-s 0"
-		if (svmType == "nuSVC" || svmType == "nu-SVC")
-			svmTypeParameter = "-s 1"
-		if (svmType == "one-class SVM" || svmType == "oneClassSVM")
-			svmTypeParameter = "-s 2"
-		
[…]

### 2016-02-05 R/trainSVM.R
small change in trainSVM, docu update, add example dir

@@ -29,7 +29,7 @@
 #' @param	trainDataY		Vector comprising the labels for the data. This cannot be used at the same time as trainDataFile.
 #' @param	trainBinaryPath		Full path to the training binary to call.
 #' @param	extraParameter		Extra parameters for solver, passed to the createTrainingArguments function of the wrapper.
-#' @param	modelFile		Path to the model file to create. If none specified, a temporary file will be created and read into memory.
+#' @param	modelFile		Path to the model file to create. If none specified, a temporary file will be created.
 #' @param	verbose		Be verbose?
 #' @param	subsamplingRate		Rate to subsample the data. 
 #' @param	subsamplingMethod		Method to subsample the data with.
@@ -56,7 +56,7 @@ trainSVM = function(
 	subsamplingMethod = "cutoff",
 	verbose = FALSE,
 	timeOut = -1,
-	readModelFile = FALSE,
+	readModelFile = TRUE,
 	...) {
 	
 	#expand possible tilde characters in the path and get rid of backslashes

### 2016-02-05 examples/quickstart.R

@@ -1,0 +1,22 @@
+
+# load the SVMBridge 
+library(SVMBridge)
+
+# add the prepackaged wrapper 
+addSVMPackage("LIBSVM", wrapperPath =  file.path (path.package("SVMBridge"), "wrapper"))
+
+# let SVMBridge find the compiled/downloaded LIBSVM package. 
+found = findSVMSoftware (method = 'LIBSVM', searchPath = "./software/LIBSVM")
+
+# use the SVMBridge to train LIBSVM
+svmObj =  trainSVM("LIBSVM", trainDataFile = './data/australian_scale', cost = 1, gamma = 1)
+
+# print the time needed for training
+cat( paste("LIBSVM took", svmObj$trainTime, "seconds to train.\n"))
+
+# use the SVMBridge to test LIBSVM (here on the same data set)
+testObj = testSVM(testDataFile = './data/australian_scale', model = svmObj$model)
+
+# print the test error
+cat( paste("Test error is ", round(testObj$testError,2), ". Testing took ", round (testObj$testTime,3) , " seconds.\n", sep = ''))
+

### 2016-02-05 examples/software/LIBSVM/svm-predict.c

@@ -1,0 +1,619 @@
+#include <math.h>
+#include <stdio.h>
+#include <ctype.h>
+#include <stdlib.h>
+#include <string.h>
+#include <errno.h>
+#include "svm.h"
+
+#define Malloc(type,n) (type *)malloc((n)*sizeof(type))
+
+// yes, this was not the best idea..
+
+
+typedef signed char schar;
+typedef float Qfloat;
+#ifndef min
+template <class T> static inline T min(T x,T y) { return (x<y)?x:y; }
+#endif
+#ifndef max
+template <class T> static inline T max(T x,T y) { return (x>y)?x:y; }
+#endif
+template <class T> static inline void swap(T& x, T& y) { T t=x; x=y; y=t; }
+template <class S, class T> static inline void clone(T*& dst, S* src, int n)
+{
+    dst = new T[n];
+    memcpy((void *)dst,(void *)src,sizeof(T)*n);
+}
+static inline double powi(double base, int times)
+{
+    double tmp = base, ret = 1.0;
+
+    for(int t=times; t>0; t/=2)
+    {
+        if(t%2==1) ret*=tmp;
+        tmp = tmp * tmp;
+    }
+    return ret;
+}
+#define INF HUGE_VAL
+#define TAU 1e-12
+
+//
+// Kernel Cache
+//
+// l is the number of total data items
+// size is the cache size limit in bytes
+//
+class Cache
+{
+public:
+    Cache(int l,long int size);
+    ~Cache();
+
+    // request data [0,len)
+    // return some position p where [p,len) need to be filled
+    // (p >= len if nothing needs to be filled)
+    int get_data(const int index, Qfloat **data, int len);
+    void swap_index(int i, int j);  
+private:
+    int l;
+    long int size;
+    struct head_t
+    {
+        head_t *prev, *next;    // a circular list
+        Qfloat *data;
+        int len;        // data[0,len) is cached in this entry
+    };
+
+    head_t *head;
+    head_t lru_head;
+    void lru_delete(head_t *h);
+    void lru_insert(head_t *h);
+};
+
+
+//
+// Kernel evaluation
+//
+// the static method k_function is for doing single kernel evaluation
+// the constructor of Kernel prepares to calculate the l*l kernel matrix
+// the member function get_Q is for getting one column from the Q Matrix
+//
+class QMatrix {
+public:
+    virtual Qfloat *get_Q(int column, int len) const = 0;
[…]

### 2016-02-05 examples/software/LIBSVM/svm-train.c

@@ -1,0 +1,432 @@
+#include <stdio.h>
+#include <stdlib.h>
+#include <string.h>
+#include <ctype.h>
+#include <errno.h>
+#include "svm.h"
+#include <math.h>
+#include <string>
+#define Malloc(type,n) (type *)malloc((n)*sizeof(type))
+
+void print_null(const char *s) {}
+
+void exit_with_help()
+{
+	printf(
+	"Usage: svm-train [options] training_set_file [model_file]\n"
+	"options:\n"
+	"-s svm_type : set type of SVM (default 0)\n"
+	"	0 -- C-SVC		(multi-class classification)\n"
+	"	1 -- nu-SVC		(multi-class classification)\n"
+	"	2 -- one-class SVM\n"
+	"	3 -- epsilon-SVR	(regression)\n"
+	"	4 -- nu-SVR		(regression)\n"
+	"-t kernel_type : set type of kernel function (default 2)\n"
+	"	0 -- linear: u'*v\n"
+	"	1 -- polynomial: (gamma*u'*v + coef0)^degree\n"
+	"	2 -- radial basis function: exp(-gamma*|u-v|^2)\n"
+	"	3 -- sigmoid: tanh(gamma*u'*v + coef0)\n"
+	"	4 -- precomputed kernel (kernel values in training_set_file)\n"
+	"-d degree : set degree in kernel function (default 3)\n"
+	"-g gamma : set gamma in kernel function (default 1/num_features)\n"
+	"-r coef0 : set coef0 in kernel function (default 0)\n"
+	"-c cost : set the parameter C of C-SVC, epsilon-SVR, and nu-SVR (default 1)\n"
+	"-n nu : set the parameter nu of nu-SVC, one-class SVM, and nu-SVR (default 0.5)\n"
+	"-p epsilon : set the epsilon in loss function of epsilon-SVR (default 0.1)\n"
+	"-m cachesize : set cache memory size in MB (default 100)\n"
+	"-e epsilon : set tolerance of termination criterion (default 0.001)\n"
+	"-h shrinking : whether to use the shrinking heuristics, 0 or 1 (default 1)\n"
+	"-b probability_estimates : whether to train a SVC or SVR model for probability estimates, 0 or 1 (default 0)\n"
+	"-wi weight : set the parameter C of class i to weight*C, for C-SVC (default 1)\n"
+	"-v n: n-fold cross validation mode\n"
+	"-q : quiet mode (no outputs)\n"
[…]

### 2016-02-05 examples/software/LIBSVM/svm.cpp

@@ -1,0 +1,3433 @@
+#include <math.h>
+#include <stdio.h>
+#include <stdlib.h>
+#include <ctype.h>
+#include <float.h>
+#include <string.h>
+#include <stdarg.h>
+#include <limits.h>
+#include <locale.h>
+#include <sys/time.h>
+#include "svm.h"
+#include <sstream>
+#include <string>
+#include <iomanip>
+
+#define safetywalltime  5
+
+// check for walltime every ...th iteration
+#define CHECKINTERVAL 10
+
+// how to convert given times into seconds, 60=minutes, 1=seconds
+#define TIMEFACTOR 60
+
+
+
+int libsvm_version = LIBSVM_VERSION;
+typedef float Qfloat;
+typedef signed char schar;
+#ifndef min
+template <class T> static inline T min(T x,T y) { return (x<y)?x:y; }
+#endif
+#ifndef max
+template <class T> static inline T max(T x,T y) { return (x>y)?x:y; }
+#endif
+template <class T> static inline void swap(T& x, T& y) { T t=x; x=y; y=t; }
+template <class S, class T> static inline void clone(T*& dst, S* src, int n)
+{
+	dst = new T[n];
+	memcpy((void *)dst,(void *)src,sizeof(T)*n);
+}
+static inline double powi(double base, int times)
+{
+	double tmp = base, ret = 1.0;
+
+	for(int t=times; t>0; t/=2)
+	{
+		if(t%2==1) ret*=tmp;
+		tmp = tmp * tmp;
+	}
+	return ret;
+}
+#define INF HUGE_VAL
+#define TAU 1e-12
+#define Malloc(type,n) (type *)malloc((n)*sizeof(type))
+
+
+extern std::string modelPath = "./";
+struct svm_model* hack_model = NULL;
+
+
+int walltime = -1;
+int savetime = -1;
+double subsamplingAmount = -1;
+
+double now()
+{
+	struct timeval t;
+	gettimeofday(&t, NULL);
+	return (t.tv_sec + 1e-6 * t.tv_usec);
+}
+
+
+static void print_string_stdout(const char *s)
+{
+	fputs(s,stdout);
+	fflush(stdout);
+}
+static void (*svm_print_string) (const char *) = &print_string_stdout;
+#if 1
+static void info(const char *fmt,...)
+{
+	char buf[BUFSIZ];
+	va_list ap;
+	va_start(ap,fmt);
[…]

### 2016-02-09 R/testSVM.R
fix multiclass and add test

@@ -54,7 +54,7 @@ testSVM = function(
 	...) {
 	
 	if (verbose == TRUE) {
-		cat("Testing SVM.")
+		cat("Testing SVM", method, ".\n")
 	}
 	
 	# sanity checks for parameters
@@ -107,8 +107,8 @@ testSVM = function(
 	
 	# general modifications
 	if (verbose == TRUE) {
-		cat("  Path of binary for testing is ", testBinaryPath)
-		cat("  Binary for testing is ", testBinary)
+		cat("    Path of binary for testing is ", testBinaryPath, "\n")
+		cat("    Binary for testing is ", testBinary, "\n")
 	}
 
 	# take care of data. if testdata or traindata is given,
@@ -128,13 +128,13 @@ testSVM = function(
 	if (is.null(testDataX) == FALSE) {
 		testDataFile = tempfile()
 		if (verbose == TRUE)
-			cat("  Writing given test data as ", testDataFile)
+			cat("    Writing given test data as ", testDataFile, "\n")
 		writeSparseData (filename = testDataFile, X = testDataX, Y = testDataY)
 		#e1071::write.matrix.csr(testDataX, testDataFile, testDataY)
 	} 
 
 	if (verbose == TRUE) 
-		cat("  Test Data is now in ", testDataFile)
+		cat("    Test Data is now in ", testDataFile, "\n")
 
 	results = list()
 
@@ -142,12 +142,12 @@ testSVM = function(
 	if (is.null(model) == FALSE) {
 		modelFile = tempfile()
 		if (verbose == TRUE)
-			cat("  Writing model in memory to disk as ", modelFile)
+			cat("    Writing model in memory to disk as ", modelFile, "\n")
 		
 		writeModel (SVMObject, model = model, modelFile = modelFile, verbose = verbose)
 	} else {
 		if (verbose == TRUE)
-			cat("  Reading model from file ", modelFile)
+			cat("    Reading model from file ", modelFile, "\n")
 	}
 
 	# check if we need to create a temporary file
@@ -163,14 +163,14 @@ testSVM = function(
 		...)
     
     if (verbose == TRUE) {
-		cat("  Generated test arguments ", args)
+		cat("    Generated test arguments ", args, "\n")
 	}
       
 	# probably FIXME because of windows systems
 	testTime = microbenchmark::microbenchmark(s <- system3(testBinaryPath, args, verbose = verbose), times = 1L)$time / 1e9
     
[…]

### 2016-02-09 R/trainSVM.R

@@ -88,8 +88,8 @@ trainSVM = function(
 	
 	# general modifications
 	if (verbose == TRUE) {
-		cat("    Path of binary for training is ", trainBinaryPath)
-		cat("    Binary for training is ", trainBinary)
+		cat("    Path of binary for training is ", trainBinaryPath, "\n")
+		cat("    Binary for training is ", trainBinary, "\n")
 	}
 
 	
@@ -131,13 +131,13 @@ trainSVM = function(
 	if (is.null(trainDataX) == FALSE) {
 		trainDataFile = tempfile()
 		if (verbose == TRUE)
-			cat("    Writing given data as ", trainDataFile)
+			cat("    Writing given data as ", trainDataFile, "\n")
 		#e1071::write.matrix.csr(trainDataX, trainDataFile, trainDataY)
 		writeSparseData (filename = trainDataFile, X = trainDataX, Y = trainDataY)
 	} 
 
 	if (verbose == TRUE) 
-		cat("    Train Data is now in ", trainDataFile)
+		cat("    Train Data is now in ", trainDataFile, "\n")
 
 	
 	# at this point we need to have a train data file on disk

### 2016-02-09 tests/testthat/test_generalWorkflow.R

@@ -38,6 +38,7 @@ test_that("general workflow works with binary packages.", {
 	source ("wrappertests.R")
 	source ("detectModeltests.R")
 	source ("downloadSoftware.R")
+	source ("multiClasstests.R")
 	
 
 
@@ -45,7 +46,7 @@ test_that("general workflow works with binary packages.", {
 	# else SVM does not work 100p the same, as the data is shuffled.
 	set.seed(42)
 
-	verbose = TRUE
+	verbose = FALSE
 	cat ("\nVerbose is", verbose, "\n")
 
 	modelFile = tempfile()
@@ -60,6 +61,12 @@ test_that("general workflow works with binary packages.", {
 	trainDataY = data.matrix(as.numeric(shufflediris [1:100, 5]))
 	testDataX = data.matrix(shufflediris [-(1:100),1:4])
 	testDataY = data.matrix(as.numeric(shufflediris [-(1:100), 5]))
+	
+	# multiclass
+	mcTestDataY = testDataY
+	mcTrainDataY = trainDataY
+	
+	# binarize labels
 	trainDataY [trainDataY==3] = 1
 	testDataY [testDataY==3] = 1
 	trainDataY [trainDataY==2] = -1
@@ -82,10 +89,11 @@ test_that("general workflow works with binary packages.", {
 	softwareBaseDir = tempdir()
 	
 	# this stuff is for debugging locally without reloading the packages each time
-	#solvers = c("LIBSVM")
-	#softwareBaseDir = "/tmp/software"
-	#if (file.exists(softwareBaseDir) == FALSE) {
-	if (TRUE == TRUE) {
+	solvers = c("LIBSVM")
+	solvers = c("LIBSVM", "LASVM", "BSGD", "SVMperf", "BVM", "CVM", "LLSVM")
+	softwareBaseDir = "/tmp/software"
+	if (file.exists(softwareBaseDir) == FALSE) {
+	#if (TRUE == TRUE) {
 		for (solver in solvers) {
 			cat ("Downloading and building software ", solver, "\n")
 			softwareDir = downloadSoftware (solver, softwareDir = softwareBaseDir, verbose = verbose)
@@ -106,19 +114,26 @@ test_that("general workflow works with binary packages.", {
 	##  now do all the thirty different ways of calling trainSVM 
 	
 	for (solver in solvers) {
-		cat ("Wrapper test for solver", solver)
+		cat ("\nWrapper test for solver", solver, ": ")
[…]

### 2016-02-11 paper/article.tex
small fixes, add sketch of paper temporarily

@@
 
-Support Vector Machines (SVM) are linear classifiers
-that try to maxime the margin of a binary problem \citep{cortes1995support}.
-As SVMs are only linear classifiers, a kernelized version
-is used for more complex data. Basically, they work by
-solving the  following convex problem:
+Kernelized Support Vector Machines (SVM) \cite{cortes1995support}
+are binary classifiers that use a kernel $k$ to allow for non-linear
+classification decisions
+$x \mapsto \sign( \langle w, \phi(x) \rangle + b)$, where $\phi$ is the
+(non-linear) feature map corresponding to the kernel
+$k(x, x') = \langle \phi(x), \phi(x') \rangle$ in the reproducing kernel
+Hilbert space $\mathcal{H}$.
+We use the RBF kernel $k(x,x') = e^{-\gamma || x-x' ||^2}$, since it
+yields excellent performance and enjoys a universal approximation
+property.
+The SVM training problem is given by
 \begin{equation}
-        \min_{w \in \mathcal{H}, b \in \R} \quad \frac{1}{2} ||w||^2 + C \cdot \sum_{i=1}^n \max \Big( 0, 1 - y_i \big( \langle w, \varphi(x_i) \rangle_{\mathcal{H}} + b \big) \Big).
-        \label{SVMproblemPrimal}
+\min_{w \in \mathcal{H}, b \in \R} \quad \frac{1}{2} ||w||^2 + C \cdot \sum_{i=1}^n \max \Big( 0, 1 - y_i \big( \langle w, \varphi(x_i) \rangle_{\mathcal{H}} + b \big) \Big),
+\label{primalproblem}
 \end{equation}
-Here 
-$\mathcal{D}=\{(x_1, y_1), \dots, (x_n, y_n)\} \in (X \times \{ \pm 1\})^n$
-is the labeled data,
-and  $\varphi : X \to \mathcal{H}$ is a feature map into a reproducing
-kernel Hilbert space $\mathcal{H}$, corresponding to a positive definite
-(Mercer) kernel function $k : X \times X \to \R$, fulfilling
-$\langle \varphi(x), \varphi(x') \rangle_{\mathcal{H}} = k(x, x')$ for
-all $x, x' \in X$. Furthermore, $C > 0$ is a regularization parameter.
-It controls the complexity of the SVM model.
+where $\big\{(x_1, y_1), \dots, (x_n, y_n)\big\}$ are the labeled
+training points and $C > 0$ is a regularization parameter that controls
+the complexity of the predictive model.
 
-Often, the problem is not directly solved, but dualized via Lagrangian theory.
-Then the problem can be stated as 
-\begin{align}
-        \max_{\alpha \in \R^n} \quad & \sum_{i=1}^n \alpha_i - \frac{1}{2} \sum_{i,j=1}^n \alpha_i \alpha_j y_i y_j k(x_i, x_j) \label{SVMproblemDual} \\
-        \text{s.t.} \quad & \sum_{i=1}^{n} y_i \alpha_i = 0 \text{~~and~~} 0 \leq \alpha_i \leq C \,\,\, \forall \, i \in \{1, \dots, n\}. \notag
-        \enspace
-\end{align}
-The solution then takes the form $w = \sum_{i=1}^n \alpha_i y_i \varphi(x_i)$
-and the offset~$b$ can be computed from the Karush-Kuhn-Tucker (KKT)
-complementarity conditions.
-
-SVMs classify by using the model given by
-$h_{w,b}(x) = \textrm{sign}(\langle w, \phi(x)\rangle_{\mathcal{H}} + b)$
-
-Different kernels can be used, e.g. the polynomial kernel or the 
-RBF kernel $k(x,x') = e^{-\gamma || x-x' ||^2}$, where
-$\gamma$ and $d$ are parameters of the kernel, that must
-be tuned to the given data.
-
-SMO.
-
-On the practical side, LIBSVM can be regarded as the 
-reference implementation of the SMO and the 'golden standard'
-of SVM solvers \citep{cc01a}.
+The problem, although convex, can be solved by different means,
+e.g. LIBSVM employs a dual
+decomposition method, while  BSGD is a primal stochastic
+gradient descent solver.
 
@@
 
-There are only a few SVM libraries readily available from within \proglang{R}.
-This includes the well-known packages like \pkg{kernlab} (via \code{ksvm()})~\citep{kernlab}, 
+In general, nearly all SVM solvers come as binary packages,
+often written in C. Therefore it is not possible to use them 
+directly from within \proglang{R}. Only a few SVM libraries are readily available,
+includ LIBSVM inside the well-known packages like \pkg{kernlab} (via \code{ksvm()})~\citep{kernlab}, 
 and \pkg{e1071} (via \code{svm()})~\citep{dimitriadou2008misc}
@@
 Other SVM solvers, like the BudgetedSVM package~\citep{djuric2014budgetedsvm}, 
-which contains BSGD and LLSVM, or SVMperf~\citep{joachims2009sparse}, cannot be called directly from  \proglang{R}.
+which contains BSGD and LLSVM, or SVMperf~\citep{joachims2009sparse}, cannot 
+be called directly from  \proglang{R}.
 Instead, a system command has to be issued.
@@
 
-\subsection{Wrapper}
+\section{General worflow}
+
+To use the SVMBridge, one usually perform the following steps:
+Make the external wrapper incl. software binaries known to the SVMBridge, 
+load the data, train a model, use the model for prediction.
+
+\subsection{Wrappers}
 
@@
 the binaries and printing.
-
-Note that the search for binaries might have a huge toll on HPC clusters,
-so it is adviced to specify the paths directly instead of relying on the 
-automatism.
-
 We provided default wrappers for LIBSVM, LASVM, CVM/BVM,
@@
 
-\section{Using a Wrapper}
-
-To use the SVMBridge, one usually perform the following steps:
-Add the external wrapper to the SVMBridge, load the data, train a model, use the
-model for prediction.
-
-Notice that the data must be written by the SVMBridge before calling 
-the corresponding SVM Solver, as there is no way to pass data via memory
-to command line tools. Therefore
-
-\subsection{Adding the Wrapper to the SVMBridge}
-
-The first step is to make the SVMBridge aware of a new wrapper.
-This is simply done by calling \code{addSVMPackage()}.
-There are several options.
+Make the SVMBridge aware of a new wrapper is
+simply done by calling \code{addSVMPackage()}.
 The SVMBridge comes with an easy mechanism to search 
@@
 
-\subsection{Reading Sparse Data}
+\subsection{Data}
 
-Most SVM solver work with the sparse data format.
-This format consists of:
-Each line (delimited by a CR/LF) is given by a label
-and the non-zero components of the data point,
-e.g. to encode the vector six dimensional vector 
- $ ( 0 , 0 , 0 , 0.2 , 0 , 1.4 ) $
-belonging to class $3$, the sparse 
-data format would contain the line {3 4:0.2 6:1.4}.
+Most SVM solver work with the LIBSVM sparse data format.
+This format saves the labels together with all non-zero indicies of every data point, e.g.
+$(0, 0, 0, 0.2 , 0, 1.4)^T$  belonging to class $3$, will be encoded as \texttt{'3 4:0.2 6:1.4'}.
 
-Although reading these files into \proglang{R} is possible via either
-the \pkg{kernlab} or the \pkg{e1071} routines, both only provide \proglang{R} solutions,
+Although reading these files into \proglang{R} is possible via e.g. the 
+\pkg{e1071} package, most only provide pure \proglang{R} solutions,
 and therefore suffer from suboptimal performance.
@@
 reading and writing, which are implemented in \proglang{C++} and
-therefore are nearly two orders of magnitudes faster than
+therefore can be nearly two orders of magnitudes faster than
 the corresponding e1071 routines.
-Notice, that currently only dense matrices are supported,
-TODO: work around this with the Sparse Matrix Package.
 
-\subsection{Training a model}
+\subsection{Models}
+
+Internally models contain support vectors and their coefficients.
+
+\subsection{Training an SVM model}
 
@@
 R> model = trainSVM (method = "LIBSVM", cost = 1.0, gamma = 2.0,
-			trainDataFile = "./data.sparse",
-			epsilon = 0.042,
-			modelFile = modelFile,
-			useBias = TRUE,
-			verbose = verbose
-		)
+	trainDataFile = "./data.sparse", epsilon = 0.042, modelFile = modelFile)
 \end{CodeInput}
-\begin{CodeOutput}
-...
-\end{CodeOutput}
 \end{CodeChunk}
@@
 
-\subsection{Predicions}
+\subsection{Testing a model}
 
-Predicting from a trained model is rather easy, as there are usually not many options.
-As with training, the predict routine will accept a model either in memory or from a file.
-The same is true for the data to predict.  Again notice that data and models in memory 
-must be written to disk prior to calling the prediction binary.
+%Predicting from a trained model is rather easy, as there are usually not many options.
+Testing (or doing predictions) needs a trained SVM model as well as test data to work on.
+As with training, testSVM will accept data and models either from memory or from a file.
+%The same is true for the data to predict.  Again notice that data and models in memory 
+%must be written to disk prior to calling the prediction binary.
 
@@
 \begin{CodeInput}
-R> predictions = testSVM (model, data = '')
-R> head(predictions)
+R> testObj = testSVM (model, testDataFile = './australian_scale', 
+	readPredictions = TRUE)
 \end{CodeInput}
-\begin{CodeOutput}
-...
-\end{CodeOutput}
 \end{CodeChunk}
 
-\subsection{Optimization Values}
+The returned object will contain meta information like testing times as well as the predictions.
 
-At times, some of the training values are of interest,
-e.g. when comparing different solvers, the primal value
-of the SVM problem as well as the dual value might  be of interest.
-These can be computed via the \code{optimizationValues()} routine.
-
-Specifically, this routine compute the following values:
-\code{weight} is given by $\frac{1}{2} w^T w$, with
-$w = \sum_i \alpha_i s_i$. Furthermore, we have
-that $C \sum_i max (0, 1 - ..) $, and 
-\code{primal} is then given by $primal = w + CHinge$.
-\code{dual} on the other hand is simply
-$w - \sum_i |\alpha_i|$. In general it must hold
-that $primal \geq dual$. 
-WHAT IF NOT? what does it mean?
-
-\subsection{Helper  Routines}
-
-A few routines have been placed into the package that were helpful
-with testing. Namely, there are several generators for synthetical data sets,
-collected from different publications. 
-WHAT ELSE? can routines be used outside? detectTilde e.g.?
-
-\subsection{Platform Considerations}
+\subsection{Other Considerations}
 
@@
 is hard to achieve. From a users perspective, it can be used on all three
-major platforms (Linux, MacOS, Windows). Additional tests should
-be started to make sure the package works as intended.
-
-\subsection{Performance Considerations}
+major platforms (Linux, MacOS, Windows). The tests should
+be executed to make sure the package works as intended.
 
@@
 
-\subsection{Training Call}
-
-A simple S3 method has to be created that needs to be called
-\code{createTrainingArguments.LIBSVM}. Apart from the very
-first argument (taking the object itself), all other paramters
-can be chosen freely. After training, from the output several
-training information will be extracted by the 
-\code{extractTrainingInfo()} function.
-
-\subsection{Testing Call}
-
-Similar to the training call, \code{createTestArguments()} needs to
-writen. Again, it should accept three variables called testDataFile and
-modelFile as well as predictionFile. There is also a
-\code{extractTestInfo()} that will extract the important informations
-from the output of the test binary.
-
-\subsection{Handling Models}
-
-As only LIBSVM models are being used, one can readily use the 
-inbuild functions of the SVMBridge. If the solver works with its
-own model, one needs to rewrite the readModel method of the wrapper.
-This will only obtain the modelFile and will return the model
-in the model format (S3 Object with XY, describe it somewhere else)
-SVMBridge needs also a writeModel functions, as models passed in
-memory needs to be written to disk. .
-Finally it is necessary to access the predictions itself. For this one
-needs to write the \code{readPredictions()} function. For LIBSVM this
-will read the prediction files that is being written by LIBSVM to disk.
-
-\subsection{Other functions}
-
-Every wrapper should also be able to find itself in a given directory tree.
-For this the S3 method \code{findSoftware()} must be provided.
-
-At last, for querying reasons, there is also a print routine
-so that wrapper can print out any further information about itself.
-To see the currently loaded wrapper, one can call 
-\code{overviewAllSVMSolver()}.
-
 \section{Conclusion}
 
-We provided a simple framework, \pkg{SVMBridge},  to attach many SVM solvers to the \proglang{R} landscape.
-\pkg{SVMBridge} does this by an easily extensible wrappers that manage the exchange of data and models
+We provided a simple framework, \pkg{SVMBridge}, to attach binary SVM solvers easily to the \proglang{R} landscape.
+The \pkg{SVMBridge} does this by an easily extensible wrappers that manage the exchange of data and models
 via command line.
-By providing such a wrapper, linking any SVM solver to \proglang{R} becomes much easier.
+%By providing such a wrapper, linking any SVM solver to \proglang{R} becomes much easier.
 

### 2016-02-11 R/trainSVM.R

@@ -27,7 +27,7 @@
 #' @param	trainDataFile		Filename to read training data from. This cannot be used at the same time as trainDataX/trainDataY.
 #' @param	trainDataX		Matrix comprising the data. This cannot be used at the same time as trainDataFile.
 #' @param	trainDataY		Vector comprising the labels for the data. This cannot be used at the same time as trainDataFile.
-#' @param	extraParameter		Extra parameters for solver, passed to the createTrainingArguments function of the wrapper.
+#' c@param	extraParameter		Extra parameters for solver, passed to the createTrainingArguments function of the wrapper.
 #' @param	modelFile		Path to the model file to create. If none specified, a temporary file will be created.
 #' @param	verbose		Be verbose?
 #' @param	subsamplingRate		Rate to subsample the data. 

### 2016-02-12 R/findSVMWrapper.R
update tutorial

@@ -17,7 +17,7 @@
 #
 
 
-#' findSVMWrapper
+#' Find a given SVM Wrapper inside a search path.
 #'	
 #' Given a search path, it will try to find the corresponding wrapper for the given method.
 #' This routine will also source the file, if specified.
@@ -37,7 +37,7 @@ findSVMWrapper <- function (method = NA, searchPath = NA, recursive = TRUE, sour
 	checkmate::assertString (searchPath)
 	
 	if (verbose == TRUE) {
-		cat("    -Finding wrapper for ", method, "\n")
+		cat("Finding wrapper for ", method, "\n")
 	}
 	
 	# look for tilde characters and expand them
@@ -46,7 +46,7 @@ findSVMWrapper <- function (method = NA, searchPath = NA, recursive = TRUE, sour
 	# get the SVM object
 	SVMObject = getSVMObject (method)
 	if (checkmate::testNull (SVMObject) == TRUE) {
-		stop ("Cannot find SVM Obejct for method ", method, ". Please create it first.")
+		stop ("Cannot find SVM Object for method ", method, ". Please create it first.")
 	}
 	checkmate::assertString (SVMObject$wrapperName)
 
@@ -57,7 +57,7 @@ findSVMWrapper <- function (method = NA, searchPath = NA, recursive = TRUE, sour
 		cat("    Looking for a wrapper with regex ", pattern, "\n")
 	}
 	
-	files <- list.files (searchPath, pattern = pattern, recursive = recursive)
+	files = list.files (searchPath, pattern = pattern, recursive = recursive)
 	if (length(files) > 1) {
 		warning ("Found multiple wrappers. Taking the first one: ", files[1])
 	}
@@ -69,7 +69,7 @@ findSVMWrapper <- function (method = NA, searchPath = NA, recursive = TRUE, sour
 	
 	wrapperPath = file.path (searchPath, files[1])
 	if (verbose == TRUE) { 
-		cat ("    -Found wrapper at ", wrapperPath, "\n") 
+		cat ("    Found wrapper at ", wrapperPath, "\n") 
 	}
 
 	if (source == TRUE) {

### 2016-02-12 R/trainSVM.R
update

@@ -27,7 +27,7 @@
 #' @param	trainDataFile		Filename to read training data from. This cannot be used at the same time as trainDataX/trainDataY.
 #' @param	trainDataX		Matrix comprising the data. This cannot be used at the same time as trainDataFile.
 #' @param	trainDataY		Vector comprising the labels for the data. This cannot be used at the same time as trainDataFile.
-#' c@param	extraParameter		Extra parameters for solver, passed to the createTrainingArguments function of the wrapper.
+#' @param	extraParameter		Extra parameters for solver, passed to the createTrainingArguments function of the wrapper.
 #' @param	modelFile		Path to the model file to create. If none specified, a temporary file will be created.
 #' @param	verbose		Be verbose?
 #' @param	subsamplingRate		Rate to subsample the data. 

### 2016-02-29 R/predictSVM.R
fix #43, fix LLSVM wrapper by removing unnecessary label inversions

@@ -1,0 +1,150 @@
+#
+# SVMBridge 
+#		(C) 2015, by Aydin Demircioglu
+#
+# SVMBridge is free software: you can redistribtrainSVMute it and/or modify
+# it under the terms of the GNU Lesser General Public License as published
+# by the Free Software Foundation, either version 3 of the License, or
+# (at your option) any later version.
+#
+# SVMBridge is distributed in the hope that it will be useful,
+# but WITHOUT ANY WARRANTY; without even the implied warranty of
+# MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
+# GNU Lesser General Public License for more details.
+#
+# Please do not use this software to destroy or spy on people, environment or things.
+# All negative use is prohibited.
+#
+ 
+
+#' Do prediction with a trained SVM on given data.
+#'
+#' This is the main routine for predictions. If needed, it will add dummy labels to the data given and then call testSVM.
+#'
+#' @param	method		Name of the method/solver to use for testing. If NULL, a detectModelTypeFromFile will be called on the given model (or modelfile).
+#' @param	testDataFile		File to read training data from. Cannot be specified at the same time as testDataX and testDataY. See notes.
+#' @param	testDataX		Matrix to read training data from. Cannot be specified at the same time as testDataFile. 
+#' @param	testDataY		Matrix to read training label from. Cannot be specified at the same time as testDataFile. Optional, see notes.
+#' @param	model		A trained SVM model. Cannot be specified at the same time as modelFile.
+#' @param	modelFile		Path of the model file. Cannot be specified at the same time as model.
+#' @param	readPredictions		Shall the predictions be read back into memory and added to the model?
+#' @param	predictionsFile		File to which the predictions will be written.
+#' @param	extraParameter			Extra parameters for solver in form of a string.
+#' @param	verbose			Be verbose?
+#' @param	...		Further parameters that will be passed to the createTestArguments function of the wrapper.
+#'
+#' @note		testDataFile and testDataX,testDataY are exclusive parameters, i.e. you cannot specify both.
[…]

### 2016-02-29 inst/wrapper/LLSVM_wrapper.R

@@ -213,17 +213,19 @@ readModel.LLSVM = function (x, modelFile = "./model", verbose = FALSE) {
 	model$totalSV = nrow(supportvectors)
 	model$label = as.numeric(unlist(strsplit(order, " ")))
 
-	# do we need to invert the labels? in this case we invert the coefficients
-	if (invertLabels == TRUE) {
-		if (verbose == TRUE)
-			cat(" Inverting Labels.")
-
-		# invert alphas
-		model$alpha = -model$alpha
-
-		# this is also needed..
-		model$bias = -bias
-	}
+# 	# do we need to invert the labels? in this case we invert the coefficients
+# 	if (invertLabels == TRUE) {
+# 		if (verbose == TRUE)
+# 			cat(" Inverting Labels.")
+# 
+# 		# invert alphas
+# 		model$alpha = -model$alpha
+# 
+# 		# this is also needed..
+# 		model$bias = -bias
+# 
+# 		model$label = c(-1, 1)
+# 	}
 
 	# close connection
 	close(con)
@@ -254,7 +256,26 @@ writeModel.LLSVM = function (x, model = NA, modelFile = "./model", verbose = FAL
 	gamma = model$gamma * 2
 	degree = 1
 	coef = 1
+# 
+# 	# labels
+# 	invertLabels = FALSE
+# 	if (model$label[1] == 1) {
+# 		invertLabels = TRUE
+# 	}
+# 	
+# 	# do we need to invert the labels? in this case we invert the coefficients
+# 	if (invertLabels == TRUE) {
+# 		if (verbose == TRUE)
+# 			cat(" Inverting Labels.")
+# 
+# 		# invert alphas
+# 		model$alpha = -model$alpha
+# 
+# 		# this is also needed..
+# 		model$bias = -model$bias
+# 	}
 
+	
 	# write header
 	modelFileHandle <- file(modelFile, open = "w+")
 	writeLines(paste ("ALGORITHM: 3", sep = ""), modelFileHandle )
@@ -269,11 +290,12 @@ writeModel.LLSVM = function (x, model = NA, modelFile = "./model", verbose = FAL
 	writeLines(paste ("KERNEL_COEF_PARAM:", coef, sep = " "), modelFileHandle )
 	writeLines(paste ("MODEL:", sep = ""), modelFileHandle )
 
+	
 	for (r in 1:nrow(model$SV)) {
 		wline = paste0 (model$w[r])
[…]

### 2016-02-29 src/SparseData.cpp

@@ -192,12 +192,24 @@ List readSparseData (std::string filename, unsigned long  skipBytes = 0, bool ve
 			inst_max_index = -1;
 			double i = strtod(p,&endptr);
 			
-			if(p == NULL) // empty line
+			// empty line 
+			if(p == NULL) {
+				if (verbose == TRUE) 
+					Rcout << "Warning: empty line.\n";  
 				break;
+			}
 			
-			if(endptr == p || *endptr != '\0')
-				break;
-
+			if(endptr == p || *endptr != '\0') {
+				// might be that we have a file without any alphas.
+				string idxp = std::string(p);
+				if (idxp.find(":") != -1) {
+					max_alphacount = 0;
+				} else {
+					break;
+				}
+			}
+			
+			
 			while(1)
 			{
 				index = 0;
@@ -246,7 +258,7 @@ List readSparseData (std::string filename, unsigned long  skipBytes = 0, bool ve
 					idx = idx2;
 					alphacount = 0;
 					//cout << "Fehlertracking: " << "idx = " << idx << endl;
-				}
+				} 
 				
 				errno = 0;
 				index = (int) strtol(idx,&endptr,10); 
@@ -310,33 +322,43 @@ List readSparseData (std::string filename, unsigned long  skipBytes = 0, bool ve
 		{
 			inst_max_index = -1; 
 			readline(fp);
-// 			if (max_alphacount > 1) {
-// 				std::cout << line << "\n";
-// 			}
-			label = strtok(line," \t\n");
 			
-			if(label == NULL) {
-				s << "Error: label = 0 " ;
-				::Rf_error(s.str().c_str());
-				return R_NilValue;
-			}
+			//if datasets contains multiple alpha values, this is the first one, else this is the class label
+			// we also have to check whether there is a label at all or not
+			if (max_alphacount > 0) {
+				label = strtok(line," \t\n");
 			
-			//if datasets contains multiple alpha values, this is the first one, else this is the class label FIXME
-			yR(i,0) = strtod(label, &endptr);
-			//cout << "label " << label << endl;
-			//cout << "endptr " << *endptr << endl;
-			if(endptr == label || *endptr != '\0'){
+				if(label == NULL) {
+					s << "Error: label = 0 " ;
+					::Rf_error(s.str().c_str());
+					return R_NilValue;
+				}
+
+				yR(i,0) = strtod(label, &endptr);
+
[…]

### 2016-02-29 tests/testthat/predicttests.R

@@ -1,0 +1,58 @@
+#
+# SVMBridge 
+#
+#		(C) 2015, by Aydin Demircioglu
+# 
+# SVMBridge is free software: you can redistribute it and/or modify
+# it under the terms of the GNU Lesser General Public License as published
+# by the Free Software Foundation, either version 3 of the License, or
+# (at your option) any later version.
+#
+# SVMBridge is distributed in the hope that it will be useful,
+# but WITHOUT ANY WARRANTY; without even the implied warranty of
+# MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
+# GNU Lesser General Public License for more details.
+#
+# Please do not use this software to destroy or spy on people, environment or things.
+# All negative use is prohibited.
+#
+
+
+predicttests = function (solver, trainDataX, trainDataY, testDataX, testDataY, verbose) {
+	trainObj =  trainSVM(
+		method = solver,
+		trainDataX = trainDataX, 
+		trainDataY = trainDataY, 
+		cost = 1, 
+		gamma = 1, 
+		epsilon = 0.01, 
+		rank = 32,
+		k = 32,
+		budget = 32,
+		readModelFile = TRUE,
+		verbose = verbose
+	)  
+	
+	testObj =  predictSVM(
+		method = solver,
+		testDataX = testDataX,
+		model = trainObj$model,
+		verbose = verbose,
+		readPredictions = TRUE
+	)  
+
+#	print (testObj$predictions)
+	
+	# expect beginning and end of predictions to be -1, -1...1, -1 (works for linux right now, hopefully everywhere)
+	# LASVM predict function is BROKEN, will not wirte predictions!
+	if ((solver != "LASVM")) {
+		testthat::expect_equal (sign(testObj$predictions[1]), -1)
+		testthat::expect_equal (sign(testObj$predictions[2]), -1)
+		testthat::expect_equal (sign(testObj$predictions[49]), 1)
+		testthat::expect_equal (sign(testObj$predictions[50]), -1)
+	}
+	
+	if (verbose == TRUE) {
+		print (testObj)
+	}
+}

### 2016-02-29 tests/testthat/test_SparseData.R

@@ -94,3 +94,12 @@ test_that(" Read/Write operations do work on datasets with non binary labels", {
 
 
 
+test_that(" Reading file without labels does work", {
+	S1 = readSparseData (filename = "../data/australian.nolabels")
+	S2 = readSparseData (filename = "../data/australian.test")
+	expect_equal(S1$X, S2$X)
+	expect_equal(S1$Y, NULL)
+})
+
+
+

### 2016-02-29 tests/testthat/test_generalWorkflow.R

@@ -35,11 +35,12 @@ test_that("general workflow works with binary packages.", {
 	
 	# call without parameters to get an exception
 	source ("cycletests.R")
-	source ("wrappertests.R")
 	source ("detectModeltests.R")
 	source ("downloadSoftware.R")
 	source ("multiClasstests.R")
-	
+	source ("predicttests.R")
+	source ("wrappertests.R")
+
 
 
 	# VERY VERY IMPORTANT!
@@ -77,7 +78,7 @@ test_that("general workflow works with binary packages.", {
 	writeSparseData(filename = trainDataFile, X = trainDataX, Y = trainDataY)
 	testDataFile = tempfile()
 	writeSparseData(filename = testDataFile, X = testDataX, Y = testDataY)
-		
+	
 
 	## 1. test finding all software first
 		
@@ -87,10 +88,12 @@ test_that("general workflow works with binary packages.", {
 
 	solvers = c("LIBSVM", "LASVM", "BSGD", "SVMperf", "BVM", "CVM", "LLSVM")
 	softwareBaseDir = tempdir()
-	
-	# this stuff is for debugging locally without reloading the packages each time
+
+	# debugging: only use a subset of solvers
 #	solvers = c("LIBSVM")
 	#solvers = c("LIBSVM", "LASVM", "BSGD", "SVMperf", "BVM", "CVM", "LLSVM")
+	
+	# this stuff is for debugging locally without reloading the packages each time
 #	softwareBaseDir = "/tmp/software"
 #	if (file.exists(softwareBaseDir) == FALSE) {
 	if (TRUE == TRUE) {
@@ -117,6 +120,11 @@ test_that("general workflow works with binary packages.", {
 		cat ("\nWrapper test for solver", solver, ": ")
 #		testthat::context (paste0(solver, "wrapper"))
 		wrappertests (solver, trainDataX, trainDataY, testDataX, testDataY, verbose)
+	}
+
+	for (solver in solvers) {
+		cat ("\nTest predictions for solver", solver, ": ")
+		predicttests (solver, trainDataX, trainDataY, testDataX, testDataY, verbose)
 	}
 	
 	mcsolvers = c("LIBSVM", "BSGD", "BVM", "CVM")

### 2016-02-29 tests/testthat/wrappertests.R

@@ -55,9 +55,9 @@ wrappertests = function (solver, trainDataX, trainDataY, testDataX, testDataY, v
 	)  
 
 	if (Sys.info()["sysname"] == "Windows") {
-		expErrors = c("LIBSVM" = 0.06, "LASVM" = 0.06, "BSGD" = 0.04, "SVMperf" = 0.06, "BVM" = 0.04, "CVM" = 0.04, "LLSVM" = 0.92) #LLSVM is no joke.
+		expErrors = c("LIBSVM" = 0.06, "LASVM" = 0.06, "BSGD" = 0.04, "SVMperf" = 0.06, "BVM" = 0.04, "CVM" = 0.04, "LLSVM" = 0.08)  
 	} else {
-		expErrors = c("LIBSVM" = 0.06, "LASVM" = 0.06, "BSGD" = 0.04, "SVMperf" = 0.06, "BVM" = 0.04, "CVM" = 0.04, "LLSVM" = 0.94) #LLSVM is no joke.
+		expErrors = c("LIBSVM" = 0.06, "LASVM" = 0.06, "BSGD" = 0.04, "SVMperf" = 0.06, "BVM" = 0.04, "CVM" = 0.04, "LLSVM" = 0.06)  
 	}
 	
 	if (verbose == TRUE) {

### 2016-02-29 examples/quickstart.R
fix #33

@@ -1,6 +1,17 @@
 
+# this tutorial assumes that the compiled LIBSVM software can be found in ./software/LIBSVM.
+# it will create a ./data directory and place the australian data set into it.
+
+
 # load the SVMBridge 
 library(SVMBridge)
+
+# download australian data set from LIBSVM data webseite
+dir.create ("data")
+download.file("https://www.csie.ntu.edu.tw/~cjlin/libsvmtools/datasets/binary/australian_scale",
+     "data/australian_scale")
+
+
 
 # add the prepackaged wrapper 
 addSVMPackage("LIBSVM", wrapperPath =  file.path (path.package("SVMBridge"), "wrapper"))

### 2016-03-03 R/LLSVM_wrapper.R
fix #13

@@ -39,8 +39,8 @@ createTrainingArguments.LLSVM = function (x,
 		sprintf("-L %.16f", (1.0 / (n * cost))),
 		sprintf("-g %.16f", 2 * gamma),
 		extraParameter,
-		bashEscape (trainDataFile),
-		bashEscape (modelFile)
+	    shQuote (trainDataFile),
+	    shQuote (modelFile)
 	)
 
 	return (args)
@@ -51,9 +51,9 @@ createTrainingArguments.LLSVM = function (x,
 createTestArguments.LLSVM = function (x, testDataFile = NULL, modelFile = NULL, predictionsFile = NULL, verbose = FALSE, ...) {
     args = c(
 		"-v 1",
-		bashEscape (testDataFile),
-		bashEscape (modelFile),
-		bashEscape (predictionsFile)
+	    shQuote (testDataFile),
+	    shQuote (modelFile),
+	    shQuote (predictionsFile)
 	)
 
 	return (args)

### 2016-03-03 R/bashEscape.R

@@ -1,30 +1,0 @@
-#
-# SVMBridge
-#		(C) 2015, by Aydin Demircioglu
-#
-# SVMBridge is free software: you can redistribute it and/or modify
-# it under the terms of the GNU Lesser General Public License as published
-# by the Free Software Foundation, either version 3 of the License, or
-# (at your option) any later version.
-#
-# SVMBridge is distributed in the hope that it will be useful,
-# but WITHOUT ANY WARRANTY; without even the implied warranty of
-# MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
-# GNU Lesser General Public License for more details.
-#
-# Please do not use this software to destroy or spy on people, environment or things.
-# All negative use is prohibited.
-#
-
-
-#' Escape a filename so it can be used on the command line.
-#'
-#' @param s		String/filename to escape.
-#' @return		Escaped string.
-#'
-#' @note		For now, only double quotes are added.
-
-bashEscape = function (s) {
-	y = paste0('"', s, '"')
-	return (y)
-}

### 2016-03-03 inst/wrapper/BSGD_wrapper.R

@@ -39,8 +39,8 @@ createTrainingArguments.BSGD = function (x,
 		sprintf("-e %.16f", epochs ),
 		sprintf("-g %.16f", 2 * gamma),
 		extraParameter,
-		trainDataFile,
-		modelFile
+		shQuote (trainDataFile),
+		shQuote (modelFile)
 	)
 	return (args)
 }
@@ -50,8 +50,8 @@ createTrainingArguments.BSGD = function (x,
 createTestArguments.BSGD = function (x, testDataFile = NULL, modelFile = NULL, predictionsFile = NULL, verbose = FALSE, ...) {
     args = c(
 		"-v 1",
-		testDataFile,
-		modelFile,
+		shQuote (testDataFile),
+		shQuote (modelFile),
 		predictionsFile
 	)
 

### 2016-03-03 inst/wrapper/BVM_wrapper.R

@@ -19,15 +19,15 @@
 
 
 createTrainingArguments.BVM = function (x,
-					trainDataFile = "",
-                                            modelFile = "",
-                                            extraParameter = "",
-                                            primalTime = 10,
-                                            wallTime = 8*60,
-                                            kernelCacheSize = 1024,
-                                            cost = 1,
-                                            gamma = 1,
-                                            epsilon = 0.001, ...) {
+    trainDataFile = "",
+    modelFile = "",
+    extraParameter = "",
+    primalTime = 10,
+    wallTime = 8*60,
+    kernelCacheSize = 1024,
+    cost = 1,
+    gamma = 1,
+    epsilon = 0.001, ...) {
 
     args = c(
         "-s 6",                         # CVM = 6, BVM = 9
@@ -37,8 +37,8 @@ createTrainingArguments.BVM = function (x,
         sprintf("-g %.16f", gamma),
         sprintf("-e %.16f", epsilon),
         extraParameter,
-        bashEscape(trainDataFile),
-        bashEscape(modelFile)
+        shQuote(trainDataFile),
+        shQuote(modelFile)
     )
 
     return (args)
@@ -48,9 +48,9 @@ createTrainingArguments.BVM = function (x,
 
 createTestArguments.BVM = function (x, testDataFile = NULL, modelFile = NULL, predictionsFile = NULL, verbose = FALSE, ...) {
     args = c(
-        bashEscape(testDataFile),
-        bashEscape(modelFile),
-        bashEscape(predictionsFile)
+        shQuote(testDataFile),
+        shQuote(modelFile),
+        shQuote(predictionsFile)
     )
     return (args)
 }

### 2016-03-03 inst/wrapper/LIBSVM_wrapper.R

@@ -109,8 +109,8 @@ createTrainingArguments.LIBSVM = function (
 			sprintf("-g %.16f", gamma),        # gamma
 			sprintf("-e %.16f", epsilon),      # epsilon tolerance
 			extraParameter,
-			trainDataFile,
-			modelFile
+			shQuote (trainDataFile),
+			shQuote (modelFile)
 		)
 		return (args)
 }
@@ -119,8 +119,8 @@ createTrainingArguments.LIBSVM = function (
 
 createTestArguments.LIBSVM = function (x, testDataFile = NULL, modelFile = NULL, predictionsFile = NULL, verbose = FALSE, ...) {
     args = c(
-        testDataFile,
-        modelFile,
+        shQuote (testDataFile),
+        shQuote (modelFile),
         predictionsFile
     )
     return (args)

### 2016-03-03 inst/wrapper/SVMperf_wrapper.R

@@ -96,8 +96,8 @@ createTrainingArguments.SVMperf = function (x,
 			# --r
 			# --s
 			extraParameter,
-			bashEscape(trainDataFile),
-			bashEscape(modelFile)
+		    shQuote(trainDataFile),
+		    shQuote(modelFile)
 		)
 	}
 
@@ -125,8 +125,8 @@ createTrainingArguments.SVMperf = function (x,
 			# --r
 			# --s
 			extraParameter,
-			bashEscape(trainDataFile),
-			bashEscape(modelFile)
+		    shQuote(trainDataFile),
+		    shQuote(modelFile)
 		)
 	}
 
@@ -154,8 +154,8 @@ createTrainingArguments.SVMperf = function (x,
 			# --r
 			# --s
 			extraParameter,
-			bashEscape(trainDataFile),
-			bashEscape(modelFile)
+		    shQuote(trainDataFile),
+		    shQuote(modelFile)
 		)
 	}
 
@@ -187,8 +187,8 @@ createTrainingArguments.SVMperf = function (x,
 			# --r
 			# --s
 			extraParameter,
-			bashEscape(trainDataFile),
-			bashEscape(modelFile)
+		    shQuote(trainDataFile),
+		    shQuote(modelFile)
 		)
 	}
 
@@ -216,8 +216,8 @@ createTrainingArguments.SVMperf = function (x,
 			# --r
 			# --s
 			extraParameter,
-			bashEscape(trainDataFile),
-			bashEscape(modelFile)
+		    shQuote(trainDataFile),
+		    shQuote(modelFile)
 		)
 	}
 
@@ -236,9 +236,9 @@ createTrainingArguments.SVMperf = function (x,
 
 createTestArguments.SVMperf = function (x, testDataFile = NULL, modelFile = NULL, predictionsFile = NULL, verbose = FALSE, ...) {
 	args = c(
-		testDataFile,
-		bashEscape (modelFile),
-		bashEscape (predictionsFile)
+		shQuote (testDataFile),
+	    shQuote (modelFile),
+	    shQuote (predictionsFile)
 	)
 
 	return (args)

### 2016-03-03 tests/testthat/bashEscapetests.R

@@ -1,0 +1,76 @@
+#
+# SVMBridge 
+#
+#		(C) 2015, by Aydin Demircioglu
+# 
+# SVMBridge is free software: you can redistribute it and/or modify
+# it under the terms of the GNU Lesser General Public License as published
+# by the Free Software Foundation, either version 3 of the License, or
+# (at your option) any later version.
+#
+# SVMBridge is distributed in the hope that it will be useful,
+# but WITHOUT ANY WARRANTY; without even the implied warranty of
+# MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
+# GNU Lesser General Public License for more details.
+#
+# Please do not use this software to destroy or spy on people, environment or things.
+# All negative use is prohibited.
+#
+
+
+bashEscapetests = function (solver, trainDataX, trainDataY, testDataX, testDataY, verbose) {
+
+    tdir = tempdir()
+    trainfile = file.path (tdir, "A B C.sparse")
+    testfile = file.path (tdir, "B C D.sparse")
+    modelfile = file.path (tdir, "D E F.sparse")
+    writeSparseData (X = trainDataX, Y = trainDataY, file = trainfile)
+    writeSparseData (X = testDataX, Y = testDataY, file = testfile)
+
+	trainObj =  trainSVM(
+		method = solver,
+		trainDataFile = trainfile,
+		cost = 1, 
+		gamma = 1, 
+		epsilon = 0.01, 
+		rank = 32,
+		k = 32,
+		budget = 32,
+		modelFile = modelfile,
+		readModelFile = TRUE,
+		verbose = verbose
+	)  
+	
+	
+	if (Sys.info()["sysname"] == "Darwin") {
+		expErrors = list("LIBSVM" = c(20, 12), "LASVM" = c(23, 12), "BSGD" = c(10, 8), "SVMperf" = c(20, 12), 
+								"BVM" = c(20,15),  "CVM" = c(20, 15),  "LLSVM" = c(17, 15)) # maybe these things are random :/
+	} else if (Sys.info()["sysname"] == "Windows") {
[…]

### 2016-03-03 tests/testthat/test_generalWorkflow.R

@@ -33,7 +33,7 @@ test_that("general workflow works with binary packages.", {
 	
 	testthat::skip_on_cran()
 	
-	# call without parameters to get an exception
+	source ("bashEscapetests.R")
 	source ("cycletests.R")
 	source ("detectModeltests.R")
 	source ("downloadSoftware.R")
@@ -118,9 +118,15 @@ test_that("general workflow works with binary packages.", {
 	
 	for (solver in solvers) {
 		cat("\nWrapper test for solver", solver, ": ")
-		
 		wrappertests (solver, trainDataX, trainDataY, testDataX, testDataY, verbose)
 	}
+	
+	
+	for (solver in solvers) {
+		cat ("\nBash escape for solver", solver, ": ")
+		bashEscapetests (solver, trainDataX, trainDataY, testDataX, testDataY, verbose)
+	}
+	
 
 	for (solver in solvers) {
 		cat ("\nTest predictions for solver", solver, ": ")
@@ -145,6 +151,7 @@ test_that("general workflow works with binary packages.", {
 #		testthat::context (paste0(solver, " detect models."))
 		detectModeltests (solver, verbose)
 	}
+	
 	
 	
 		

### 2016-03-03 wrapper/CVM_wrapper.R

@@ -19,15 +19,16 @@
 
 
 createTrainingArguments.CVM = function (x,
-						trainDataFile = "",
-						modelFile = "",
-						extraParameter = "",
-						primalTime = 10,
-						wallTime = 8*60,
-						kernelCacheSize = 1024,
-						cost = 1,
-						gamma = 1,
-						epsilon = 0.001, ...) {
+    trainDataFile = "",
+    modelFile = "",
+    extraParameter = "",
+    primalTime = 10,
+    wallTime = 8*60,
+    kernelCacheSize = 1024,
+    cost = 1,
+    gamma = 1,
+    epsilon = 0.001, ...)
+{
     args = c(
         "-s 6",                         # CVM = 6, BVM = 9
         "-t 2",
@@ -36,8 +37,8 @@ createTrainingArguments.CVM = function (x,
         sprintf("-g %.16f", gamma),
         sprintf("-e %.16f", epsilon),
         extraParameter,
-        bashEscape (trainDataFile),
-        bashEscape (modelFile)
+        shQuote (trainDataFile),
+        shQuote (modelFile)
     )
 
     return (args)
@@ -47,9 +48,9 @@ createTrainingArguments.CVM = function (x,
 
 createTestArguments.CVM = function (x, testDataFile = NULL, modelFile = NULL, predictionsFile = NULL, verbose = FALSE, ...) {
     args = c(
-        bashEscape (testDataFile),
-        bashEscape (modelFile),
-        bashEscape (predictionsFile)
+        shQuote (testDataFile),
+        shQuote (modelFile),
+        shQuote (predictionsFile)
     )
     return (args)
 }

### 2016-03-03 wrapper/LASVM_wrapper.R

@@ -44,8 +44,8 @@ createTrainingArguments.LASVM = function (x,
 		sprintf("-e %.16f", epsilon),
 		sprintf("-p %.16f", epochs),
 		extraParameter,
-		trainDataFile,
-		modelFile
+		shQuote (trainDataFile),
+		shQuote (modelFile)
 	)
 
 	return (args)
@@ -55,8 +55,8 @@ createTrainingArguments.LASVM = function (x,
 
 createTestArguments.LASVM = function (x, testDataFile = NULL, modelFile = NULL, predictionsFile = NULL, verbose = FALSE, ...) {
     args = c(
-        testDataFile,
-        modelFile,
+        shQuote (testDataFile),
+        shQuote (modelFile),
         predictionsFile
     )
     return (args)
